### 2019-07-30 README.Rmd
moved over all material from private repository

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# Resources for Morey & Hoekstra (2019) <img src="man/figures/logo.png" align="right" alt="" width="120" />

Data, analysis code, and demonstrations for Morey & Hoekstra's (2019) statistical cognition experiment, which was run over the Christmas season in 2018. You can find more details at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/.

* [Installation](#installation)
* [Task demonstration](#task-demonstration)
* [Shiny app](#shiny-app)

## Installation

If you're using a recent version MacOS, you'll need to install [Xquartz](https://www.xquartz.org/index.html) for the fonts.

To fully use the package, you'll probably need to install several packages that are not on CRAN:

``` r
# install.packages('devtools') # if needed
devtools::install_github("crsh/papaja")
devtools::install_github("rstudio/gt")
```

Then you can install this package with

``` r
devtools::install_github("richarddmorey/Morey_Hoekstra_StatCognition")
```

## Task demonstration

A short video showing the task can be found under [`img/vignettes/santa_task_example.mp4`](https://github.com/richarddmorey/Morey_Hoekstra_StatCognition/blob/master/vignettes/img/santa_task_example.mp4)  (click 'View raw' to download).

You can find an [online version of the task demonstration](articles/task_demo.html).

Users of the package can also load a version of the experimental task from within R:

``` r
task_demo()
```

## Shiny app

Running the included shiny app will allow you to explore individual participants' response patterns.

``` r
explore_data()
```
You can access an already-deployed version of the shiny app at https://richarddmorey.shinyapps.io/explore/.

The app will let you:

* Search, sort, and filter the data set by experimental condition, response, etc.
* See the selected participant's interface at the poiint of decision (both experimental samples and random shuffle reports)
* See an animation showing the selected participant's behaviour over time, up to the point of decision
* Read the selected participant's responses to the questions about their strategy
* Search the strategy text using [POSIX regular expressions](https://stat.ethz.ch/R-manual/R-devel/library/base/html/regex.html), optionally (try "chance" to search for all instances of "chance"; "color|colour" will search for mentions of either of the two spelling variants of "color")

### 2019-07-30 R/MoreyHoekstra2019-package.R

@@ -1,0 +1,16 @@
+
+#' MoreyHoekstra2019: A package containing data, code, and demos for Morey & Hoekstra 2019
+#'
+#' @section MoreyHoekstra2019 functions:
+#' The MoreyHoekstra2019 functions ...
+#'
+#' @docType package
+#' @name MoreyHoekstra2019
+#' @importFrom graphics abline axis curve mtext par plot
+#' @importFrom graphics points polygon rasterImage rect
+#' @importFrom graphics segments text
+#' @importFrom stats coef density dlogis dnorm
+#' @importFrom stats lm pchisq qlogis qnorm reorder
+#' @importFrom stats rmultinom time
+#' @importFrom utils URLencode read.csv read.table
+NULL

### 2019-07-30 R/christmas_stats_public-data.R

@@ -1,0 +1,127 @@
+
+#' Cleaned, participant-level data
+#'
+#' Participant-level data for the Christmas statistical
+#' cognition experiment
+#'
+#' \itemize{
+#' \item duration: Qualtrics-reported time taken in the experiment (seconds)
+#' \item id: participant id
+#' \item consent: Response to informed consent
+#' \item times: "Have you participated previously"?
+#' \item shuffle: "Do you understand use of shuffle reports?"
+#' \item shuffle_other" (text) response for "Other" responses to `shuffle`
+#' \item response: primary response: which team is faster?
+#' \item confidence: participant's confidence in their `response`
+#' \item salient_factors: (text) "What facts or observations were most salient in coming to the conclusion that you did?"
+#' \item expt_strategy: (text) "Please describe your experimental strategy, if any, in your own words."
+#' \item shuffle_desc: (text) "Did you make use of the "random shuffle reports"? If so, how?"
+#' \item is_science: "Is your work in a field that would typically be considered scientific?"
+#' \item is_science_other: (text) response for "Other" responses to `is_science`
+#' \item education: highest level of formal education in scientific field
+#' \item formal_training: years of formal statistical training
+#' \item how_use: "How do statistics play a role in your work?" Concatenated string of responses
+#' \item field: "In what applied field(s) do you use statistics?" Concatenated string of responses
+#' \item preferred: "What sort of inferential procedures would you typically prefer?"
+#' \item preferred_other: (text) response for "Other" responses to `preferred`
+#' \item sig_testing: "What is your opinion about statistical significance testing?" Concatenated string of responses
+#' \item mobile: Qualtrics flag for mobile browsers. Should be blank for all participants
+#' \item evidence_power: Randomly-assigned transformation power. 3 or 7
+#' \item effect_size: "true" effect size in standard deviation units, hidden from the participant
+#' \item finished_practice_timestamp: javascript timestamp reported by browser when participant ended the instructions. Indexed to initial_timestamp
+#' \item downloaded_expt: Did the participant download their experimental samples?
+#' \item downloaded_null: Did the participant download their null samples?
+#' \item initial_timestamp: initial timestamp (should be 0, start of experiment)
+#' \item true_winner: Which "team" was the true winner? Corresponds to sign of effect_size
+#' \item true_null: Was the effect size 0?
[…]

### 2019-07-30 R/compile_complete_dataset.R

@@ -1,0 +1,273 @@
+
+#' Complete data set after extracting samples and making some columns "neater"
+#'
+#'
+#' @return List containing two data frames:
+#' compiled_data (the participant summaries) and samples (participants'
+#' clicks in the experimental task)
+#' @export
+#' @import dplyr
+#' @importFrom urltools url_decode
+#' @importFrom rjson fromJSON
+#' @importFrom purrr map map_int
+#'
+compile_complete_dataset <- function() {
+
+  dat <- load_complete_dataset()
+  # Convert columns to appropriate types ----------------------------------------------------------
+
+  dat$effect_size = as.numeric(dat$effect_size)
+  dat$evidence_power = as.integer(dat$evidence_power)
+  dat$formal_training = as.numeric(dat$formal_training)
+  dat$initial_timestamp <- as.numeric(dat$initial_timestamp)
+  dat$finished_practice_timestamp <-
+    as.numeric(dat$finished_practice_timestamp)
+
+  # Recode primary response ----------------------------------------------------------
+
+  dat %>% mutate(
+    response = case_when(
+      response == "" ~ NA_character_,
+      response == "...I will tell him that I conclude that the JINGLIES are faster when making ${e://Field/toy_name}.\"" ~ "jinglies",
+      response == "...I will tell him that I conclude that the SPARKLIES are faster when making ${e://Field/toy_name}.\"" ~ "sparklies",
+      response == "...I will tell him that I CANNOT DETECT A DIFFERENCE between the two teams when making ${e://Field/toy_name}.\"" ~ "no_detect",
+      response == "...I will tell him that I conclude that BOTH TEAMS ARE THE SAME SPEED when making ${e://Field/toy_name}.\"" ~ "same",
+      response == "...I will tell him that I got bored and want to leave.\"" ~ "bored"
+    ),
+    true_winner = case_when(
+      sign(effect_size) == 0 ~ "null",
+      sign(effect_size) == 1 ~ "jinglies",
+      sign(effect_size) == -1 ~ "sparklies"
+    ),
+    true_null = case_when(true_winner == "null" ~ TRUE,
+                          TRUE ~ FALSE),
+    response_null = case_when(
+      response %in% c("same", "no_detect") ~ TRUE,
+      response %in% c("jinglies", "sparklies") ~ FALSE,
+      response == "bored" ~ NA
+    ),
+    response_alt = !response_null,
[…]

### 2019-07-30 R/explore_data.R

@@ -1,0 +1,12 @@
+
+
+#' Run shiny app to explore the (cleaned) data set
+#'
+#' @return Returns the result of shiny::runApp()
+#' @export
+#' @importFrom shiny runApp
+#'
+explore_data <- function(){
+  shiny::runApp(system.file("shiny/explore",
+                          package = "MoreyHoekstra2019"))
+}

### 2019-07-30 R/fig_confidence.R

@@ -1,0 +1,135 @@
+
+#' Response confidence ratings
+#'
+#' @param which_resp Which responses to plot (responses indicating null or alt)
+#' @param shade_limit N for column below which will be shaded to indicate interpretability
+#' problems from small N
+#' @param filter_func function to filter the data before plotting
+#'
+#' @return Invisibly returns NULL
+#' @import dplyr
+#' @importFrom viridis viridis
+#' @importFrom stringr str_pad
+#'
+#' @export
+#'
+#' @examples
+#' fig_confidence( which_resp = "alt" )
+fig_confidence <- function(which_resp = c("null", "alt"),
+                           shade_limit = 20,
+                           filter_func = function(data, ...) return(data)) {
+  which_resp = which_resp[1]
+  stopifnot(which_resp %in% c("null", "alt"))
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  x = table(
+    dat$confidence,
+    abs(dat$effect_size),
+    dat$response %in% c("jinglies", "sparklies")
+  )
+  x0 = lapply(1:2, function(i)
+    t(t(x[, , i]) / colSums(x[, , i])))
+  names(x0) = c("Reported no detection/no difference", "Detected difference")
+
+  cx = .7
+  par(
+    mar = c(5.1, 11.1, 4.1, 2.1),
+    cex = cx,
+    cex.axis = .8,
+    cex.lab = 1
+  )
+
+  scaling = .4
+  cols = viridis::viridis(4)
+
+  d = as.numeric(colnames(x))
+  J = length(d)
+  mids = (d[-J] + d[-1]) / 2
+
+  if (which_resp == "null") {
+    i = 1
+  } else{
+    i = 2
+  }
+
+  plot(
+    d,
+    d,
+    ty = 'n',
+    ylim = c(.5, 4.5),
+    axes = FALSE,
+    xlab = "Effect size",
+    ylab = "",
+    xlim = c(d[1] - (mids[1] - d[1]), d[J] + (d[J] - mids[J - 1])),
+    xaxs = "i"
+  )
+  abline(v = mids, col = rgb(0, 0, 0, 1))
+  pin = par()$pin
+  usr = par()$usr
+  x.s = diff(usr[1:2]) / pin[1]
+  y.s = diff(usr[3:4]) / pin[2]
+  asp.s = x.s / y.s
+
[…]

### 2019-07-30 R/fig_desc_evidence.R

@@ -1,0 +1,210 @@
+#' Show the transformation between the z/p value and the visual location
+#'
+#' @return Invisibly returns NULL
+#' @importFrom viridis viridis
+#'
+#' @export
+#'
+#' @examples
+#' fig_desc_evidence_transform()
+fig_desc_evidence_transform <- function() {
+  cols = viridis::viridis(3)
+
+  par(las = 1)
+  curve(
+    evp(x, 3),
+    0,
+    1,
+    500,
+    type = "n",
+    xaxs = "i",
+    yaxs = "i",
+    ylab = "Evidence / x location",
+    xlab = "One-sided p value [Pr(Z<z)]"
+  )
+  rect(0, -1, .025, 1, border = NA, col = rgb(0, 0, 0, .2))
+  rect(.975, -1, 1, 1, border = NA, col = rgb(0, 0, 0, .2))
+  curve(evp(x, 3),
+        0,
+        1,
+        500,
+        lwd = 2,
+        col = cols[1],
+        add = TRUE)
+  curve(
+    evp(x, 7),
+    0,
+    1,
+    500,
+    add = TRUE,
+    lwd = 2,
+    col = cols[2],
+    lty = 2
+  )
+  text(.2, 0, "q = 7", col = cols[2], cex = 1.3)
+  text(.2, -.4, "q = 3", col = cols[1], cex = 1.3)
+
+  axis(3, at = pnorm(seq(-2,2,1)), lab = seq(-2,2,1))
+  mtext("Z test statistic", 3, 2.5, adj = .5)
+
+  invisible(NULL)
+
+}
+
+
+
+#' Show distributions of the "evidence" as a function of sample size, evidence scale/power, and true effect size
+#'
+#' @param n sample sizes (single value or vector)
+#' @param q evidence scale/power (single value or vector)
+#' @param d true effect size (single value or vector)
+#' @param letter Label to add to the graph
+#' @param scale_y_on How to choose a scale for the y axis; see details
+#' @param d_scale Scale multiplier to be applied to effect sizes
+#'
+#' @details \code{scale_y_on} can be one of several values:
+#' \itemize{
+#' \item NULL: scale on the the median maximum density value across distributions
+#' \item 0: scale on the maximum density value across distributions
[…]

### 2019-07-30 R/fig_edu_training.R

@@ -1,0 +1,97 @@
+
+
+#' Distribution of attained educational level in the sample
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import ggplot2
+#' @import dplyr
+#' @export
+#'
+#' @examples
+#' fig_edu_training_edu()
+fig_edu_training_edu <- function(filter_func = function(data, ...) return(data),
+                                 print_plot = TRUE) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    ggplot(aes(x = education)) +
+    geom_bar(mapping = aes(x = education, y = ..prop.., group = 1)) +
+    geom_text(stat = "count",
+              aes(y = ..prop.., label = ..count.., group = 1),
+              hjust = -.2) +
+    ylim(0, 1) +
+    #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
+    theme_bw() +
+    theme(
+      panel.grid.major.y = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      #panel.background = element_blank(),
+      text = element_text()
+    ) +
+    geom_hline(yintercept = c(0, 1)) +
+    xlab("Highest education") +
+    ylab("Proportion of participants") +
+    coord_flip() -> g
+
+  if(print_plot){
+    print(g)
+  }
+
+  invisible(g)
+
+}
+
+#' Distribution of reported number of years of statistical training in the sample
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import ggplot2
+#' @import dplyr
+#' @export
+#'
+#' @examples
+#' fig_edu_training_stats()
+fig_edu_training_stats <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    summarise(missing = round(100 * mean(is.na(formal_training)), 1)) %>%
+    unlist() %>% paste0("No response: ", . , "%") -> missing
+
+  dat %>% ggplot(aes(x = formal_training)) +
[…]

### 2019-07-30 R/fig_error_rates.R

@@ -1,0 +1,174 @@
+
+
+#' Error/correct response rates as a function of effect size
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#' @importFrom tidyr complete
+#' @export
+#'
+#' @examples
+#' fig_error_rates()
+fig_error_rates <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  summaries <-
+    dat %>%
+    mutate(effect_size = abs(effect_size)) %>%
+    group_by(effect_size, true_winner, response) %>%
+    tally() %>%
+    complete(effect_size, true_winner, response, fill = list(n = 0)) %>%
+    ungroup() %>%
+    filter(response != "bored") %>%
+    mutate(
+      dec_type = case_when(
+        true_winner != "null" &
+          true_winner != response &
+          response %in% c("jinglies", "sparklies") ~ "sign_error",
+        response %in% c("jinglies", "sparklies") ~ "report_difference",
+        response == "same" ~ "same",
+        response == "no_detect" ~ "no_detect"
+      )
+    ) %>%
+    group_by(dec_type, effect_size) %>%
+    tally( wt = n ) %>%
+    complete(dec_type, effect_size, fill = list(n = 0)) %>%
+    group_by(effect_size) %>%
+    mutate(total = sum(n))
+
+  summaries_all_null <-
+    summaries %>%
+    mutate(dec_type = ifelse(dec_type %in% c("same", "no_detect"), "null", dec_type)) %>%
+    filter(dec_type == "null") %>%
+    group_by(dec_type, effect_size) %>%
+    summarize(n = sum(n), total = first(total))
+
+  rbind(summaries, summaries_all_null) %>%
+    mutate(
+      p = n / total,
+      stderr = sqrt(p * (1 - p) / total),
+      lo = qlogis(p),
+      lo.stderr = stderr / dlogis(qlogis(p))
+    ) -> summaries
+
+  summaries %>%
+    filter(dec_type != "same") %>%
+    mutate(
[…]

### 2019-07-30 R/fig_evidence_p_vals.R

@@ -1,0 +1,111 @@
+
+# load it so everything can use it
+ima <- png::readPNG(system.file("pkg_html/img/clean_interface.png",
+                                package = "MoreyHoekstra2019"))
+
+#' Generate image to show the interface with p values overlaid
+#'
+#' @param q Evidence scale/power
+#' @param ps vector of p values to plot
+#' @param letter label to add to plot
+#'
+#' @return Returns a file path to the generated PNG file
+#' @importFrom viridis magma
+#' @importFrom png readPNG
+#' @importFrom showtext showtext_begin showtext_end
+#'
+#' @export
+#'
+#' @examples
+#' fig_evidence_p_vals(q = 7)
+fig_evidence_p_vals <- function(q,
+                                ps = c(.005, .025, .05, .125, .25, .5, .75, .875, .95, .975, .995),
+                                letter = ""){
+
+  tf = tempfile(fileext = ".png")
+
+  # ima is a global variable
+  img_dim = c(dim(ima)[2], dim(ima)[1])
+  res = 144
+
+  lower_left = c(239, img_dim[2] - 929)
+  lower_right = c(2028, lower_left[2])
+  upper_left = c(lower_left[1], img_dim[2] - 36)
+  plot_width = lower_right[1] - lower_left[1]
+  plot_height = upper_left[2] - lower_left[2]
+
+
+  evs <- q.ev(ps, d = 0, n = 10, q = q)
+  locs <- plot_width / 2 * evs + lower_left[1] + plot_width / 2
+  locs0 <- c(lower_left[1], locs, lower_right[1])
+
+  cols = rev(viridis::plasma(ceiling(length(locs) / 2), alpha = .05))
+
+  png(
+    filename = tf,
+    width = img_dim[1],
+    height = img_dim[2],
+    units = "px",
+    res = res
+  )
+
+  showtext::showtext_begin()
+
+  #Set up the plot area
+  par(mar = c(0, 0, 0, 0),
+      pin = img_dim / res)
+  plot(
+    0,
+    0,
+    type = 'n',
+    main = "",
+    xlab = "",
+    ylab = "",
+    axes = FALSE,
+    xaxs = "i",
+    yaxs = "i",
+    xlim = c(0, img_dim[1]),
+    ylim = c(0, img_dim[2])
+  )
+
[…]

### 2019-07-30 R/fig_evidence_scale.R

@@ -1,0 +1,338 @@
+
+
+
+
+#' Number of samples as a function of sample type, evidence scale
+#' and true effect size
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#' @importFrom purrr map map_df
+#' @importFrom gridExtra grid.arrange
+#' @importFrom MASS rlm
+#' @importFrom scales hue_pal
+#' @export
+#'
+#' @examples
+#' fig_evidence_scale_boxes()
+fig_evidence_scale_boxes <- function(
+        filter_func = function(data, ...) return(data),
+        print_plot = TRUE
+        ) {
+        dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+                filter_func() %>%
+                mutate(
+                        effect_size = abs(effect_size),
+                        ln_null = log10(n_null),
+                        ln_expt = log10(n_expt)
+                )
+
+
+        cnull <- dat %>%
+                split(.$evidence_power) %>%
+                map( ~ MASS::rlm(ln_null ~ effect_size, data = .)) %>%
+                map_df(coef)
+
+
+        cexpt <- dat %>%
+                split(.$evidence_power) %>%
+                map( ~ MASS::rlm(ln_expt ~ effect_size, data = .)) %>%
+                map_df(coef)
+
+        # grouped boxplot
+        p1 = ggplot(dat,
+                    aes(
+                            x = effect_size,
+                            y = n_null,
+                            group = interaction(evidence_power, effect_size),
+                            fill = factor(evidence_power)
+                    )) +
+                geom_boxplot() +
+                geom_abline(
+                        intercept = cnull$`3`[1],
+                        slope = cnull$`3`[2],
+                        color = scales::hue_pal()(2)[1]
+                ) +
+                geom_abline(
+                        intercept = cnull$`7`[1],
+                        slope = cnull$`7`[2],
+                        color = scales::hue_pal()(2)[2]
+                ) +
+                #geom_point(position=position_dodge(width=0.06),aes(group=interaction(evidence_power, effect_size))) +
+                scale_y_continuous(trans = 'log10') +
+                theme(legend.position = "none") +
[…]

### 2019-07-30 R/fig_field.R

@@ -1,0 +1,63 @@
+
+
+#' Distribution of self-reported scientific field in the sample
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#'
+#' @export
+#'
+#' @examples
+#' fig_field()
+fig_field <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    select(starts_with("field."), -matches("wordcount") ) %>%
+    mutate(field.other = !is.na(field.other)) %>%
+    mutate("field.no response" = rowSums(.) == 0) %>%
+    summarise_all(mean) %>%
+    unlist() %>% sort() -> summaries
+
+  tibble(
+    Field = substring(names(summaries), 7),
+    count = summaries * nrow(dat),
+    prop = summaries,
+    "No response" = factor(names(summaries) == "field.no response")
+  ) %>%
+    mutate(order = case_when(`No response` == TRUE ~ -Inf,
+                             TRUE ~ prop)) %>% ggplot(aes(
+                               x = reorder(Field, order),
+                               y = prop,
+                               fill = `No response`
+                             )) +
+    geom_bar(stat = "identity") +
+    xlab("Field") + ylab("Proportion of participants") +
+    ylim(0, 1) +
+    theme_bw() +
+    theme(legend.position = "none") +
+    geom_hline(yintercept = c(0, 1)) +
+    geom_text(
+      stat = "identity",
+      aes(
+        y = summaries,
+        label = summaries * nrow(dat),
+        group = 1
+      ),
+      hjust = -.2
+    ) +
+    coord_flip() -> g
+
+  if(print_plot) print(g)
+
+  invisible(g)
+
+}

### 2019-07-30 R/fig_field_scientific.R

@@ -1,0 +1,46 @@
+
+#' Participants' reports of whether others would call their field scientific
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import ggplot2
+#' @import dplyr
+#' @export
+#'
+#' @examples
+#' fig_field_scientific()
+fig_field_scientific <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+
+  dat %>%
+    ggplot(aes(x = is_science)) +
+    geom_bar(mapping = aes(x = is_science, y = ..prop.., group = 1)) +
+    geom_text(stat = "count",
+              aes(y = ..prop.., label = ..count.., group = 1),
+              hjust = -.2) +
+    ylim(0, 1) +
+    #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
+    theme_bw() +
+    theme(
+      panel.grid.major.y = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      #panel.background = element_blank(),
+      text = element_text()
+    ) +
+    geom_hline(yintercept = c(0, 1)) +
+    xlab("Field is considered scientific") +
+    ylab("Proportion of participants") +
+    coord_flip() -> g
+
+  if(print_plot) print(g)
+
+  invisible(g)
+
+}

### 2019-07-30 R/fig_how_use_stat.R

@@ -1,0 +1,67 @@
+
+
+#' Distribution of self-reported ways participants use statistics
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#'
+#' @export
+#' @examples
+#' fig_how_use_stat()
+fig_how_use_stat <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+
+  dat %>%
+    select(starts_with("how_use."), -matches("wordcount") ) %>%
+    mutate(how_use.other = !is.na(how_use.other)) -> use_cols
+
+  colnames(use_cols) = substring(colnames(use_cols), 9)
+
+  use_cols %>%
+    mutate("no response" = rowSums(.) == 0) %>%
+    summarise_all(mean) %>%
+    unlist() %>% sort() -> summaries
+
+  tibble(
+    How = names(summaries),
+    count = summaries * nrow(dat),
+    prop = summaries,
+    "No response" = factor(names(summaries) == "no response")
+  ) %>%
+    mutate(order = case_when(`No response` == TRUE ~ -Inf,
+                             TRUE ~ prop)) %>% ggplot(aes(
+                               x = reorder(How, order),
+                               y = prop,
+                               fill = `No response`
+                             )) +
+    geom_bar(stat = "identity") +
+    xlab("Reported use of statistics in position") + ylab("Proportion of participants") +
+    ylim(0, 1) +
+    theme_bw() +
+    theme(legend.position = "none") +
+    geom_hline(yintercept = c(0, 1)) +
+    geom_text(
+      stat = "identity",
+      aes(
+        y = summaries,
+        label = summaries * nrow(dat),
+        group = 1
+      ),
+      hjust = -.2
+    ) +
+    coord_flip() -> g
+
+  if(print_plot) print(g)
+
+  invisible(g)
+
+}

### 2019-07-30 R/fig_preferred_method.R

@@ -1,0 +1,56 @@
+
+
+#' Distribution of self-reported preferences for statistical methods
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#'
+#' @export
+#' @examples
+#' fig_preferred_method()
+fig_preferred_method <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    group_by(preferred) %>%
+    summarise(n = n()) %>%
+    mutate(total = sum(n),
+           p = n / total) %>%
+    ggplot(aes(x = preferred)) +
+    geom_bar(stat = "identity",
+             mapping = aes(
+               x = reorder(preferred, p),
+               y = p,
+               group = 1
+             )) +
+    geom_text(stat = "identity",
+              aes(y = p, label = n, group = 1),
+              hjust = -.2) +
+    ylim(0, 1) +
+    #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
+    theme_bw() +
+    theme(
+      panel.grid.major.y = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      #panel.background = element_blank(),
+      text = element_text()
+    ) +
+    geom_hline(yintercept = c(0, 1)) +
+    xlab("Preferred statistical method") +
+    ylab("Proportion of participants") +
+    coord_flip() -> g
+
+  if(print_plot)
+    print(g)
+
+  invisible(g)
+
+}

### 2019-07-30 R/fig_recreate_display.R

@@ -1,0 +1,121 @@
+
+
+
+
+#' Generate image to show a participant's sampling behaviour
+#'
+#' @param id ID of participant whose samples to show
+#' @param type String indicating type of samples ("expt" for experiments; "null" for random shuffle reports)
+#' @param ps p values to overlay on the display
+#'
+#' @return Returns a file path to the generated PNG file
+#' @import dplyr
+#' @importFrom magrittr %<>%
+#' @importFrom png readPNG
+#' @importFrom showtext showtext_begin showtext_end
+#' @importFrom grDevices dev.off png rgb
+#'
+#' @export
+#' @examples
+#' fig_recreate_display(id = "R_7V3ufsG6NYL27Ul")
+fig_recreate_display <- function(id,
+                                 type = "expt",
+                                 ps =
+                                   c(.005, .025, .05, .125, .25, .5, .75, .875, .95, .975, .995),
+                                 letter = "") {
+  tf = tempfile(fileext = ".png")
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants
+  samples <- MoreyHoekstra2019::christmas_stats_samples
+
+  q <- dat %>% filter(id == !!id) %>% select( evidence_power )
+  q <- q[1,1]
+
+  max_n_idx <- max(samples$n_idx)
+  samples %<>%
+    filter(id == !!id, type == !!type) %>%
+    mutate(ev = sgn * ev)
+
+  # ima is a global variable
+  img_dim = c(dim(ima)[2], dim(ima)[1])
+  res = 144
+
+  lower_left = c(239, img_dim[2] - 929)
+  lower_right = c(2028, lower_left[2])
+  upper_left = c(lower_left[1], img_dim[2] - 36)
+  plot_width = lower_right[1] - lower_left[1]
+  plot_height = upper_left[2] - lower_left[2]
+
+
+  if (length(ps)) {
+    evs <- q.ev(ps, d = 0, n = 10, q = q)
+    locs <- plot_width / 2 * evs + lower_left[1] + plot_width / 2
+    locs0 <- c(lower_left[1], locs, lower_right[1])
+
+    cols = rev(viridis::plasma(ceiling(length(locs) / 2), alpha = .05))
+  }
+
+  png(
+    filename = tf,
+    width = img_dim[1],
[…]

### 2019-07-30 R/fig_response_evidence.R

@@ -1,0 +1,160 @@
+
+
+#' Distribution of "evidence" in the display by response
+#'
+#' @param filter_func function to filter the data before plotting
+#' @return Invisibly returns NULL
+
+#' @import dplyr
+#' @importFrom purrr map map_dbl
+#'
+#' @export
+#' @examples
+#' fig_response_evidence()
+fig_response_evidence <- function(filter_func = function(data, ...) return(data)) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+  samples <- MoreyHoekstra2019::christmas_stats_samples %>%
+    filter(id %in% dat$id)
+
+
+  w_stats = samples %>%
+    mutate(ev0 = ev * sgn) %>%
+    split(.$id) %>%
+    map( ~ wilcox.test(ev0 ~ type, alternative = "less", data = .)) %>%
+    map_dbl("p.value")
+
+  w_stats = w_stats[match(dat$id, names(w_stats))]
+
+  j = dat$response == "jinglies"
+  s = dat$response == "sparklies"
+  n = dat$response %in% c("same", "no_detect")
+
+  dns = density(qlogis(w_stats[n]), na.rm = TRUE)
+
+  par(cex.axis = .8, cex.lab = 1)
+
+  plot(
+    dns,
+    ty = 'n',
+    xlim = c(-40, 40),
+    axes = FALSE,
+    ylim = c(-.03, max(dns$y) * 1.0),
+    xlab = "Wilcoxon two-sided p value",
+    ylab = "",
+    main = ""
+  )
+  polygon(c(dns$x, rev(dns$x)),
+          c(dns$y, dns$y * 0),
+          col = rgb(0, 0, 0, .4),
+          border = NA)
+
+  expt = seq(-15, -3, 3)
+  x.val = qlogis((10 ^ expt / 2))
+  x.val = c(x.val, 0,-rev(x.val))
+  expt = c(expt, 0, rev(expt))
+  strng = sapply(expt, function(e)
+    substitute(10 ^ q, list(q = e)))
+  strng[[ceiling(length(strng) / 2)]] = 1
+  for (i in 1:length(expt))
+    axis(1, at = x.val[i], labels = strng[[i]])
+
+  points(
+    qlogis(w_stats[n]),
[…]

### 2019-07-30 R/fig_sampling_behavior.R

@@ -1,0 +1,87 @@
+
+
+
+#' Number of samples per participant for each activity, by effect size
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns a list containing the ggplot objects
+#' and the result of the robust regression analysis
+#' @import dplyr
+#' @import ggplot2
+#' @importFrom ggpubr ggarrange
+#' @importFrom MASS rlm
+#'
+#' @export
+#' @examples
+#' fig_sampling_behavior()
+fig_sampling_behavior <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ) {
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    MASS::rlm(log(n_null) ~ abs(effect_size), data = .) -> rlm_obj_null
+
+  rlm_obj_null %>% coef() %>% `*`(., log10(exp(1))) -> coefs_null_lm
+
+  dat %>%
+    mutate(effect_size = abs(effect_size)) %>%
+    ggplot(aes(y = n_null, x = effect_size, group = effect_size)) +
+    geom_boxplot(outlier.shape = NA) +
+    geom_jitter(position = position_jitter(width = .01, height = 0), alpha = .4) +
+    geom_abline(
+      intercept = coefs_null_lm[1],
+      slope = coefs_null_lm[2],
+      linetype = 2,
+      col = "red"
+    ) +
+    scale_y_continuous(trans = "log10") +
+    ylab("Null samples") +
+    xlab("Effect size") -> g1
+
+  dat %>%
+    MASS::rlm(log(n_expt) ~ abs(effect_size), data = .) -> rlm_obj_expt
+
+  rlm_obj_expt %>% coef() %>% `*`(., log10(exp(1))) -> coefs_expt_lm
+
+  dat %>%
+    mutate(effect_size = abs(effect_size)) %>%
+    ggplot(aes(y = n_expt, x = effect_size, group = effect_size)) +
+    geom_boxplot(outlier.shape = NA) +
+    geom_jitter(position = position_jitter(width = .01, height = 0), alpha = .4) +
+    geom_abline(
+      intercept = coefs_expt_lm[1],
+      slope = coefs_expt_lm[2],
+      linetype = 2,
[…]

### 2019-07-30 R/fig_shuffle_understanding.R

@@ -1,0 +1,47 @@
+
+
+#' Distribution of self-reported understanding of what usefulness of random shuffle reports before experiment
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import ggplot2
+#' @import dplyr
+#' @export
+#'
+#' @examples
+#' fig_shuffle_understanding()
+fig_shuffle_understanding <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ) {
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    ggplot(aes(x = shuffle)) +
+    geom_bar(mapping = aes(x = shuffle, y = ..prop.., group = 1)) +
+    geom_text(stat = "count",
+              aes(y = ..prop.., label = ..count.., group = 1),
+              hjust = -.2) +
+    ylim(0, 1) +
+    #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
+    theme_bw() +
+    theme(
+      panel.grid.major.y = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      #panel.background = element_blank(),
+      text = element_text()
+    ) +
+    geom_hline(yintercept = c(0, 1)) +
+    xlab("Understood use of shuffle reports") +
+    ylab("Proportion of participants") +
+    coord_flip() -> g
+
+  if(print_plot)
+    print(g)
+
+  invisible(g)
+
+}

### 2019-07-30 R/fig_sig_testing_opinion.R

@@ -1,0 +1,67 @@
+
+
+#' Distribution of opinions about significance testing in sample
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#'
+#' @export
+#' @examples
+#' fig_sig_testing_opinion()
+fig_sig_testing_opinion <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ){
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    select(starts_with("sig_testing."), -matches("wordcount")) %>%
+    mutate(sig_testing.other = !is.na(sig_testing.other)) -> sig_cols
+
+  colnames(sig_cols) = substring(colnames(sig_cols), 13)
+
+  sig_cols %>%
+    mutate("no response" = rowSums(.) == 0) %>%
+    summarise_all(mean) %>%
+    unlist() %>% sort() -> summaries
+
+  tibble(
+    Opinion = names(summaries),
+    count = summaries * nrow(dat),
+    prop = summaries,
+    "No response" = factor(names(summaries) == "no response")
+  ) %>%
+    mutate(order = case_when(`No response` == TRUE ~ -Inf,
+                             TRUE ~ prop)) %>% ggplot(aes(
+                               x = reorder(Opinion, order),
+                               y = prop,
+                               fill = `No response`
+                             )) +
+    geom_bar(stat = "identity") +
+    xlab("Opinion about significance testing") + ylab("Proportion of participants") +
+    ylim(0, 1) +
+    theme_bw() +
+    theme(legend.position = "none") +
+    geom_hline(yintercept = c(0, 1)) +
+    geom_text(
+      stat = "identity",
+      aes(
+        y = summaries,
+        label = summaries * nrow(dat),
+        group = 1
+      ),
+      hjust = -.2
+    ) +
+    coord_flip() -> g
+
+  if(print_plot)
+    print(g)
+
+  invisible(g)
+
+}

### 2019-07-30 R/fig_strategies.R

@@ -1,0 +1,69 @@
+
+
+#' Frequencies of self-reported strategies
+#'
+#' Coded responses to the three open-ended text questions
+#' about participants' strategies.
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#'
+#' @export
+#'
+#' @examples
+#' fig_strategies()
+fig_strategies <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+  ){
+
+  dat <- MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func()
+
+  dat %>%
+    mutate(text_strong = text_comparison | text_sampling_var,
+                  text_only_weak = (text_asymmetry | text_inc_asymmetry ) & !text_strong,
+                  text_none = !(text_strong | text_only_weak),
+                  text_missing = text_missing | text_irrelevant) %>%
+    mutate(text_response_type = case_when(
+      text_strong ~ "Strong",
+      text_only_weak ~ "Only weak",
+      text_none ~ "None",
+      TRUE ~ "."
+    ),
+    text_response_subtype = case_when(
+      text_no_shuffle ~ "...denies using shuffles at all",
+      text_missing ~ "...is missing/irrelevant",
+      TRUE ~ "other"
+    ),
+    text_response_type = factor(text_response_type, ordered = TRUE),
+    text_response_subtype = factor(text_response_subtype,
+                                   levels = c("...is missing/irrelevant","...denies using shuffles at all","other"),
+                                   ordered = TRUE)
+    ) %>%
+    count(text_response_type,
+          text_response_subtype,
+          .drop = FALSE) -> for_barplot
+
+
+  for_barplot %>%
+    rename(`Response...` = text_response_subtype) %>%
+    ggplot(aes(fill=`Response...`, y=n, x=text_response_type)) +
+    geom_bar( stat="identity", color="black") +
+    scale_y_continuous(expand = c(0, 0), limits = c(0, 400)) +
+    coord_flip() +
+    theme_classic() +
+    theme(legend.position = "top",
+          text = element_text(size=12),
+          plot.margin = unit(c(0,1,0,0), "cm")) +
+    xlab("Sig. test content in text") -> g
+
+    if(print_plot)
+      print(g)
+
+  invisible(g)
+
+}

### 2019-07-30 R/load_clean_dataset.R

@@ -1,0 +1,60 @@
+
+#' Load the data set and apply all cleaning steps
+#'
+#'
+#' @return Returns a list of four objects:
+#' \itemize{
+#' \item clean_data: data frame with cleaned data, per participant
+#' \item samples: Each click for each participant in the experimental task
+#' \item eliminations: number of participants that fail to meet each criterion
+#' \item original rows: the number of rows in the data set before cleaning
+#' }
+#' @export
+#' @import dplyr
+#' @importFrom purrr map_lgl pmap_lgl
+#' @importFrom magrittr %<>%
+#'
+load_clean_dataset <- function(){
+
+  data_list <- compile_complete_dataset()
+
+  dat <- data_list$compiled_data
+  samples <- data_list$samples
+
+  original_rows <- nrow(dat)
+
+  # Perform main data filtering ---------------------------------------------------
+
+  # Perform data filtering by elimination criteria
+  dat %<>%
+    mutate(
+      exclusion_firsttime = times != "No, this is my FIRST TIME doing this survey.",
+      exclusion_engaged = (n_null == 0 | n_expt == 0),
+      exclusion_isscience = is_science == "No",
+      exclusion_education = education == "No university"
+      )
+
+  # Compute number of eliminations
+  dat %>%
+    select( matches("exclusion_") ) %>%
+    mutate_all( function(z){
+      z[is.na(z)] = TRUE
+      z }
+      ) %>%
+    mutate(exclusion_any = pmap_lgl(., any )) -> excl_tbl
+
+  eliminations <- colSums( excl_tbl, na.rm = TRUE )
+
+  dat %<>%
+    filter( !excl_tbl$exclusion_any ) %>%
+    select( -matches("exclusion_") ) %>%
+    merge(MoreyHoekstra2019::text_coding, by = "id")
+
+  # Eliminate participants in record of sampling behaviour
+  samples %<>% filter(id %in% dat$id )
+
+  list( clean_data = dat,
+        samples = samples,
+        eliminations = eliminations,
+        original_rows = original_rows )
+}

### 2019-07-30 R/load_complete_dataset.R

@@ -1,0 +1,61 @@
+
+#' Read the complete data set from the (compressed) CSV
+#'
+#' @return Returns a data from containing the entire data set,
+#' with invalid rows (mobile users and non-consents) filtered out
+#' @export
+#' @import dplyr
+#' @importFrom utils unzip
+#' @importFrom tools file_path_sans_ext
+#'
+load_complete_dataset <- function(){
+
+  zip_file_path <- system.file("extdata/qualtrics_31-5-2019_20.40.csv.zip",
+                                 package = "MoreyHoekstra2019")
+  csv_dir_path <- tempdir()
+  csv_file_name <- tools::file_path_sans_ext(basename(zip_file_path))
+  csv_file_path <- file.path(csv_dir_path, csv_file_name)
+
+  md5_file_path <- system.file("md5/md5sum.txt",
+                               package = "MoreyHoekstra2019")
+  colnames_file_path <- system.file("extdata/colnames.csv",
+                               package = "MoreyHoekstra2019")
+
+  md5s <- read.table(md5_file_path, header = FALSE,
+                     row.names = 1, stringsAsFactors = FALSE)
+
+  # check md5 of zipped file
+  if( md5s["Original_zip",1] != tools::md5sum( zip_file_path ) )
+    stop( "Something is wrong; md5 of zip file does not check out! Aborting." )
+
+  utils::unzip(zip_file_path, exdir = csv_dir_path)
+
+  # check md5 of unzipped file
+  if( md5s["Original",1] != tools::md5sum( csv_file_path ) )
+    stop( "Something is wrong; md5 of csv file does not check out! Aborting." )
+
+  # Read in data twice to get header; qualtrics is weird
+  dat <- read.csv(csv_file_path, skip = 1, header = TRUE,
+                  stringsAsFactors = FALSE)
+  dat <- dat[-1,]
+  header <- read.csv(csv_file_path,skip = 1, header = TRUE)
+  colnames(dat) <- colnames(header)
+
+  # Clean columns/initial filter
+
+  # Rename columns for convenience and
+  # delete unnecessary columns
+  cnames <- read.csv(file = colnames_file_path,
+                     header = TRUE, stringsAsFactors = FALSE)
+  cnames0 <- cnames$names[cnames$keep]
+  names(cnames0) <- cnames$new_names[cnames$keep]
+
+  dat %>%
+    select( !!!cnames0 ) %>%
+    filter(mobile != "true") %>%
[…]

### 2019-07-30 R/load_text_responses.R

@@ -1,0 +1,31 @@
+
+#' Text responses  for each participant from the cleaned data
+#'
+#' @return A data frame containing the only text responses from the cleaned data set
+#' @export
+#' @import dplyr
+#'
+load_text_responses <- function(){
+
+  data_list <- load_clean_dataset()
+
+  dat <- data_list$clean_data
+  samples <- data_list$samples
+
+  colnames_file_path <- system.file("extdata/colnames.csv",
+                                    package = "MoreyHoekstra2019")
+  cnames <- read.csv(file = colnames_file_path,
+                     header = TRUE, stringsAsFactors = FALSE)
+
+  dat_names <- colnames(dat)
+
+  has_text = ( dat_names %in% cnames$new_names[ cnames$free_text ] |
+                 grepl("other", dat_names, fixed = TRUE) ) &
+    !( dat_names %in% "sig_testing.prefer_other" )
+
+  dat %>%
+    select( c("id", dat_names[ has_text ] ) ) -> data_text_responses
+
+  return(data_text_responses)
+
+}

### 2019-07-30 R/santaDisplay_widget.R

@@ -1,0 +1,75 @@
+#' HTML widget for displaying an animation of participant behavior
+#'
+#' @param value Participant ID to display
+#' @param label label for the widget
+#' @param tooltip tooltip for the widget
+#' @param width width of div
+#' @param height height of div
+#' @param elementId widget div element id
+#'
+#' @return An object of class \code{htmlwidget} for printing into HTML
+#' @export
+#' @import dplyr
+#' @import htmlwidgets
+#' @importFrom rjson toJSON
+#'
+santaDisplay <- function(value, label = "", tooltip = "", width = NULL, height = NULL, elementId = NULL)
+{
+
+  snippet_fn <- system.file("htmlwidgets/html/interface_snippet.html",
+                            package = "MoreyHoekstra2019")
+  snippet <- paste(readLines(snippet_fn, n = -1), collapse = "\n")
+
+  stopifnot( value %in% MoreyHoekstra2019::christmas_stats_participants$id )
+
+  df <- MoreyHoekstra2019::christmas_stats_samples %>%
+    filter(id == value) %>%
+    arrange(time)
+
+
+
+  # create widget
+  htmlwidgets::createWidget(
+    name = 'santaDisplay',
+    rjson::toJSON(list(df = df, snippet = snippet)),
+    width = width,
+    height = height,
+    package = 'MoreyHoekstra2019',
+    elementId = elementId,
+    sizingPolicy = htmlwidgets::sizingPolicy(
+      defaultWidth = 200,
+      defaultHeight = 200
+    )
+  )
+}
+
+
+#' Shiny bindings for santaDisplay
+#'
+#' Output and render functions for using santaDisplay within Shiny
+#' applications and interactive Rmd documents.
+#'
+#' @param outputId output variable to read from
+#' @param width,height Must be a valid CSS unit (like \code{'100\%'},
+#'   \code{'400px'}, \code{'auto'}) or a number, which will be coerced to a
+#'   string and have \code{'px'} appended.
+#' @param expr An expression that generates a ct_test
+#' @param env The environment in which to evaluate \code{expr}.
+#' @param quoted Is \code{expr} a quoted expression (with \code{quote()})? This
+#'   is useful if you want to save an expression in a variable.
+#'
+#' @name santaDisplay-shiny
+#'
+#' @import htmlwidgets
+#' @export
[…]

### 2019-07-30 R/task_demo.R

@@ -1,0 +1,61 @@
+#' Start a web server with a demo of the experimental task and open browser
+#'
+#' @param toy_name Name of the toy; determines effect size
+#' @param evidence_power Power for the p value transform
+#'
+#' @return Invisibly returns a list of the job number of the HTTP server and url
+#' @export
+#' @importFrom servr httd
+#'
+task_demo <- function(toy_name = NULL, evidence_power = NULL){
+
+  toy_names = c("whizbang balls",
+                "constructo bricks",
+                "rainbow clickers",
+                "doodle noodles",
+                "singing bling rings",
+                "brahma buddies",
+                "magic colorclay",
+                "moon-candy makers")
+
+  if(is.null(toy_name)){
+    probs = c(1/4, 3/4*rep(1/7, 7))
+    idx = which(rmultinom(1, 1, probs)>0)
+    toy_name = toy_names[idx]
+  }
+
+  if(is.null(evidence_power)){
+    evidence_power = sample(c(3,7), size = 1)
+  }
+
+  if( !(evidence_power %in% c(3,7) ) ){
+    stop("evidence_power must be either 3 or 7.")
+  }
+  if( !( toy_name %in% toy_names ) ){
+    stop("toy_name must be one of", toy_names,".")
+  }
+
+  sf = tempfile()
+  zz <- file(sf, open = "wt")
+  dn <- system.file("pkg_html/",
+                    package = "MoreyHoekstra2019")
+
+  initpath = paste0("/task.html?toy_name=", URLencode(toy_name), "&q=",evidence_power)
+
+  sink(file = zz, type = "message")
+  servr::httd(dir = dn, initpath = initpath, browse = FALSE)
+  sink(type = "message")
+  close(zz)
+  mesg <- readLines(sf, -1)
+  job <- regmatches(mesg[1],
+                    regexpr("[0-9]+", mesg[1], perl = TRUE)
+  )
+  url_actual <- regmatches(mesg[2],
+                           regexpr("http://.+$", mesg[2], perl = TRUE)
+  )
+  utils::browseURL(url_actual)
+  message(mesg)
+
+  invisible(c(job = job, url = url_actual))
+
+}

### 2019-07-30 inst/font/font_setup.R

@@ -1,0 +1,26 @@
+library(showtext)
+
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.showtext=TRUE,
+  dpi=300
+)
+
+font_paths(
+  system.file("font/", package = "MoreyHoekstra2019")
+)
+
+font_add("lato",
+         regular = "Lato-Light.ttf",
+         italic = "Lato-LightItalic.ttf",
+         bold = "Lato-Semibold.ttf",
+         bolditalic = "Lato-SemiboldItalic.ttf",
+         symbol = "Lato-LightItalic.ttf")
+
+knitr::knit_hooks$set(
+  lato = function(before, options, envir)
+    if (before) par(family = "lato")
+)
+
+knitr::opts_chunk$set(lato = TRUE)

### 2019-07-30 inst/htmlwidgets/lib/santa_utility.js

@@ -1,0 +1,190 @@
+
+function transformFunc(e){
+	return 1 - Math.pow(e, colorTransformPower);
+}
+
+function identityFunc(e){
+	return e;
+}
+
+function getColor(evidence, r, g, b){
+	r = (typeof r !== 'undefined') ?  r : 0;
+	g = (typeof g !== 'undefined') ?  g : 0;
+	b = (typeof b !== 'undefined') ?  b : 0;
+
+	var intensity = evidence;
+    var color = "rgba(" + r + "," + g + "," + b + ", " + intensity +")";
+    return color;
+}
+
+function getEvidence(d, z, n, idx, transformFunc){
+	transformFunc = (typeof transformFunc !== 'function') ?  identityFunc : transformFunc;
+	var p = std_n_cdf( z.nextGaussian() + d * Math.sqrt(n[idx]/2) );
+	var sgn = (p > 0.5) * 2 - 1;
+	p = (p < 0.5) ? p * 2 : (1 - p) * 2 ;
+  return { sgn: sgn, ev: transformFunc( p ) };
+}
+
+function drawX(x, y, r, w, h, red, green, blue, alpha, canvas) {
+
+    canvas.beginPath();
+    canvas.strokeStyle = "rgba("+red+","+green+","+blue+","+alpha+")";
+    canvas.lineWidth=1;
+
+    r = r * (w + h)/2;
+    x = x * w;
+    y = y * h;
+    canvas.moveTo(x - r, y - r);
+    canvas.lineTo(x + r, y + r);
+
+    canvas.moveTo(x + r, y - r);
+    canvas.lineTo(x - r, y + r);
+    canvas.stroke();
+}
+
+function point(x, y, r, w, h, red, green, blue, alpha, canvas, fill, stroke){
+  canvas.beginPath();
+  canvas.arc(x * w, y * h, r * (w + h)/2, 0, 2 * Math.PI, true);
+  canvas.fillStyle = "rgba("+red+","+green+","+blue+","+alpha+")";
+  canvas.strokeStyle = "rgba(0,0,0,.3)";
+  canvas.lineWidth=1;
+  if(fill) canvas.fill();
+  if(stroke) canvas.stroke();
+}
+
+function plot_new_point(d, z, n_idx, n, c, h, w){
+  var ctx = c.getContext("2d");
+
+  var n_ran_idx = 0;
+  var n_unif = 0;
+
[…]

### 2019-07-30 inst/htmlwidgets/santaDisplay.js

@@ -1,0 +1,134 @@
+HTMLWidgets.widget({
+
+  name: 'santaDisplay',
+  type: 'output',
+
+  initialize: function(el, width, height)
+  {
+
+  },
+
+
+  renderValue: function(el, x, obj) {
+
+        var ms_per_update = 200;
+
+        var interval = 0;
+
+        function increment_slider(){
+          var status = $( "#ani_play_button" ).data("status");
+
+          if(status === undefined){
+            clearInterval(interval);
+            interval = 0;
+            return;
+          }
+
+          var max = $( "#ani_slider_container" )
+            .slider( "option", "max");
+          var m = $( "#ani_slider_container" )
+            .slider( "option", "value" );
+
+          if(max>m){
+            $( "#ani_slider_container" )
+              .slider("value", m+1 )
+              .trigger("slidechange");
+          }else if( (max == m) ){
+            clearInterval(interval);
+            interval = 0;
+            $( "#ani_play_button" ).removeData("status");
+            $( "#ani_play_button" ).removeAttr("data-status");
+            $( "#ani_play_button" ).html("▶");
+          }
+        }
+
+        function reset_play_button(el){
+          var status = $( el ).data("status");
+
+          if(status === undefined){
+            $( el ).data("status", "playing");
+            interval = setInterval(increment_slider, ms_per_update);
+            $( "#ani_play_button" ).html("■");
+          }else{
+            $( el ).removeData("status");
+            $( el ).removeAttr("data-status");
+            $( "#ani_play_button" ).html("▶");
+            clearInterval(interval);
+            interval = 0;
+          }
+        }
+
+        x = JSON.parse(x);
+        el.innerHTML = x.snippet;
+
+        var df = x.df;
+        var n = df.id.length;
+        var type="";
+        var ev = 0;
+        var n_idx = 0;
+        var n_unif = 0;
+        var sgn = 0;
+
+        var c = $(".plot_canvas")[0];
+        var ctx = c.getContext("2d");
+        var h = c.offsetHeight;
+				var w = c.offsetWidth;
+
+        var blue = 0;
+        red = 255;
+        var green = 255 - red;
+
+        $( "#ani_slider_container" )
+        .css({ width: "75%"} )
+        .slider({
+          min: 0,
+          max: n
+        }).on( "slidechange", function( event, ui ) {
+          ctx.clearRect(0, 0, w, h);
[…]

### 2019-07-30 inst/pkg_html/js/getUrlParameter.js

@@ -1,0 +1,8 @@
+
+// Obtained from https://davidwalsh.name/query-string-javascript
+function getUrlParameter(name) {
+    name = name.replace(/[\[]/, '\\[').replace(/[\]]/, '\\]');
+    var regex = new RegExp('[\\?&]' + name + '=([^&#]*)');
+    var results = regex.exec(location.search);
+    return results === null ? '' : decodeURIComponent(results[1].replace(/\+/g, ' '));
+};

### 2019-07-30 inst/shiny/explore/server.R

@@ -1,0 +1,222 @@
+#
+# This is the server logic of a Shiny web application. You can run the
+# application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(dplyr)
+library(DT)
+library(purrr)
+library(gt)
+library(magrittr)
+
+ima <- MoreyHoekstra2019:::ima
+dat <- MoreyHoekstra2019::christmas_stats_participants
+samples <- MoreyHoekstra2019::christmas_stats_samples
+
+repl_text = c(
+    text_comparison = "Comparison to null",
+    text_asymmetry = "Symmetry/asymmetry",
+    text_sampling_var = "Shuffles as sampling variability/null,chance distribution",
+    text_inc_asymmetry = "Increasing asymmetry",
+    text_no_shuffle = "Explicitly said no use of random shuffles",
+    text_irrelevant = "Text irrelevant",
+    text_missing = "Text missing"
+)
+
+
+suppressWarnings({w_stats = samples %>%
+    mutate(ev0 = ev * sgn) %>%
+    split(.$id) %>%
+    map( ~ wilcox.test(ev0 ~ type, alternative = "less", data = .)) %>%
+    map_dbl("p.value")})
+
+w_stats <- w_stats[match(dat$id, names(w_stats))]
+
+dat$wilcoxon = ifelse(w_stats < .5,
+                      w_stats*2,
+                      (1 - w_stats)/2)
+
+values <- reactiveValues(dat = dat)
+
+shinyServer(function(input, output) {
+
+    output$santaDisplay <- MoreyHoekstra2019::renderSantaDisplay({
+        idx = input$dataTable_rows_selected
+
+        if(length(idx)==0 | nrow(values$dat)==0)
+            stop("Please select a row in the Data tab.")
+
+        id = values$dat[idx, "id"]
+
+        MoreyHoekstra2019::santaDisplay(id)
+    })
+
+    output$distributionsPlot <- renderPlot({
+        idx = input$dataTable_rows_selected
+
+        if(length(idx)==0 | nrow(values$dat)==0){
+            plot(c(0, 1), c(0, 1), ann = F, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
+            text(x = 0.5, y = 0.5, "Please select a row in the Data tab.",
+                 cex = 1.6, col = "black")
+            return(NULL)
+        }
+
[…]

### 2019-07-30 inst/shiny/explore/ui.R

@@ -1,0 +1,42 @@
+#
+# This is the user-interface definition of a Shiny web application. You can
+# run the application by clicking 'Run App' above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+
+# Define UI for application that draws a histogram
+shinyUI(fluidPage(
+
+    # Application title
+    titlePanel("Morey & Hoekstra 2019 Data"),
+    withMathJax(),
+
+    # Sidebar with a slider input for number of bins
+    sidebarLayout(
+        sidebarPanel(
+            selectInput("type", "Sample type in 'Display' tab",
+                        c(Experiments="expt",
+                          "Null samples" = "null") ),
+            checkboxInput("pvals", "p values in 'Display' tab",
+                          value = FALSE),
+            textInput("strategyTextFilter", "Filter by strategy text", width = 200)
+        ),
+
+        # Show a plot of the generated distribution
+        mainPanel(
+            tabsetPanel(
+                tabPanel("Data", DT::dataTableOutput("dataTable")),
+                tabPanel("Display", imageOutput("displayPlot")),
+                tabPanel("Animate", MoreyHoekstra2019::santaDisplayOutput(outputId = "santaDisplay")),
+                tabPanel("Strategy text", uiOutput("strategyText")),
+                tabPanel("Distributions", plotOutput("distributionsPlot"))#,
+                #tabPanel("Explanation", uiOutput("explText"))
+            )
+        )
+    )
+))

### 2019-08-01 README.Rmd
rework vignette building

@@
 
-A short video showing the task can be found under [`img/vignettes/santa_task_example.mp4`](https://github.com/richarddmorey/Morey_Hoekstra_StatCognition/blob/master/vignettes/img/santa_task_example.mp4)  (click 'View raw' to download).
+A short video showing the task can be found under [`inst/manuscript/img/santa_task_example.mp4`](https://github.com/richarddmorey/Morey_Hoekstra_StatCognition/blob/master/inst/manuscript/img/santa_task_example.mp4)  (click 'View raw' to download).
 

### 2019-08-02 inst/manuscript/manuscript.Rmd
updated vignette/website compilation

---
title: Use of significance test logic by scientists in a novel reasoning task
titlerunning: Statistical reasoning in scientists
authorrunning: Morey & Hoekstra
authors: 
- name: Richard D. Morey
  address: School of Psychology, Cardiff University
  email: moreyr@cardiff.ac.uk
- name: Rink Hoekstra
  address: Faculty of Behavioural and Social Sciences, University of Groningen
  email: r.hoekstra@rug.nl
keywords:
- cognitive science
- psychology
- science
- statistics
- statistical cognition
#PACS:

MSC:
- 91E10    
abstract: |
     Although statistical significance testing is one of the most widely-used techniques across science, previous research has suggested that scientists have a poor understanding of how it works. If scientists misunderstand one of their primary inferential tools the implications are dramatic: potentially unchecked, unjustified conclusions and wasted resources. Scientists' apparent difficulties with significance testing have led to calls for its abandonment or increased reliance on alternative tools, which would represent a substantial, untested, shift in scientific practice. However, if scientists' understanding of significance testing is truly as poor as thought, one could argue such drastic action is required. We show using a novel experimental method that scientists do, in fact, understand the logic of significance testing and can use it effectively. This suggests that scientists may not be as statistically-challenged as often believed, and that reforms should take this into account.
bibliography: "`r system.file('bib/bibfile.bib', package = 'MoreyHoekstra2019')`"
csl: nature.csl
output: rticles::springer_article
vignette: >
  %\VignetteIndexEntry{Manuscript}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
#nocite: | 
#  @Venables:Ripley:2002
header-includes: |
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{subfig}
  \renewcommand{\thesubfigure}{\Alph{subfigure}}
pkgdown:
  as_is: true
  extension: pdf
---

```{r include = FALSE}
source(
  system.file("font/font_setup.R", package = "MoreyHoekstra2019")
)

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      out.width = "0.95\\textwidth")

library(dplyr)
library(papaja)
library(tidyr)
library(broom)
```

Despite statistical significance testing (SST) use in a majority of research projects across fields, there is debate over whether scientists understand SST and can use it competently. Methodologists and statistical cognition researchers point to evidence from questionnaires and vignette studies to argue that researchers do not, in fact, grasp the core logic of SST. In one highly influential study of research psychologists, Oakes [-@Oakes:1986] presented six statements about a hypothetical significance test result to be categorized as true or false (e.g., "[The $p$ value provides] the probability of the null hypothesis being true"). Despite all of these statements being false, 97% of the research psychologists categorized at least one as true. Oakes argues that this shows that the participants have an "[un]sound understanding of the logic of the significance test" (p. 82).

Oakes' basic method and results have been replicated and extended with various groups, showing that students [@Falk:Greenbaum:1995], instructors [@Haller:Krauss:2002], and statisticians [@Lecoutre:etal:2003] all misinterpret SST results. Moreover, these misinterpretations are difficult to eliminate even through targeted interventions [@Kalinowski:etal:2008]. As a result, many have argued that use of SST should be discontinued or dramatically reduced, and may even contribute to wide-spread replication problems in the sciences [@Carver:1978;@Oakes:1986;@Fidler:2006;@The:2011;@Wasserstein:Lazar:2016].

The interpretation of studies of researchers' understanding of SST are limited, however, by their methodology. A typical study presents a vignette describing research results. Statistical results are offered to the participants (e.g., a $t$ statistic and $p$ value), who are then asked to explicitly give or endorse various interpretations that are thought to represent participants' understanding of SST. However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication). It is unclear how well vignette studies (including ones by the present authors; @Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results. Conceptual understanding and fluency with common representations are both important, but are distinct.

A second major piece of evidence for misunderstandings of SST logic is reasoning errors in published papers. [@Gelman:Stern:2006;@Hoekstra:etal:2006;@Nieuwenhuis:etal:2011;@Weisburd:etal:2003]. Like evidence from vignette studies, however, these errors are difficult to interpret as misunderstandings of SST logic *per se*. To avoid conflating problems with statistical reasoning with motivated scientific reasoning, one must test scientists' statistical reasoning where these incentives are not in play.

using a simulated experimental task, we tested researchers' and trainees' understanding and application of SST. Our key innovation allowing us to focus on SST was to prevent---or make difficult---the use of alternative strategies. If participants have poor understanding of SST, one would predict they would perform poorly and would not be sensitive to SST-related information. In fact, scientists performed well, explicitly reported using SST strategies, and were sensitive to SST-related information. Our results strongly suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.

# Testing reasoning by withholding information

In tests of perception --- e.g., of colorblindness [@Ishihara:1972] --- it is common to eliminate one cue (e.g., brightness) in order assess the ability to use another (e.g., color). If color is the only useful cue for reading a number, deficits in color vision make the number difficult to read. We adopt a similar strategy to test statistical reasoning: we eliminate numerical information from statistical results to test scientists' ability to interpret results with reference to a null sampling distribution (i.e., SST logic). Without numerical information, many other strategies and heuristics, such as confidence intervals or Bayesian inference, are difficult or impossible to apply.

Participants were scientists or trainees recruited via social media. Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of Christmas elves --- "Jinglies" or "Sparklies" --- could make more of a particular toy, where the true effect size was a randomly-assigned group difference between 0 and 1 standard deviations ($\delta$). The numerical information for an experiment, including sample sizes and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}). Participants could change the sample size per group for each experiment (increasing the time required to return a result), but did not know its numerical value. The interpretation of these colors and locations was unknown to the participants, aside from the monotone relationship with the statistical evidence.

```{r interface, fig.cap="The interface on which samples were reported. The $x$-axis monotonically (but nonlinearly) related to the strength of the statistical evidence ($z$ statistic) favoring one group; the $y$-axis is monotonically (but nonlinearly) related to the sample size. Underlying numerical values of the statistical evidence and sample sizes were unknown to the participant."}
fn = system.file("pkg_html/img/clean_interface.png", package = "MoreyHoekstra2019")
knitr::include_graphics(fn)

```

Participants were also offered the opportunity to sample "random shuffle reports" that were described as the results of experiments with random assignment of elves to groups: that is, the result of experiments in which the null hypothesis was true. These results took no time to return. Participants were not instructed how to use these samples, only that they might use them.

After sampling as many "experiments" and "random shuffle reports" as they liked, participants could report whether they believed Jinglies or Sparklies were better, that they could not detect a difference, that there was no difference, or that they were bored and wanted to stop. Following their decision they were asked several open-ended questions about their strategy. Our key questions are whether participants can effectively find the "truth", whether they report strategies consistent with SST, and whether their behaviour shows evidence of SST use.

```{r}
total_n = nrow(MoreyHoekstra2019::christmas_stats_participants)
```

Here, we report the results of `r total_n` scientists or trainees who completed the statistical reasoning task. We provide an interactive app for exploring participants' responses at https://richarddmorey.shinyapps.io/explore/.

# Success rates identifying effect sign

```{r error_rates, fig.cap="Decision error rates as a function of true effect size. The red solid line represents the rate at which participants {\\em correctly} determined the direction of the effect, except for at an effect size of 0, where it represents an incorrect claim of an effect. The shaded region around the solid line is one standard error wide. The difference between the black and the gray dashed lines shows the proportion of participants who decided that the two groups were the same (as opposed to failing to detect a difference)."}
MoreyHoekstra2019::fig_error_rates()
```

Decision error rates as a function of true effect size are shown in Figure \ref{fig:error_rates}.

```{r}
null_n =  sum(MoreyHoekstra2019::christmas_stats_participants$true_null)

null_resp_count = MoreyHoekstra2019::christmas_stats_participants %>% 
  filter(true_null) %>% 
  group_by(response) %>%
  summarise(n=n())

t1_err_count = (MoreyHoekstra2019::christmas_stats_participants %>% 
  filter(true_null, response_alt) %>% 
  summarise(n=n()))$n

null_same_count = (null_resp_count %>% filter(response=="same"))$n
null_nodet_count = (null_resp_count %>% filter(response=="no_detect"))$n

sign_err_count = nrow(MoreyHoekstra2019::christmas_stats_participants %>% 
  filter( (effect_size<0 & response == "jinglies") | (effect_size>0 & response == "sparklies")))

```

Of the `r null_n` participants for whom the null hypothesis was true (i.e. $\delta=0$), `r t1_err_count` participants (`r round(t1_err_count/null_n * 100, 0)`%) incorrectly indicated an effect. This is larger than the typically-accepted 5% false positive rate in many sciences; however, participants were performing a novel task with no recourse to numbers or statistical software. Those who did not indicate an effect when $\delta=0$ tended to indicated that they *did not detect* an effect (`r null_nodet_count`; `r round(null_nodet_count/null_n * 100,0)`%), which is the correct conclusion from the SST perspective. The other `r null_same_count` (`r round(null_same_count/null_n * 100,0)`%) indicated that the groups were the same, which under SST is a fallacy.

When there was a true effect ($|\delta|>0$), correct decisions increased as a function of effect size, plateauing at about 95%. Of the `r total_n - null_n` participants for whom $|\delta|>0$, only `r sign_err_count` (`r round(sign_err_count/(total_n - null_n)*100,1)`%) indicated the incorrect team [a sign, or Type S, error; @Gelman:Carlin:2014]. For larger effect sizes, participants never incorrectly indicated that the two groups were the same.

Another way one can evaluate the participants' responses is whether they reflect the information in the display at the time the decision is made, taking into account all points. To roughly quantify the evidence for a difference for each participant, we computed $p$ values from the Wilcoxon rank-sum test between the shuffle reports and the fictitious experimental results as they stood when the participant made their decision. Small Wilcoxon $p$ values suggest a difference between the shuffle reports and the experiments.

```{r response_ev, fig.cap="Statistical evidence underlying participants' decisions. The Wilcoxon rank-sum test $p$ value ($x$ axis) between the participants' null samples and experimental samples is used as a rough index of evidential strength. Kernel density estimates for the evidence are shown for three relevant conclusions. Each point at the bottom represents a single participant. Filled circles show correct decisions; hollow circles, incorrect decisions. The two hollow squares show sign errors."}
MoreyHoekstra2019::fig_response_evidence()
```

Figure \ref{fig:response_ev} shows the distribution of Wilcoxon $p$ values (arranged by the direction of the decision). Kernel density estimates show the distributions of $p$ values when participants indicated that Sparklies were faster, no detection/same, or that Jinglies were faster. With few exceptions, participants' conclusions appear reasonable given the information in the display, though a few participants appear to ignore clear evidence of an effect.

# Self-Reported SST Strategies

After they reported their decision regarding which team they believed was faster, we asked participants three questions about how they performed the task: what was the most salient information for their decision, what was their general strategy, and whether/how they used the shuffle reports.

We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategy), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports.

```{r strategy_table}
MoreyHoekstra2019::christmas_stats_participants %>% 
  mutate(text_strong = text_comparison | text_sampling_var,
               text_only_weak = (text_asymmetry | text_inc_asymmetry ) & !text_strong,
               text_none = !(text_strong | text_only_weak),
               text_missing = text_missing | text_irrelevant) %>%
  summarise(Strong = sum(text_strong),
            `Only weak` = sum(text_only_weak),
            Neither = sum(text_none),
            `No shuffles` = sum(text_no_shuffle),
            `Missing` = sum(text_missing),
            Total = n()) -> for_table

for_table = rbind(for_table, paste0(round(for_table/for_table$Total * 100, 2),"%"))
rownames(for_table) = c("Count","%")

knitr::kable(for_table[,c(1:3,6,4,5)], caption = "\\label{tab:strategies} Frequencies of self-reported strategies.", format = "latex",  align =c("r", "r", "r|", "r||", "r", "r"), booktabs = TRUE)

```

As Table \ref{tab:strategies} shows that a large majority of participants (`r for_table[1,1]`, `r round( 100*as.numeric(for_table[1,1]) / as.numeric(for_table[1,6]), 2)`%) indicated using strong significance testing strategies. We should be cautious, however, because participants were told in the instructions that the shuffle reports could be used for assessing sampling variability. We did this to make clear what the shuffle reports were, but without explaining how to use them. To some extent, then, the text responses may reflect the instructions. However, the data strongly suggest a deeper understanding; first, among the responses were richer, lucid descriptions of SST logic, such as:

> "[t]he random [shuffles] showed quite often such 'strong evidence', even at high sample sizes. That should not happen when the evidence is really strong, so probably the end of the scale was not [so] strong evidence...The random [shuffles] helped me to judge how common misleading evidence in that order of magnitude is, and after 5 samples from the real experiment I concluded that this result is probably not misleading evidence."

The instructions did not tell the participants *how* they should use the shuffles reports, yet many participants gave detailed accounts. Moreover, as we show next, participants' decisions were sensitive to a key manipulation of the shuffles reports in the manner that one would expect if they were using SST logic.

# Sensitivity to SST-Relevant Information

```{r response_ev2, fig.height=4, fig.cap="The effect of the evidence transformation manipulation on responding. Points on top (narrow scale; $q=7$) and bottom (wide scale; $q=3$) represent participants' decisions as a function of the most extreme experiment sampled. ``N'' indicates a ``no detect'' or ``same'' response; ``J/S'' indicates a response in favor of a difference between the groups. Curves show predicted probability by a logistic regression fit with standard errors.", fig.show = "hold", fig.subcap = c("Predicted response probabilities relative to visual extremity. Vertical lines show the critical 0.05 for the corresponding null sampling distribution.","Predicted response probabilities relative to the null sampling distributions (implicit $p$ values)."), out.width = "0.90\\textwidth", fig.ncol = 1}
glm_obj_x <- MoreyHoekstra2019::fig_evidence_scale_logistic_x()

pval = .05

glm_pars_x = broom::tidy(glm_obj_x)
glm_anova_x = broom::tidy(anova(glm_obj_x, test = "Chisq"))
glm_apa_x = papaja::apa_print.glm(glm_obj_x)

pred_p = outer(c(3,7), c(3,7), function(true, z_calc){
  plogis(predict.glm(glm_obj_x, 
              newdata = list(ev = MoreyHoekstra2019:::evz(qnorm(1 - pval / 2), z_calc), 
                 evidence_power = true)
              )
  )
})
dimnames(pred_p) <- list("true"=c("Wide (3)","Narrow (7)"),
                         "z_calc"=c("Wide (3)","Narrow (7)"))

glm_obj_p <- MoreyHoekstra2019::fig_evidence_scale_logistic_p()

glm_pars_p = broom::tidy(glm_obj_p)
glm_anova_p = broom::tidy(anova(glm_obj_p, test = "Chisq"))
glm_apa_p = papaja::apa_print.glm(glm_obj_p)

```

In addition to a random effect size, participants were randomly assigned to one of two transformations of the location/color test statistic from an underlying $z$ statistic. Of particular interest was how the transformation affected responding for the same visual deviation from the center. In one condition ("wide"), the null sampling distribution was visually about twice as wide as in the other ("narrow"). If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence.

Figure \ref{fig:response_ev2} (top) shows responses (no detect/same or Jinglies/Sparklies) as a function of the most extreme experiment sampled ($x$ axis) and the transformation. There was a strong effect of the transformation consistent with use of the null sampling distribution; participants randomly assigned to the "narrow" evidence transformation responded "Jinglies/Sparklies" for much less visually extreme evidence (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_x[3:4,3]),3)`, p<.001$).

A logistic regression relating responses to the visual extremeness of the evidence and the transformation provides predicted probabilities of responding "Jinglies/Sparklies" when the visual evidence corresponded to $p=0.05$ for the null sampling distribution. In both the wide and the narrow conditions, the predicted probability of a "Jinglies/Sparklies" response at the critical value was about 22%, despite that in the wide transformation condition this point was about twice as visually extreme.

Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response_ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).

# Discussion

Although it has previously been suggested that scientists have dramatic misunderstandings of SST logic, scientists and trainees in our experiment demonstrate both understanding and the ability to use it to come to the correct conclusion in a simulated statistical task. Moreover, they report strategies consistent with SST, and the signature of SST reasoning can be seen in their responses. Because we removed numerical effect size and sample size information --- making strategies other than pure significance testing difficult or impossible to apply --- our results are strong evidence that scientists *can* successfully deploy SST logic. It is still an open question what causes SST to be misunderstood so often, but it seems comprehension of its underlying logic can be ruled out for many scientists.

Our findings echo other work showing that human reasoning can, under some conditions, be better than previously understood. [@Cosmides:Tooby:1992;@Gigerenzer:Hoffrage:1995]. Suggestions that SST be discontinued due to scientists' apparent misunderstandings may have been hasty. Of course, there may be other reasons why some favor abandoning SST, but our work shows that given the opportunity, scientists successfully deploy basic SST logic. In spite of scientists' real-life statistical behaviour often resembling a "ritual" [@Gigerenzer:etal:2004], when we eliminate the ritual --- no $p$ value, or any other familiar number, was offered --- they think statistically, very often arriving at the correct conclusion about the sign of the effect. Effective methodological reform in the sciences may be easier than abandoning significance testing, if we can harness scientists' already existing statistical competence.

# Methods

## Participants

Participants were recruited via social media platforms such as Twitter and Facebook. All participants gave informed consent. Data inclusion criteria included sampling at least one shuffle report and experimental result, working in a scientific field, having at least some University education in science, and that it was their first time participating. Details are given in Supplement B.

After applying all inclusion criteria, `r nrow(MoreyHoekstra2019::christmas_stats_participants)` participants remained for analysis.

## Experimental Design and Procedure

Each participant was randomly assigned to one of eight true effect sizes (from $\delta=0$ to $\delta=1$) and one of two evidence powers ("wide" $q=3$ or "narrow" $q=7$; see "Evidence Distributions" below).  The probability of being assigned $\delta=0$ was 25%, while the remaining effect sizes were equally probable at 11%. The probability of assignment to either evidence power was 50%.  Details are given in Table 1.1 in Supplement A.

After offering informed consent, participants read the cover story and instructions. During the instructions, the participant was introduced to the task through sampling random shuffle reports. After a brief recap of the instructions, participants performed the main task --- sampling either random shuffles or experiments --- until they made a decision about which, if either, elf group was faster. They were then asked several open-ended questions about their strategy, some informational questions (results in Supplement B) and debriefed.

```{r}
med_dur = median(as.numeric(MoreyHoekstra2019::christmas_stats_participants$duration) / 60)

```

Qualtrics' duration estimate indicated that the median time spent on the experiment was `r round(med_dur,0)` minutes.

All materials, code, and data are freely available at https://github.com/richarddmorey/Morey_Hoekstra_StatCognition.

## Evidence distributions

The evidence/horizontal ($x$) location test statistic presented to the participant was derived from a transformed $Z$ statistic:
\[
Z \sim \mbox{Normal}(\delta\sqrt{n/2}, 1)
\]
where $\delta$ is a true effect size (randomly assigned to each participant, from 0 to 1) and $n$ is the selected but unknown sample size (from 10 to 200 participants per group). $Z$ then transformed to the (-1,1) space:
\[
x = \mbox{sgn}(Z)\left[1 - \left(1 - F_{\chi_1^2}\left(Z^2\right)\right)^{\frac{1}{q}}\right],\ -1\leq x \leq 1.
\]
where $F_{\chi_1^2}$ is the cumulative distribution function of a $\chi^2_1$ random variable, and $q\in\{3,7\}$ was randomly assigned for each participant. $x=-1$ represented the left edge of the interface, $x=0$ the middle, and $x=1$ the right edge. The setting of $q$ determined how spread out the test statistic was on the display. This arbitrary transformation was done to ensure that the test statistic's distribution was unfamiliar to the participant. See Supplement A for more details, including graphical depictions of the evidence distributions.

## Coding of open-ended strategy questions

We determined the coding scheme and independently categorized the first 20 participant, discussing the source of disagreements. After categorizing the remaining participants, some disagreements were resolved through agreement, and a discussion between the authors was had over what caused the disagreements. The remainder of the disagreements were re-coded separately, and a final round of discussion resolved the remaining disagreements. Participants' codings described in detail in Supplement B.

# References

<div id="refs"></div>

# End notes

* **Data and materials availability**:

* Supplementary Information is linked to the online version of the paper at www.nature.com/nature

* **Author Contributions**:

* Reprints and permissions information is available at www.nature.com/reprints.

* The authors declare no competing interests.

* Correspondence and requests for materials should be addressed to Richard D. Morey (richarddmorey@gmail.com).

* Compiled `r Sys.time()` (`r Sys.timezone()`) under `r version$version.string`.

### 2019-08-02 inst/manuscript/supplementA.Rmd
updated vignette/website compilation

---
title: "Supplement A to 'Researchers understand and use the logic of significance testing in a novel statistical reasoning task'"
subtitle: "Methods details and extra results"
author: "Richard D. Morey"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: kate
    dev: svg
    fig_width: 6
    fig_height: 4
    use_bookdown: true
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Supplement A}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
bibliography: "`r system.file('bib/bibfile.bib', package = 'MoreyHoekstra2019')`"
nocite: | 
  @Venables:Ripley:2002
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

source(
  system.file("font/font_setup.R", package = "MoreyHoekstra2019")
)

```

```{r echo = FALSE, include = FALSE}
library(ggplot2)
library(dplyr)
library(papaja)
library(knitr)
library(kableExtra)
library(tidyr)
library(broom)
library(magrittr)
library(purrr)

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

dat <- MoreyHoekstra2019::christmas_stats_participants
samples <- MoreyHoekstra2019::christmas_stats_samples
```

# Extended description of experiment

The Christmas 2018 statistical cognition experiment ran continuously from 16 December 2018 to 1 January 2019. Participants were recruited via social media (particularly Twitter and Facebook). We used [Qualtrics](https://www.qualtrics.com) to deploy the experiment, which was written in HTML/CSS/Javascript. Participants were asked to perform a series of fictitious experiments with a two-group design and come to a conclusion regarding which of the two groups was "faster". The way the fictitious results were reported to the participants was unusual --- no numbers were given --- to test participants' ability to use significance testing logic.

Of particular interest to us were:

* whether participants sought information relevant to a significance test, and ignored irrelevant information,
* whether participants could come to the right conclusion with high probability,
* whether participants' conclusions were reasonable given the information they were given, and
* whether participants' descriptions of their strategies were consistent with significance testing logic.

## Basic task setup

```{r EStable, results='asis'}
prob = c(.25, 
         .75 * rep(1/7, 7)
)

samples %>% 
  filter( type == "expt" ) %>%
  group_by( toy_name ) %>%
  summarise( `Hidden effect size (δ)` = first(abs(d)) ) %>%
  arrange( `Hidden effect size (δ)` ) %>%
  rename( `Toy name` = "toy_name") %>%
  mutate(Probability = !!prob) %>%
  kable(caption="\"Toy\" name, corresponding true effect sizes used in the study (in standard deviation units) and the probability that a participant would be randomly assigned to that effect size.", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, position = "float_right", font_size = 16)

```

Participants were randomly assigned to one of two evidence powers (wide, $q=3$; or narrow, $q=7$) with equal probability. Participants were also randomly assigned to one of eight "true" effect sizes. Because the behaviour of participants when there is no true effect was particularly of interest, the probability of assignment to no effect ($\delta=0$) was 25%. Across the seven other effect sizes listed in Table \@ref(tab:EStable), the remaining 75% probability was evenly distributed.

The cover story (which can be read here) presented a problem in which it was desired to know which of two groups of elves ("Sparklies" or "Jinglies") was faster. Participants were presented with the results of fictitious experiments as they requested them.

Participants could increase or decrease the sample size for the experiments as well; importantly, they were *not* aware of the actual sample size. Participants could adjust the sample size with a slider that had 20 divisions. The corresponding 20 hidden sample sizes are shown in Table  \@ref(tab:Ntable).

```{r Ntable,results='asis'}
samples %>% 
  group_by(n) %>%
  mutate(`Time (s)` = ceiling(n / 10) ) %>% summarise(`Time (s)` = first(`Time (s)`)) %>% 
  mutate(`n index` = order(n)) %>%
  select(`n index`, n, `Time (s)`) %>%
  kable(caption="Sample size index, underlying hidden per-group sample size, and corresponding time delay in seconds taken to return the experimental result, if the result requested was an experimental sample. Random shuffle reports were returned instantaneously.", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, position = "float_right", font_size = 16)

```

The results were returned to the participants in a visual fashion: instead of being presented with a test statistic, each result was associated with a color intensity (from white to red) and a horizontal location that we will describe as left (-1) to right (1). Results on the far left were red and were associated with maximum evidence for the Sparklies being faster; results in the center were white were not evidence for either group; and results on the right were again red and were evidence for the Jinglies being faster. The intensity of the red color was defined by a linear gradient on the transparency (alpha) from 0 to -1 or 1 as defined using [CSS](https://developer.mozilla.org/en-US/docs/Web/CSS/linear-gradient).[^1] The resulting horizintal axis is shown in the figure below.

[^1]: The CSS code for the transition between red (-1) to white (0) was `linear-gradient(to right, rgba(255,0,0,1), rgba(255,0,0,0))`; from 0 to 1 the transition was the reverse.

```{r cleanInterface, fig.cap = "The interface on which fictitious results were presented to the participants."}
knitr::include_graphics(system.file("pkg_html/img/clean_interface.png", package = "MoreyHoekstra2019"))
```

## Underlying statistical details

Each fictitious result was the result of applying a transformation to randomly sampled $Z$ statistic. The distribution of the $Z$ statistic was a function of the randomly-assigned (but unknown to the participant) effect size $\delta$ and a group sample size $n$ (adjustable by, but unknown to, the participant).

\[
Z \sim \mbox{Normal}(\delta\sqrt{n/2}, 1)
\]

The $x$ location of the result, and hence the color, was then defined by the transformation:
\[
x = \mbox{sgn}(Z)\left[1 - \left(1 - F_{\chi_1^2}\left(Z^2\right)\right)^{\frac{1}{q}}\right]
\]
where evidence power $q\in\{3,7\}$, the location $x \in (-1,1)$, and $F_{\chi_1^2}$ is the cumulative distribution function of the $\chi_1^2$ distribution.

The figure below shows the transformation from $Z$ statistics (top axis) and one-sided $p$ values (bottom axis) to $x$ locations.

```{r transformation, fig.height = 6, fig.cap="Transformation between traditional test statistics ($Z$, $p$) and the $x$ location"}
MoreyHoekstra2019::fig_desc_evidence_transform()
```

```{r ev-distributions,fig.cap="Evidence distributions as functions of underlying effect size, for non-negative effect sizes. Distributions from left to right correspond to increasingly large true effect sizes. Evidence distributions for negative effect sizes were mirror images of these distributions. A: Smallest sample size, wide evidence distribution; B: Largest sample size, wide evidence distribution; C: Smallest sample size, narrow evidence distribution; D: Largest sample size, narrow evidence distribution", fig.height = 10}
par(mfrow = c(4,1))
ymax = MoreyHoekstra2019::fig_desc_evidence(n = 10, q = 3, letter = "A", scale_y_on = 0)
MoreyHoekstra2019::fig_desc_evidence(n = 200, q = 3, letter = "B", scale_y_on = -ymax)
MoreyHoekstra2019::fig_desc_evidence(n = 10, q = 7, letter = "C")
MoreyHoekstra2019::fig_desc_evidence(n = 200, q = 7, letter = "D")

```

```{r pvals-q7, echo = FALSE, out.width='100%', fig.cap="Selected two-sided $p$ values for the null distribution of the evidence for the narrow evidence distribution ($q=7$)"}
fn = MoreyHoekstra2019::fig_evidence_p_vals(q = 7, letter = "q=7")
knitr::include_graphics(fn)
```

```{r pvals-q3, echo = FALSE, out.width='100%', fig.cap="Selected two-sided $p$ values for the null distribution of the evidence for the wide evidence distribution ($q=3$)"}
fn = MoreyHoekstra2019::fig_evidence_p_vals(q = 3, letter = "q=3")
knitr::include_graphics(fn)

```

# Visual examples of task

## Video example

<video style="width: 90%" controls>
  <source src="img/santa_task_example.mp4" type="video/mp4">
</video>

## One participant's final result

```{r null-ex,echo = FALSE, out.width='100%', fig.cap="Caption"}
idx = 7
id = MoreyHoekstra2019::christmas_stats_participants$id[idx]
fn = MoreyHoekstra2019::fig_recreate_display(id, "null", letter = "N")
knitr::include_graphics(fn)
```

```{r expt-ex,echo = FALSE, out.width='100%', fig.cap="Caption"}
fn = MoreyHoekstra2019::fig_recreate_display(id, "expt", letter = "E")
knitr::include_graphics(fn)

```

This participant's (id: *`r id`*) final decision was "`r MoreyHoekstra2019::christmas_stats_participants[idx,"response"]`".

# Difficulties for likelihood or Bayesian accounts

The task is set up specifically to make it difficult to apply other modes of inference to the problem, while *encouraging* significance testing. We would not argue that significance testing is the only mode of inference that participants might use in other contexts, or that it is their preferred mode; rather, the goal was to test whether they had enough of an understanding of significance testing to apply it.

In order to test this, it was necessary to block other kinds of inference. Two aspects of the task make other inference methods difficult: the arbitrary transformation of the test statistic, and the removal of sample size information.

## Likelihood inference

Consider what would be necessary for a likelihood inference from these data. A likelihoodist would require a model with a parameter $\theta$:
\[
l(\theta; \mathbf x) \propto \prod_{i=1}^M f_{n_i}(x_i;\theta)
\]
where $\mathbf x$ is the vector of length $M$ of all evidence/$x$-locations of experimental samples produced by the participant, $x_i$ represents the $i$th element of $\mathbf x$, and $f_{n_i}$ represents the density function of experiments for the hidden sample size $n_i$. The density functions are unknown, as are the sample sizes. There is no obvious measure of effect size on which to build a model.

One might choose an impoverished model that throws out information about $x$ and only uses the ordering of null samples and experimental samples:
\[
\theta = Pr(X>X_0),
\]
where $X$ is a draw from an experiment and $X_0$ is a draw from the shuffle reports. But then $\theta$ would be dependent on $n_i$ in an unknown way except under the null hypothesis where $\theta=0.5$ (because under the null hypothesis experiments and random shuffle reports have the same distribution, by definition). Thus, the null is the only thing the participant knows.

Even a clever participant who knew about the arbitrary transformation from the $z$ statistic would have difficulty. They might use the following strategy:

1. Sample many random shuffle reports, estimate their distribution
2. Find a transformation $z_i = g(x_i)$ to take the shuffle reports to standard normal deviates
3. Use this transformation to transform experimental samples to $z$ statistics
4. Use likelihood inference on the mean of the $z$ statistics from experimental samples

This strategy would require extensive knowledge of statistical theory and, likely, sophisticated programming skills used during the task. It is difficult to imagine anyone applying it (and certainly no participant reported doing such a strategy). But even this strategy is frustrated, because the transformation conflates the effect size and sample size. Only $\delta\sqrt{n/2}$ is estimable, and $n$ is unknown. No inference about $\delta$ is possible, except that it is different from 0.

## Bayesian inference

> "For [statistical] hypotheses, Bayes' theorem tells us this: Unless the observed facts are absolutely impossible on hypothesis $H_0$, it is meaningless to ask how much those facts tend 'in themselves' to confirm or refute $H_0$. Not only the mathematics, but also our innate common sense (if we think about it for a moment) tell us that we have not asked any definite, well-posed question until we specify the possible alternatives to $H_0$." -- @Jaynes:2003

Bayesian inference, being dependent on the likelihood, is saddled with the difficulties outlined above, as well as an additional one: the Bayesian has need of a prior. As Jaynes points out, Bayesian inference requires specified alternative. Typically, this would be a set of alternatives rendered as a prior distribution over some parameter, $p(\theta)$. In this problem, there is no clear parameter and hence it is not clear what the prior would be placed on. Additionally, the sample sizes are unknown, leading to the problem that it is unclear how a Bayesian would update their prior to the posterior; the sample sizes are clearly relevant information, but hidden from the analyst.  The final difficulty is that the likelihood is unknown due to the arbitrary transform applied to the $z$ statistic.

Suppose, however, that the Bayesian analyst knew about the transformation, and applied the strategy described in the previous section to reverse-engineer the transformation back to $z$ scores. The inference about $\delta$ would be dependent on their prior about the sample sizes. As emphasized by Bayesians who appeal to the Jeffreys-Lindley paradox [@Lindley:1957], the statistical support for the null hypothesis $\delta=0$ depend on the sample size such that assuming larger sample sizes guarantees evidence for the null hypothesis. The inference about $\delta$ is completely confounded with $n$: assuming larger $n$ means inferring smaller $\delta$.

It is possible that a Bayesian making just the right assumptions, and working very hard to reverse-engineer the transformation, could perform the task well. However, their inference would be dependent on these strong assumptions.

No participant reported using such a strategy (or anything like it).

# Self-reported understanding of shuffle reports

Question: "Do you understand why the random shuffle reports could be useful?"

```{r shuffle-understanding, fig.caption="Caption"}
MoreyHoekstra2019::fig_shuffle_understanding()
```

# Participants' confidence in their responses

Question: How confident are you in your assessment above?

* Not confident at all 
* Somewhat doubtful 
* Somewhat confident 
* Very confident

```{r confidence-null, fig.cap="Caption"}
MoreyHoekstra2019::fig_confidence(which_resp = "null")
```

```{r confidence-alt, fig.cap="Caption"}
MoreyHoekstra2019::fig_confidence(which_resp = "alt")
```

# Participant sampling behavior

```{r sampling, fig.cap="Sampling behavior by effect size. Each point represents a participant. A: Number of samples from the null distribution as a function of true effect size. B: Number of samples of fictional 'experiments' as a function of true effect size. Note that the $y$ axis is logarithmically scaled. Lines are robust regression fits (Venables \\& Ripley, 2002)."}
MoreyHoekstra2019::fig_sampling_behavior()
```

```{r}
kt_null <- kruskal.test(n_null ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
kt_expt <- kruskal.test(n_expt ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
```

Participants sought out information that would be necessary for significance tests. They made heavy use of shuffle reports (Figure A above). Across all true effect sizes, participants sampled a median of 152 shuffle reports (range: 1-2034).

Participants also made use of "replications" of the fictional experiments. (Figure B above) shows the distribution of the number of experiments sampled as a function of the true effect size. Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`). When the effect size is small and difficult to detect, participants do more experimentation before deciding.

# Open-ended strategy questions

## Coding of strategy-related questions

After they reported their decision, participants were asked three open-ended questions related to their overall experimental strategy:

1. "What facts or observations were most salient in coming to the conclusion that you did (whatever that was)?"
2. "Please describe your experimental strategy, if any, in your own words."
3. "Did you make use of the 'random shuffle reports'? If so, how?"

Authors RM and RH coded the responses as one or more of the following. Responses were judged holistically across the three questions; occasionally a response was unclear to one question, but their response to another question clarified their meaning.

Coding was performed without reference to the participant's assigned condition and without reference to their decision (to the greatest extent possible; sometimes the text of their answer suggested the condition or decision).

1. *Comparison to the random shuffle reports*: Participant mentioned comparing the experimental results to the random shuffle reports
    * Example (to Q1): "A comparison of the experimental results when compared to the 'random shuffle' arrangements."
    * Example (to Q2): "...[experiments were] much more extreme than would be predicted by chance given the sampling distribution from the random experiments"
2. *Random shuffle reports to assess sampling variability*: Participant mentioned using the random shuffle reports as a guide to the expected variability of the evidence/sampling distribution/noise/random error etc.
    * Example (to Q3): "Yup - smashed it to bits to see the variation in difference 'scores'."
    * Example (to Q2): "Get an idea of the variability (and expected difference, which seemed to be zero difference) using the 'random shuffle' results, and then see how the experimental results varied."
3. *Asymmetry of experiments*: Participant noted that the observations tend to be on one side or another, or favoring one team or another; that there was a *lack* of asymmetry; or that symmetry was in some way considered by the participant.
    * Example (to Q2): "Conducted experiment with max participants 4 times; all observations well to the Jinglies' side..."
    * Example (to Q2): "take samples see whether consistently left or right of middle..."
    * Example (to Q1): "all [experiments] came out 1 way"
    * Example (to Q1): "Approx same number of wins for each team ; no clear visual assymetry"
4. *Increasing asymmetry as N increased*: Participant mentioned noting a pattern where the extremeness of results appeared to increase as the sample size increased, a signature of a true effect, or noting that such information would have been of interest.
    * Example (to Q1): "...larger sample sizes tended to show clearer evidence of difference"
5. *Response not relevant*: Responses in this category did not answer the questions posed.
6. *Response missing*: No response was given to any of the three questions.
7. *Explicitly denied using shuffle reports*: In describing their strategy, the participant said that they *did not use* the shuffle reports or did not find them useful in their decision.

Categories 1, 2, 3, and 4 are consistent with significance testing logic. Categories 1 and 2 represent high-quality SST strategies, whereas Categories 3 and 4 may represent less efficient (e.g., ignoring variance of null distribution), unclear, or invalid SST strategies.

We must note that although many or most of the responses clearly fit into one category or another, sometimes categorization was not clear. We attempted to ensure the most reasonable coding for each participant through multiple rounds of coding and discussions. Although individual participant's codings might be up for debate, in the aggregate they provide a guide to the strategies deployed by participants and a rough estimate of their frequencies.

## Frequencies of strategies

The figure below shows the frequencies of various self-reported strategies as coded by the authors.

```{r strategies, fig.cap="Coded frequencies of different strategies in the open-ended responses. A \"Strong\" indicates use of comparison to the null distribution or  its use to assess sampling variability/distribution. A \"Weak only\" response indicates some indication of symmetry or asymmetry (or its use as a strategy)."}
MoreyHoekstra2019::fig_strategies()
```

```{r results="asis"}
MoreyHoekstra2019::text_coding %>%
  select(-id) -> lst_dat

marg_vec <- sapply(lst_dat,sum)/nrow(lst_dat)

prob_mat <- diag(length(lst_dat))*NA
rownames(prob_mat) <- 
  colnames(prob_mat) <- 
  c("Comparison","Asymmetry","Samp. variance", "Inc. asymmetry","No shuffles","Irrelevant","Missing")

for(i in 1:length(lst_dat))
  for(j in 1:length(lst_dat))
    prob_mat[i,j] = switch((i == j) + 1,
                           mean(lst_dat[[i]] & lst_dat[[j]]) / marg_vec[j],
                           1
      )

prob_mat = prob_mat[c(1,3,2,4,5),c(1,3,2,4,5)]
prob_mat %>% 
  as_tibble() %>%
  mutate(` ` = rownames(prob_mat)) %>%
  select(` `, everything()) %>%
  mutate_if(is.numeric, function(x) {
    x = c(0,round(x, 2),1)
    z = cell_spec(x, bold = TRUE, 
              color = spec_color(x, end = 0.9),
              font_size = spec_font_size(x))
    z[c(-1, -length(x))]
  }) %>%
  kable(caption=paste0("The conditional probabilities of the various responses. Each entry shows the probability of the response type on the *row* **given** the response type on *column*. These probabilities exclude the ",sum(marg_vec[6:7]*nrow(lst_dat))," missing/irrelevant responses."), escape = FALSE, align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE, position = "float_right", font_size = 16) %>%
  column_spec(1, bold = TRUE, border_right = TRUE) %>%
  add_header_above(c(" " = 1, "(given)" = 5))

```

# Exploratory sanity checks

This section reports a few analyses based on the coded open-text responses. Analyses in this section should be treated with skepticism due to theie exploratory nature and the fact that they are conditioning on an observed variable that is likely correlated in interesting ways with the decision. Also, some people did not respond to these questions, so their strategy is unknown; yet they will be categorized with those who responded and offered an invalid strategy.

## Error rates

We can compare error rates of those who reported using a strong significance-testing strategy versus ones who did not.

```{r fig.cap="Proportion who have a 'difference' response as a function of effect size, for participants who were coded as reported 0, 1, or 2 strong significance testing strategies. Ribbons are standard errors."}

dat <- MoreyHoekstra2019::christmas_stats_participants %>%
  mutate(n_strong = text_sampling_var + text_comparison) %>%
  mutate(
    effect_size = abs(effect_size),
    ln_null = log10(n_null),
    ln_expt = log10(n_expt)
  )

dat %>%
    mutate(effect_size = abs(effect_size),
           `# strong` = factor(n_strong)) %>%
    group_by(effect_size, true_winner, response, `# strong`) %>%
    tally() %>%
    complete(effect_size, true_winner, response, `# strong`, fill = list(n = 0)) %>%
    ungroup() %>%
    filter(response != "bored") %>%
    mutate(
      dec_type = case_when(
        true_winner != "null" &
          true_winner != response &
          response %in% c("jinglies", "sparklies") ~ "sign_error",
        response %in% c("jinglies", "sparklies") ~ "report_difference",
        response == "same" ~ "same",
        response == "no_detect" ~ "no_detect"
      )
    ) %>%
    group_by(dec_type, effect_size, `# strong`) %>%
    tally( wt = n ) %>%
    complete(dec_type, effect_size, `# strong`, fill = list(n = 0)) %>%
    group_by(effect_size, `# strong`) %>%
    mutate(total = sum(n)) %>%
  filter(dec_type == "report_difference") %>%
  mutate(
      p = n / total,
      stderr = sqrt(p * (1 - p) / total),
      lo = qlogis(p),
      lo.stderr = stderr / dlogis(qlogis(p))
    ) %>%
  select(-dec_type) %>% 
  ggplot(aes(x = effect_size, y = p, group = `# strong`)) +
      geom_rect(
      xmin = 0,
      xmax = 1,
      ymin = 0,
      ymax = 1,
      color = NA,
      fill = "#EEEEEE",
      alpha = .1
    ) +
    geom_rect(
      xmin = 0,
      xmax = 1,
      ymin = .8,
      ymax = 1,
      color = NA,
      fill = "#DDDDDD",
      alpha = .1
    ) +
    geom_rect(
      xmin = 0,
      xmax = 1,
      ymin = 0,
      ymax = .2,
      color = NA,
      fill = "#DDDDDD",
      alpha = .1
    ) +
    geom_rect(
      xmin = 0,
      xmax = 1,
      ymin = .95,
      ymax = 1,
      color = NA,
      fill = "#CCCCCC",
      alpha = .1
    ) +
    geom_rect(
      xmin = 0,
      xmax = 1,
      ymin = 0,
      ymax = .05,
      color = NA,
      fill = "#CCCCCC",
      alpha = .1
    ) +
    geom_segment(aes(
      x = 0,
      y = 0,
      xend = 0,
      yend = 1
    )) +
    geom_hline(yintercept = c(0, 1)) +
    geom_line(aes(linetype = `# strong`, col = `# strong`), size = .7) +
    geom_point(aes(col = `# strong`)) +
    geom_ribbon(aes(
      ymin = p - stderr,
      ymax = p + stderr,
      fill = `# strong`
    ),
    alpha = 0.2) +
      xlab("Effect size") + ylab("Proportion of responses") +
    coord_cartesian(
      ylim = c(0, 1.2),
      xlim = c(0, 1),
      expand = FALSE,
      clip = "off"
    ) +
    scale_y_continuous(breaks = seq(0, 1, .1)) +
    scale_x_continuous(breaks = seq(0, 1, .2)) +
    theme(
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.background = element_blank(),
      text = element_text()
    )

```

## Sampling behaviour

Did participants who reported more strong significance testing strategies use more null samples, on average? It appears so (left figure below; note the log scale).

Did participants who reported more strong significance testing strategies use more (or fewer) experimental samples, on average? There's not much evidence either way, but importantly (see plot) those who reported *fewer* strong significance testing strategies didn't also sample fewer experiments on average. There is nothing here to suggest they were less engaged on average.

```{r fig.cap="Boxes, in order from left to right (red, green, blue) are 0, 1, and 2 strong significance testing strategies, respectively. Lines are robust regression fits. Stars show participants who indicated they did not use the random shuffle reports. Null samples (left) order from bottom to top (0, 1, 2); experimental samples (right) order is (3, 0, 1) from bottom to top."}
cnull <- dat %>%
  split(.$n_strong) %>%
  map( ~ MASS::rlm(ln_null ~ effect_size, data = .)) %>%
  map_df(coef)

cexpt <- dat %>%
  split(.$n_strong) %>%
  map( ~ MASS::rlm(ln_expt ~ effect_size, data = .)) %>%
  map_df(coef)

dat %>% filter(text_no_shuffle) -> only_no_shuffle

# grouped boxplot
p1 = ggplot(dat,
            aes(
              x = effect_size,
              y = n_null,
              group = interaction(n_strong, effect_size),
              fill = factor(n_strong)
              )) +
  geom_boxplot() +
  scale_y_continuous(trans = 'log10') +
  theme(legend.position = "none") +
  ylab("Number of samples") + ggtitle("Null") +
  xlab("Effect size") +
  geom_abline(
    intercept = cnull$`0`[1],
    slope = cnull$`0`[2],
    color = scales::hue_pal()(2)[1]
    ) +
  geom_abline(
    intercept = cnull$`1`[1],
    slope = cnull$`1`[2],
    color = scales::hue_pal()(2)[2]
    ) +
  geom_abline(
    intercept = cnull$`2`[1],
    slope = cnull$`2`[2],
    color = scales::hue_pal()(3)[3]
    ) +
  geom_point(data = only_no_shuffle,
             mapping = aes(x = effect_size, 
                           y = n_null), 
             shape = 8,
             size = 3)

        # grouped boxplot
p2 = ggplot(dat,
            aes(
              x = effect_size,
              y = n_expt,
              group = interaction(n_strong, effect_size),
              fill = factor(n_strong)
              )) +
  geom_boxplot() +
  scale_y_continuous(trans = 'log10') +
  theme(legend.position = "none") +
  ylab("Number of samples") + ggtitle("Experiments") +
  xlab("Effect size") +
  geom_abline(
    intercept = cexpt$`0`[1],
    slope = cexpt$`0`[2],
    color = scales::hue_pal()(2)[1]
    ) +
  geom_abline(
    intercept = cexpt$`1`[1],
    slope = cexpt$`1`[2],
    color = scales::hue_pal()(2)[2]
    ) +
  geom_abline(
    intercept = cexpt$`2`[1],
    slope = cexpt$`2`[2],
    color = scales::hue_pal()(3)[3]
    ) +
    geom_point(data = only_no_shuffle,
             mapping = aes(x = effect_size, 
                           y = n_expt), 
             shape = 8,
             size = 3)

gridExtra::grid.arrange(p1, p2, ncol = 2)

```

-------

*Compiled `r Sys.time()` (`r Sys.timezone()`) under `r version$version.string`.*

-------

# References

### 2019-08-02 inst/manuscript/supplementB.Rmd
updated vignette/website compilation

---
title: "Supplement B to 'Researchers understand and use the logic of significance testing in a novel statistical reasoning task'"
subtitle: "Inclusion criteria and demographic/opinion questions"
author: "Richard D. Morey"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: kate
    dev: svg
    fig_width: 6
    fig_height: 4
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Supplement B}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

source(
  system.file("font/font_setup.R", package = "MoreyHoekstra2019")
)

```

```{r echo = FALSE, include = FALSE}
library(ggplot2)
library(dplyr)
library(papaja)

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

dat <- MoreyHoekstra2019::christmas_stats_participants
```

# Inclusion criteria

The following criteria were all required to be met by participants.

## Requirements of participation

* **Informed consent**. Participants were given the opportunity to stop the experiment if they did not consent to participation. If they declined consent, the experiment ended immediately. 
* **A desktop browser**. Qualtrics allows the exclusion of users of mobile devices from participation. We disallowed mobile users because the small size of many mobile screens would make the task difficult. Mobile users were shown a message inviting them to participate later from a desktop device.

## Requirements of data inclusion

* **First time doing experiment**: Must have answered "No, this is my FIRST TIME doing this survey" to the question "Have you completed this survey previously?" 
* **Following directions/engagement**: Must have performed at least 1 null sample and 1 experimental sample. Participants eliminated by this criterion started the instructions, but did not perform the experimental task. 
* **Field is arguably scientific**: Must have offered an answer to the question "Is your work in a field that would typically be considered scientific?", and that answer must *not* have been "No". 
* **Education at University level**: Must have offered an answer to the question "What is the highest level of formal education you have achieved in your scientific field?", and that answer must *not* have been "I have no formal scientific education at the University level".

# Deleted/edited records

After data collection was complete, we were made aware of the possibility that our data set included minors. Because our experiment was targeted at scientists, we did not include indication of parental consent for minors in our consent form. Although there is nothing ethically problematic about our experiment, out of an abundance of caution and in consultation with the Cardiff University School of Psychology Board of Ethics we deleted 32 records that may have been minors. We used the following criteria to select these records:

* The participant indicated no university-level education, and
* In their reply to the question about their scientific field, their response included any of the following terms: "ap" (as in *advanced placement*), "student", "school", "high", or "class".

These records were deleted from Qualtrics. Because none of the deleted records indicated University-level education, this did not affect the data that met the criteria for analysis.

The record deletion was done on 2 May, 2019.

An additional 15 records were deleted on 31 May, 2019 when it was found that some students had indicated their student status in other fields.

One record was further edited to remove personal information on 31 May, 2019. This record (id: *R_2a8OIVnAjcZXHyU*) now includes a note indicating the redaction.

# Demographic and informational questions

This section outlines the responses to the informational questions in the Christmas 2018 statistical cognition experiment. Only the "valid" participants are included below.

## Field is scientific

Question: "Is your work in a field that would typically be considered scientific?"

* Yes 
* No 
* Depends/Maybe/Other [specify]

```{r}
MoreyHoekstra2019::fig_field_scientific()
```

## Education and training

Question: "What is the highest level of formal education you have achieved in your scientific field?"

* I have no formal scientific education at the University level 
* I have some education at the Bachelor level (or equivalent) 
* I have completed a Bachelor's degree (or equivalent) 
* I have some education above the Bachelor level (e.g., Master's, PhD, or equivalent) 
* I have completed a PhD (or equivalent)

```{r}
MoreyHoekstra2019::fig_edu_training_edu()
```

Question: "How many years of formal statistical training do you have?"

[enter a number]

```{r}
MoreyHoekstra2019::fig_edu_training_stats()

summary(dat[,"formal_training"])
```

## Scientific field

Question: "In what applied field(s) do you use statistics? (check all that apply)"

* Biological sciences 
* Medical sciences 
* Physical sciences 
* Social/Behavioral sciences 
* Computer science / technology 
* Other [specify]

```{r}
MoreyHoekstra2019::fig_field()
```

## Role of statistics

Question: "How do statistics play a role in your work? (check all that apply)"

* Statistics do not play a role in my work 
* I use statistics in practice for the analysis of data 
* I comment on statistical practice, but do not develop them myself 
* I develop statistical methods 
* I comment on the philosophy of statistics 
* Other [specify]

```{r}
MoreyHoekstra2019::fig_how_use_stat()
```

## Preferred statistical method

Question: "If you use statistics regularly in your work, what sort of inferential procedures would you typically prefer?"

* Classical, frequentist, or error statistical 
* Bayesian (of any sort) 
* None; I prefer only descriptive 
* I do not use statistics in my work 
* Other [specify]

```{r}
MoreyHoekstra2019::fig_preferred_method()
```

## Significance testing

Question: "What is your opinion about statistical significance testing? (check all that apply)"

* I do not have a strong opinion for or against significance testing 
* I think significance testing is necessary for science 
* I think significance testing is fine, but prefer other approaches to statistical inference 
* I think significance testing is fine, but people misunderstand/misuse it 
* I think significance testing should be discontinued or performed very rarely 
* I think the logic of significance testing is fatally flawed 
* I do not understand the question 
* Other [specify]

```{r}
MoreyHoekstra2019::fig_sig_testing_opinion()
```

## Data downloads

During debriefing, participants were given the option of downloading their data for their own analysis. The table below shows how many took advantage of the option for both shuffle reports and experimental samples.

<style>
.table_label td {
  padding: 5px;
}
</style>

```{r}
X = with(MoreyHoekstra2019::christmas_stats_participants, 
     table(downloaded_null, downloaded_expt) )
dimnames(X) = list(`Shuffle reports` = c("Did not download", "Downloaded"),
                   `Experiments` = c("Did not download", "Downloaded"))

# https://stackoverflow.com/questions/26429740/preserve-row-column-labels-from-table-using-kable-and-knitr
table_label <- function(tbl) {

  # table dimensions
  rows <- dim(tbl)[1]
  cols <- dim(tbl)[2]

  # get started
  html_out <- '<table class = "table_label">\n'

  # first row: label only
  blank_cell <- '<td>&nbsp;</td>'
  html_out <- 
    paste0(html_out,
           '\t<tr>',
           blank_cell, 
           '<td>', names(dimnames(tbl))[2], '</td>', # column label
           rep(blank_cell, cols-2),
           '</tr>\n')

  # second row:
  html_out <- 
    paste0(html_out,
           '\t<tr>',
           # label...
           '<td>', names(dimnames(tbl))[1], '</td>',
           # ...and headers
           paste0('<td>', dimnames(tbl)[[2]], '</td>', collapse=''),
           '</tr>\n')

  # subsequent rows
  for (i in 1:rows) {
    html_out <- 
      paste0(html_out,
             '\t<tr>',
             # header... 
             '<td>', dimnames(tbl)[[1]][i], '</td>',                        
             # ...and values
             paste0('<td>', tbl[i,], '</td>', collapse=''),
             '</tr>\n')
  }

  # last row
  html_out <- paste0(html_out, '</table>')
  return(html_out)
}

```

```{r, results='asis'}
cat(table_label(X))
```

-------

*Compiled `r Sys.time()` (`r Sys.timezone()`) under `r version$version.string`.*

### 2019-08-02 inst/build_site.R

@@ -1,0 +1,31 @@
+library(here)
+
+pkgdown::build_site( here() )
+
+rmarkdown::render( input = here("README.Rmd"),
+                   output_file = here("README.md"),
+                   output_format = "md_document")
+
+tmpdir = tempdir()
+
+which_files = c("MoreyHoekstra2019/inst/doc/manuscript.pdf",
+                "MoreyHoekstra2019/inst/doc/supplementA.html",
+                "MoreyHoekstra2019/inst/doc/supplementB.html")
+which_dirs = c("MoreyHoekstra2019/inst/pkg_html")
+
+pkg_file = devtools::build( pkg = here(),
+                            path = tmpdir,
+                            quiet = TRUE )
+
+untar( pkg_file,
+       files = c( which_files, which_dirs ),
+       exdir = tmpdir )
+
+file_paths = paste( tmpdir, which_files, sep = "/" )
+dir_paths = paste( tmpdir, which_dirs, sep = "/" )
+
+fs::dir_create( here("docs/extras") )
+file.copy( c(file_paths, dir(dir_paths, full.names = TRUE) ),
+           here("docs/extras"),
+           recursive = TRUE )
+

### 2019-08-02 README.Rmd
updated vignette/website compilation
fix link in readme

@@
 ``` r
-devtools::install_github("richarddmorey/Morey_Hoekstra_StatCognition")
+devtools::install_github("richarddmorey/Morey_Hoekstra_StatCognition", build_vignettes = TRUE, depedencies = TRUE)
 ```
@@
 
-You can find an [online version of the task demonstration](articles/task_demo.html).
+You can find an [online version of the task demonstration](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html).
 

### 2019-08-06 inst/manuscript/manuscript.Rmd
Add project file, fix typo

@@
 This is larger than the typically-accepted 5% false positive rate in many sciences; however, participants were performing a novel task with no recourse to numbers or statistical software.
-Those who did not indicate an effect when $\delta=0$ tended to indicated that they *did not detect* an effect (`r null_nodet_count`; `r round(null_nodet_count/null_n * 100,0)`%), which is the correct conclusion from the SST perspective.
+Those who did not indicate an effect when $\delta=0$ tended to indicate that they *did not detect* an effect (`r null_nodet_count`; `r round(null_nodet_count/null_n * 100,0)`%), which is the correct conclusion from the SST perspective.
 The other `r null_same_count` (`r round(null_same_count/null_n * 100,0)`%) indicated that the groups were the same, which under SST is a fallacy.

### 2020-02-05 inst/manuscript/manuscript.Rmd
small changes ahead of resubmission

@@
 However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication).
-It is unclear how well vignette studies (including ones by the present authors; @Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results.
+It is unclear how well vignette studies (including ones by the present authors [@Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results.
 Conceptual understanding and fluency with common representations are both important, but are distinct.
@@
 
-using a simulated experimental task, we tested researchers' and trainees' understanding and application of SST.
+Using a simulated experimental task, we tested researchers' and trainees' understanding and application of SST.
 Our key innovation allowing us to focus on SST was to prevent---or make difficult---the use of alternative strategies.
@@
 Participants were scientists or trainees recruited via social media.
-Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of Christmas elves --- "Jinglies" or "Sparklies" --- could make more of a particular toy, where the true effect size was a randomly-assigned group difference between 0 and 1 standard deviations ($\delta$).
-The numerical information for an experiment, including sample sizes and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}).
+Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy, where the true effect size was a randomly-assigned group difference between 0 and 1 standard deviations ($\delta$).
+Because the study was run around the Christmas holiday season, it was hoped that the theme would make the task more enjoyable.
+The numerical information for an experiment, including sample size and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}).
 Participants could change the sample size per group for each experiment (increasing the time required to return a result), but did not know its numerical value.
-The interpretation of these colors and locations was unknown to the participants, aside from the monotone relationship with the statistical evidence.
+Importantly, the meaning of the colors and locations was unknown to the participants, aside from the monotone relationship with the sample size and statistical evidence.
 
@@
 These results took no time to return.
-Participants were not instructed how to use these samples, only that they might use them.
+Participants were not told how to use these samples, only that they might use them.
 
@@
 
-We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategy), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports.
+We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategies), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports.
 
@@
 We determined the coding scheme and independently categorized the first 20 participant, discussing the source of disagreements.
-After categorizing the remaining participants, some disagreements were resolved through agreement, and a discussion between the authors was had over what caused the disagreements.
+After categorizing the remaining participants, some disagreements were resolved through mutual agreement, and a discussion between the authors was had over what caused the disagreements.
 The remainder of the disagreements were re-coded separately, and a final round of discussion resolved the remaining disagreements.
@@
 
-* **Data and materials availability**:
+* **Data and materials availability**: All materials, code, and data are freely available at https://github.com/richarddmorey/Morey_Hoekstra_StatCognition.
 
-* Supplementary Information is linked to the online version of the paper at www.nature.com/nature
-
-* **Author Contributions**:
-
-* Reprints and permissions information is available at www.nature.com/reprints.
+* **Author Contributions**: RDM conceptualized and designed the study in consultation with RH. RDM analysed the data and created the materials and figures. The manuscript was written by RDM and RH.
 

### 2020-02-05 inst/manuscript/supplementA.Rmd
small changes ahead of resubmission

@@
 
-```{r null-ex,echo = FALSE, out.width='100%', fig.cap="Caption"}
+```{r null-ex,echo = FALSE, out.width='100%', fig.cap="Selected participant's random shuffle report samples."}
 idx = 7
@@
 
-```{r expt-ex,echo = FALSE, out.width='100%', fig.cap="Caption"}
+```{r expt-ex,echo = FALSE, out.width='100%', fig.cap="Selected participant's experimental samples."}
 fn = MoreyHoekstra2019::fig_recreate_display(id, "expt", letter = "E")
@@
 The inference about $\delta$ would be dependent on their prior about the sample sizes.
-As emphasized by Bayesians who appeal to the Jeffreys-Lindley paradox [@Lindley:1957], the statistical support for the null hypothesis $\delta=0$ depend on the sample size such that assuming larger sample sizes guarantees evidence for the null hypothesis.
+As emphasized by Bayesians who appeal to the Jeffreys-Lindley paradox [@Lindley:1957], the statistical support for the null hypothesis $\delta=0$ depends on the sample size such that assuming larger sample sizes guarantees evidence for the null hypothesis.
 The inference about $\delta$ is completely confounded with $n$: assuming larger $n$ means inferring smaller $\delta$.
@@
 
-```{r shuffle-understanding, fig.caption="Caption"}
+```{r shuffle-understanding, fig.caption="Frequencies of responses to the initial question, \"Do you understand why the random shuffle reports could be useful?\""}
 MoreyHoekstra2019::fig_shuffle_understanding()
@@
 
-```{r confidence-null, fig.cap="Caption"}
+```{r confidence-null, fig.cap="Reported confidence ratings (vertical axis) in their response for participants who judged that they either could not detect a difference between the groups, or that the groups were the same, by true effect size (horizontal axis).
+Columns with very vew participants in them (when the effect size was large, but they did not respond that there was a difference) are faded to indicate lack of trustworthiness of the frequencies."}
 MoreyHoekstra2019::fig_confidence(which_resp = "null")
@@
 
-```{r confidence-alt, fig.cap="Caption"}
+```{r confidence-alt, fig.cap="Reported confidence ratings (vertical axis) in their response for participants who judged that they either could not detect a difference between the groups, or that the groups were the same, by true effect size (horizontal axis)."}
 MoreyHoekstra2019::fig_confidence(which_resp = "alt")
@@
 
-```{r fig.cap="Proportion who have a 'difference' response as a function of effect size, for participants who were coded as reported 0, 1, or 2 strong significance testing strategies.
+```{r fig.cap="Proportion who have a \"difference\" response as a function of effect size, for participants who were coded as reported 0, 1, or 2 strong significance testing strategies.
 Ribbons are standard errors."}

### 2020-02-05 inst/build_site.R

@@ -10,7 +10,8 @@ tmpdir = tempdir()
 
 which_files = c("MoreyHoekstra2019/inst/doc/manuscript.pdf",
                 "MoreyHoekstra2019/inst/doc/supplementA.html",
-                "MoreyHoekstra2019/inst/doc/supplementB.html")
+                "MoreyHoekstra2019/inst/doc/supplementB.html",
+                "MoreyHoekstra2019/inst/reprohack/reprohack2019_feedback_letter.pdf")
 which_dirs = c("MoreyHoekstra2019/inst/pkg_html")
 
 pkg_file = devtools::build( pkg = here(),

### 2020-02-05 inst/manuscript/manuscript.Rmd
manuscript, added sentence for clarity

@@
 In one condition ("wide"), the null sampling distribution was visually about twice as wide as in the other ("narrow").
-If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence.
+If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is visually wider.
 

### 2020-02-05 inst/manuscript/manuscript.Rmd
added sentences to manuscript for clarity

@@
 Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response_ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).
+It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with a small number of participants incorrectly using the nondiagnostic visual extremeness to perform the task.
+If more people had been fooled by the irrelevant visual manipulation of the sampling distribution, we would expect this effect to be substantially larger.
 

### 2020-02-07 inst/manuscript/manuscript.Rmd
made Lato the default font for ggplot graphs

@@
 
+MoreyHoekstra2019::pkg_options(ggplot_family="Lato Light")
+
 knitr::opts_chunk$set(echo = FALSE, 
@@
                       message = FALSE,
-                      out.width = "0.95\\textwidth")
+                      out.width = "0.95\\textwidth",
+                      dev = "cairo_pdf")
 
@@
 The shaded region around the solid line is one standard error wide.
-The difference between the black and the gray dashed lines shows the proportion of participants who decided that the two groups were the same (as opposed to failing to detect a difference)."}
+The difference between the black and the gray dashed lines shows the proportion of participants who decided that the two groups were the same (as opposed to failing to detect a difference).", fig.showtext=FALSE}
 MoreyHoekstra2019::fig_error_rates()

### 2020-02-07 inst/manuscript/supplementA.Rmd
made Lato the default font for ggplot graphs

@@
 knitr::include_graphics(fn)
-
 ```
@@
 
-```{r shuffle-understanding, fig.caption="Frequencies of responses to the initial question, \"Do you understand why the random shuffle reports could be useful?\""}
+```{r shuffle-understanding, fig.caption="Frequencies of responses to the initial question, \"Do you understand why the random shuffle reports could be useful?\"", fig.showtext=FALSE}
 MoreyHoekstra2019::fig_shuffle_understanding()
@@
 Note that the $y$ axis is logarithmically scaled.
-Lines are robust regression fits (Venables \\& Ripley, 2002)."}
+Lines are robust regression fits (Venables \\& Ripley, 2002).", fig.showtext=FALSE}
 MoreyHoekstra2019::fig_sampling_behavior()
@@
 A \"Strong\" indicates use of comparison to the null distribution or  its use to assess sampling variability/distribution.
-A \"Weak only\" response indicates some indication of symmetry or asymmetry (or its use as a strategy)."}
+A \"Weak only\" response indicates some indication of symmetry or asymmetry (or its use as a strategy).", fig.showtext=FALSE}
 MoreyHoekstra2019::fig_strategies()
@@
 ```{r fig.cap="Proportion who have a \"difference\" response as a function of effect size, for participants who were coded as reported 0, 1, or 2 strong significance testing strategies.
-Ribbons are standard errors."}
+Ribbons are standard errors.", fig.showtext=FALSE}
 
@@
       panel.background = element_blank(),
-      text = element_text()
+      text = element_text(
+        family = MoreyHoekstra2019::pkg_options("ggplot_family")
+        )
     )
@@
 Stars show participants who indicated they did not use the random shuffle reports.
-Null samples (left) order from bottom to top (0, 1, 2); experimental samples (right) order is (3, 0, 1) from bottom to top."}
+Null samples (left) order from bottom to top (0, 1, 2); experimental samples (right) order is (3, 0, 1) from bottom to top.", fig.showtext=FALSE}
 cnull <- dat %>%
@@
   scale_y_continuous(trans = 'log10') +
-  theme(legend.position = "none") +
+  theme(legend.position = "none",
+        text = element_text(
+          family = MoreyHoekstra2019::pkg_options("ggplot_family"))
+        ) +
   ylab("Number of samples") + ggtitle("Null") +
@@
   scale_y_continuous(trans = 'log10') +
-  theme(legend.position = "none") +
+  theme(legend.position = "none",
+        text = element_text(
+          family = MoreyHoekstra2019::pkg_options("ggplot_family"))
+        ) +
   ylab("Number of samples") + ggtitle("Experiments") +

### 2020-02-07 R/fig_edu_training.R

@@ -10,7 +10,9 @@
 #' @export
 #'
 #' @examples
-#' fig_edu_training_edu()
+#' suppressWarnings(
+#'   fig_edu_training_edu()
+#' )
 fig_edu_training_edu <- function(filter_func = function(data, ...) return(data),
                                  print_plot = TRUE) {
 
@@ -22,7 +24,8 @@ fig_edu_training_edu <- function(filter_func = function(data, ...) return(data),
     geom_bar(mapping = aes(x = education, y = ..prop.., group = 1)) +
     geom_text(stat = "count",
               aes(y = ..prop.., label = ..count.., group = 1),
-              hjust = -.2) +
+              hjust = -.2,
+              family = pkg_options("ggplot_family")) +
     ylim(0, 1) +
     #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
     theme_bw() +
@@ -30,7 +33,7 @@ fig_edu_training_edu <- function(filter_func = function(data, ...) return(data),
       panel.grid.major.y = element_blank(),
       panel.grid.minor.y = element_blank(),
       #panel.background = element_blank(),
-      text = element_text()
+      text = element_text(family = pkg_options("ggplot_family"))
     ) +
     geom_hline(yintercept = c(0, 1)) +
     xlab("Highest education") +
@@ -55,7 +58,9 @@ fig_edu_training_edu <- function(filter_func = function(data, ...) return(data),
 #' @export
 #'
 #' @examples
-#' fig_edu_training_stats()
+#' suppressWarnings(
+#'   fig_edu_training_stats()
+#' )
 fig_edu_training_stats <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -77,7 +82,8 @@ fig_edu_training_stats <- function(
       label = missing,
       hjust = 1.1,
       vjust = 1.5,
-      size = 5
+      size = 5,
+      family = pkg_options("ggplot_family")
     ) +
     xlab("Self-reported years of statistical training") +
     ylab("Frequency") +
@@ -86,7 +92,7 @@ fig_edu_training_stats <- function(
       panel.grid.major.y = element_blank(),
       panel.grid.minor.y = element_blank(),
       #panel.background = element_blank(),
-      text = element_text()
+      text = element_text(family = pkg_options("ggplot_family"))
     ) +
     geom_hline(yintercept = 0) -> g
 

### 2020-02-07 R/fig_error_rates.R

@@ -11,7 +11,9 @@
 #' @export
 #'
 #' @examples
-#' fig_error_rates()
+#' suppressWarnings(
+#'   fig_error_rates()
+#' )
 fig_error_rates <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -132,14 +134,16 @@ fig_error_rates <- function(
       x = summaries_all_null$effect_size,
       y = 1.06,
       label = paste0("(", summaries_all_null$total, ")"),
-      size = 3
+      size = 3,
+      family = pkg_options("ggplot_family")
     ) +
     annotate(
       "text",
       x = .5,
       y = 1.15,
       label = "(Number of participants)",
-      size = 3
+      size = 3,
+      family = pkg_options("ggplot_family")
     ) +
     geom_ribbon(aes(
       ymin = p - stderr,
@@ -163,7 +167,7 @@ fig_error_rates <- function(
       panel.grid.major.x = element_blank(),
       panel.grid.minor.x = element_blank(),
       panel.background = element_blank(),
-      text = element_text()
+      text = element_text(family = pkg_options("ggplot_family"))
     ) -> g
 
 

### 2020-02-07 R/fig_evidence_scale.R

@@ -17,7 +17,9 @@
 #' @export
 #'
 #' @examples
-#' fig_evidence_scale_boxes()
+#' suppressWarnings(
+#'   fig_evidence_scale_boxes()
+#' )
 fig_evidence_scale_boxes <- function(
         filter_func = function(data, ...) return(data),
         print_plot = TRUE
@@ -63,7 +65,8 @@ fig_evidence_scale_boxes <- function(
                 ) +
                 #geom_point(position=position_dodge(width=0.06),aes(group=interaction(evidence_power, effect_size))) +
                 scale_y_continuous(trans = 'log10') +
-                theme(legend.position = "none") +
+                theme(legend.position = "none",
+                      text = element_text(family = pkg_options("ggplot_family"))) +
                 ylab("Number of samples") + ggtitle("Null") +
                 xlab("Effect size")
 
@@ -87,7 +90,8 @@ fig_evidence_scale_boxes <- function(
                         color = scales::hue_pal()(2)[2]
                 ) +
                 scale_y_continuous(trans = 'log10') +
-                theme(legend.position = "none") +
+                theme(legend.position = "none",
+                      text = element_text(family = pkg_options("ggplot_family"))) +
                 ylab("Number of samples") + ggtitle("Experimental") +
                 xlab("Effect size")
 

### 2020-02-07 R/fig_field.R

@@ -11,7 +11,9 @@
 #' @export
 #'
 #' @examples
-#' fig_field()
+#' suppressWarnings(
+#'   fig_field()
+#' )
 fig_field <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -43,7 +45,8 @@ fig_field <- function(
     xlab("Field") + ylab("Proportion of participants") +
     ylim(0, 1) +
     theme_bw() +
-    theme(legend.position = "none") +
+    theme(legend.position = "none",
+          text = element_text(family = pkg_options("ggplot_family"))) +
     geom_hline(yintercept = c(0, 1)) +
     geom_text(
       stat = "identity",
@@ -52,7 +55,8 @@ fig_field <- function(
         label = summaries * nrow(dat),
         group = 1
       ),
-      hjust = -.2
+      hjust = -.2,
+      family = pkg_options("ggplot_family")
     ) +
     coord_flip() -> g
 

### 2020-02-07 R/fig_field_scientific.R

@@ -9,7 +9,9 @@
 #' @export
 #'
 #' @examples
-#' fig_field_scientific()
+#' suppressWarnings(
+#'   fig_field_scientific()
+#' )
 fig_field_scientific <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -24,7 +26,8 @@ fig_field_scientific <- function(
     geom_bar(mapping = aes(x = is_science, y = ..prop.., group = 1)) +
     geom_text(stat = "count",
               aes(y = ..prop.., label = ..count.., group = 1),
-              hjust = -.2) +
+              hjust = -.2,
+              family = pkg_options("ggplot_family")) +
     ylim(0, 1) +
     #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
     theme_bw() +
@@ -32,7 +35,7 @@ fig_field_scientific <- function(
       panel.grid.major.y = element_blank(),
       panel.grid.minor.y = element_blank(),
       #panel.background = element_blank(),
-      text = element_text()
+      text = element_text(family = pkg_options("ggplot_family"))
     ) +
     geom_hline(yintercept = c(0, 1)) +
     xlab("Field is considered scientific") +

### 2020-02-07 R/fig_how_use_stat.R

@@ -10,7 +10,9 @@
 #'
 #' @export
 #' @examples
-#' fig_how_use_stat()
+#' suppressWarnings(
+#'   fig_how_use_stat()
+#' )
 fig_how_use_stat <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -47,7 +49,8 @@ fig_how_use_stat <- function(
     xlab("Reported use of statistics in position") + ylab("Proportion of participants") +
     ylim(0, 1) +
     theme_bw() +
-    theme(legend.position = "none") +
+    theme(legend.position = "none",
+          text = element_text(family = pkg_options("ggplot_family"))) +
     geom_hline(yintercept = c(0, 1)) +
     geom_text(
       stat = "identity",
@@ -56,7 +59,8 @@ fig_how_use_stat <- function(
         label = summaries * nrow(dat),
         group = 1
       ),
-      hjust = -.2
+      hjust = -.2,
+      family = pkg_options("ggplot_family")
     ) +
     coord_flip() -> g
 

### 2020-02-07 R/fig_preferred_method.R

@@ -10,7 +10,9 @@
 #'
 #' @export
 #' @examples
-#' fig_preferred_method()
+#' suppressWarnings(
+#'   fig_preferred_method()
+#' )
 fig_preferred_method <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -33,7 +35,8 @@ fig_preferred_method <- function(
              )) +
     geom_text(stat = "identity",
               aes(y = p, label = n, group = 1),
-              hjust = -.2) +
+              hjust = -.2,
+              family = pkg_options("ggplot_family")) +
     ylim(0, 1) +
     #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
     theme_bw() +
@@ -41,7 +44,7 @@ fig_preferred_method <- function(
       panel.grid.major.y = element_blank(),
       panel.grid.minor.y = element_blank(),
       #panel.background = element_blank(),
-      text = element_text()
+      text = element_text(family = pkg_options("ggplot_family"))
     ) +
     geom_hline(yintercept = c(0, 1)) +
     xlab("Preferred statistical method") +

### 2020-02-07 R/fig_response_evidence.R

@@ -10,7 +10,9 @@
 #'
 #' @export
 #' @examples
-#' fig_response_evidence()
+#' suppressWarnings(
+#'   fig_response_evidence()
+#' )
 fig_response_evidence <- function(filter_func = function(data, ...) return(data)) {
 
   dat <- MoreyHoekstra2019::christmas_stats_participants %>%

### 2020-02-07 R/fig_sampling_behavior.R

@@ -14,7 +14,9 @@
 #'
 #' @export
 #' @examples
-#' fig_sampling_behavior()
+#' suppressWarnings(
+#'   fig_sampling_behavior()
+#' )
 fig_sampling_behavior <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -40,7 +42,8 @@ fig_sampling_behavior <- function(
     ) +
     scale_y_continuous(trans = "log10") +
     ylab("Null samples") +
-    xlab("Effect size") -> g1
+    xlab("Effect size") +
+    theme(text = element_text(family = pkg_options("ggplot_family"))) -> g1
 
   dat %>%
     MASS::rlm(log(n_expt) ~ abs(effect_size), data = .) -> rlm_obj_expt
@@ -60,7 +63,8 @@ fig_sampling_behavior <- function(
     ) +
     scale_y_continuous(trans = "log10") +
     ylab("Experimental samples") +
-    xlab("Effect size") -> g2
+    xlab("Effect size") +
+    theme(text = element_text(family = pkg_options("ggplot_family"))) -> g2
 
   g0 <- ggpubr::ggarrange(
     g1,

### 2020-02-07 R/fig_shuffle_understanding.R

@@ -10,7 +10,9 @@
 #' @export
 #'
 #' @examples
-#' fig_shuffle_understanding()
+#' suppressWarnings(
+#'   fig_shuffle_understanding()
+#' )
 fig_shuffle_understanding <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -24,7 +26,8 @@ fig_shuffle_understanding <- function(
     geom_bar(mapping = aes(x = shuffle, y = ..prop.., group = 1)) +
     geom_text(stat = "count",
               aes(y = ..prop.., label = ..count.., group = 1),
-              hjust = -.2) +
+              hjust = -.2,
+              family = pkg_options("ggplot_family")) +
     ylim(0, 1) +
     #coord_cartesian(ylim = c(0,1), expand = FALSE, clip = "off") +
     theme_bw() +
@@ -32,7 +35,7 @@ fig_shuffle_understanding <- function(
       panel.grid.major.y = element_blank(),
       panel.grid.minor.y = element_blank(),
       #panel.background = element_blank(),
-      text = element_text()
+      text = element_text(family = pkg_options("ggplot_family"))
     ) +
     geom_hline(yintercept = c(0, 1)) +
     xlab("Understood use of shuffle reports") +

### 2020-02-07 R/fig_sig_testing_opinion.R

@@ -10,7 +10,9 @@
 #'
 #' @export
 #' @examples
-#' fig_sig_testing_opinion()
+#' suppressWarnings(
+#'   fig_sig_testing_opinion()
+#' )
 fig_sig_testing_opinion <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -46,7 +48,8 @@ fig_sig_testing_opinion <- function(
     xlab("Opinion about significance testing") + ylab("Proportion of participants") +
     ylim(0, 1) +
     theme_bw() +
-    theme(legend.position = "none") +
+    theme(legend.position = "none",
+          text = element_text(family = pkg_options("ggplot_family"))) +
     geom_hline(yintercept = c(0, 1)) +
     geom_text(
       stat = "identity",
@@ -55,7 +58,8 @@ fig_sig_testing_opinion <- function(
         label = summaries * nrow(dat),
         group = 1
       ),
-      hjust = -.2
+      hjust = -.2,
+      family = pkg_options("ggplot_family")
     ) +
     coord_flip() -> g
 

### 2020-02-07 R/fig_strategies.R

@@ -14,7 +14,9 @@
 #' @export
 #'
 #' @examples
-#' fig_strategies()
+#' suppressWarnings(
+#'   fig_strategies()
+#' )
 fig_strategies <- function(
   filter_func = function(data, ...) return(data),
   print_plot = TRUE
@@ -57,7 +59,7 @@ fig_strategies <- function(
     coord_flip() +
     theme_classic() +
     theme(legend.position = "top",
-          text = element_text(size=12),
+          text = element_text(size=12, family = pkg_options("ggplot_family")),
           plot.margin = unit(c(0,1,0,0), "cm")) +
     xlab("Sig. test content in text") -> g
 

### 2020-02-07 R/pkg_options.R

@@ -1,0 +1,24 @@
+# Variable, global to package's namespace. 
+# This function is not exported to user space and does not need to be documented.
+MYPKGOPTIONS <- settings::options_manager(
+  ggplot_family = "Lato Light"
+)
+
+# User function that gets exported:
+
+#' Set or get options for my package
+#' 
+#' @param ... Option names to retrieve option values or \code{[key]=[value]} pairs to set options.
+#'
+#' @section Supported options:
+#' The following options are supported
+#' \itemize{
+#'  \item{\code{ggplot_family}}{(\code{character};"Lato Light") Font family to use for ggplot2 graphs }
+#' }
+#'
+#' @export
+pkg_options <- function(...){
+  # protect against the use of reserved words.
+  settings::stop_if_reserved(...)
+  MYPKGOPTIONS(...)
+}

### 2020-02-07 inst/build_site.R

@@ -16,7 +16,7 @@ which_dirs = c("MoreyHoekstra2019/inst/pkg_html")
 
 pkg_file = devtools::build( pkg = here(),
                             path = tmpdir,
-                            quiet = TRUE )
+                            quiet = FALSE )
 
 untar( pkg_file,
        files = c( which_files, which_dirs ),

### 2020-02-08 R/pkg_options.R
enabled font switching for base and ggplot fonts

@@ -1,7 +1,8 @@
 # Variable, global to package's namespace. 
 # This function is not exported to user space and does not need to be documented.
 MYPKGOPTIONS <- settings::options_manager(
-  ggplot_family = "Lato Light"
+  ggplot_family = "Lato Light",
+  base_family   = "lato"
 )
 
 # User function that gets exported:
@@ -14,6 +15,7 @@ MYPKGOPTIONS <- settings::options_manager(
 #' The following options are supported
 #' \itemize{
 #'  \item{\code{ggplot_family}}{(\code{character};"Lato Light") Font family to use for ggplot2 graphs }
+#'  \item{\code{base_family}}{(\code{character};"lato") Font family to use for base graphs }
 #' }
 #'
 #' @export

### 2020-02-08 inst/build_site.R

@@ -1,12 +1,10 @@
 library(here)
 
-pkgdown::build_site( here() )
+pkgdown::build_site( here::here() )
 
 rmarkdown::render( input = here("README.Rmd"),
                    output_file = here("README.md"),
                    output_format = "md_document")
-
-tmpdir = tempdir()
 
 which_files = c("MoreyHoekstra2019/inst/doc/manuscript.pdf",
                 "MoreyHoekstra2019/inst/doc/supplementA.html",
@@ -14,9 +12,11 @@ which_files = c("MoreyHoekstra2019/inst/doc/manuscript.pdf",
                 "MoreyHoekstra2019/inst/reprohack/reprohack2019_feedback_letter.pdf")
 which_dirs = c("MoreyHoekstra2019/inst/pkg_html")
 
-pkg_file = devtools::build( pkg = here(),
+tmpdir = tempdir()
+
+pkg_file = devtools::build( pkg = here::here(),
                             path = tmpdir,
-                            quiet = FALSE )
+                            quiet = TRUE )
 
 untar( pkg_file,
        files = c( which_files, which_dirs ),

### 2020-02-08 inst/font/font_setup.R

@@ -1,12 +1,7 @@
 library(showtext)
+library(MoreyHoekstra2019)
 
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>",
-  fig.showtext=TRUE,
-  dpi=300
-)
-
+## Set up lato font
 font_paths(
   system.file("font/", package = "MoreyHoekstra2019")
 )
@@ -18,9 +13,17 @@ font_add("lato",
          bolditalic = "Lato-SemiboldItalic.ttf",
          symbol = "Lato-LightItalic.ttf")
 
+# Set default font
 knitr::knit_hooks$set(
-  lato = function(before, options, envir)
-    if (before) par(family = "lato")
+  MH2019 = function(before, options, envir)
+    if (before) par(family = MoreyHoekstra2019::pkg_options("base_family"))
 )
 
-knitr::opts_chunk$set(lato = TRUE)
+knitr::opts_chunk$set(
+  collapse = TRUE,
+  comment = "#>",
+  fig.showtext=TRUE,
+  dpi=300,
+  ### This sets the font use
+  MH2019 = TRUE 
+)

### 2020-05-20 inst/manuscript/manuscript.Rmd
enabled font switching for base and ggplot fonts
moved to RSOS format

--- 2020-02-05 inst/manuscript/manuscript.Rmd
@@
 title: Use of significance test logic by scientists in a novel reasoning task
-titlerunning: Statistical reasoning in scientists
-authorrunning: Morey & Hoekstra
-authors: 
+author:
   - name: Richard D. Morey
+    affiliation: "1"
+  - name: Rink Hoekstra
+    affiliation: "2"
+address:
+  - code: "1"
     address: School of Psychology, Cardiff University
-  email: moreyr@cardiff.ac.uk
-- name: Rink Hoekstra
+  - code: "2"
     address: Faculty of Behavioural and Social Sciences, University of Groningen
-  email: r.hoekstra@rug.nl
+corresp_author_name:  "Richard D. Morey"
+corresp_author_email: "moreyr@cardiff.ac.uk"
+
+subject:
+  - "subject 1"
+  - "subject 2"
+  - "subject 3"
+
 keywords:
-- cognitive science
-- psychology
-- science
+  - reasoning
   - statistics
   - statistical cognition
-#PACS:
+  - decision making
+  - significance testing
 
-MSC:
-- 91E10    
 abstract: |
@@
 This suggests that scientists may not be as statistically-challenged as often believed, and that reforms should take this into account.
+
+ethics: |
+  This research project was evaluated by the Cardiff University School of Psychology (application number EC.18.12.11.5526G).
+It was found to be within the ethical guidelines for experiments with human participants.
+All participants gave informed consent prior to their participation.
+
+data_accessibility: |
+  All materials, code, and data are freely available at https://github.com/richarddmorey/Morey_Hoekstra_StatCognition.
+
+author_contributions: |
+  RDM conceptualized and designed the study in consultation with RH. RDM analysed the data and created the materials and figures. The manuscript was written by RDM and RH.
+
+conflict_of_interest: |
+  The authors declare no conflicts of interest.
+
 bibliography: "`r system.file('bib/bibfile.bib', package = 'MoreyHoekstra2019')`"
-csl: nature.csl
-output: rticles::springer_article
+
+lineno: true
+
+site: bookdown::bookdown_site
+output: 
+  bookdown::pdf_book:
+    base_format: rticles::rsos_article
+pkgdown:
+  as_is: true
+  extension: pdf
 vignette: >
@@
   chunk_output_type: console
-#nocite: | 
-#  @Venables:Ripley:2002
 header-includes: |
@@
   \renewcommand{\thesubfigure}{\Alph{subfigure}}
-pkgdown:
-  as_is: true
-  extension: pdf
 ---
@@
                       message = FALSE,
-                      out.width = "0.95\\textwidth")
+                      out.width = "0.95\\textwidth",
+                      dev = "cairo_pdf",
+                      fig.width = 6,
+                      fig.height = 4)
 
@@
 Methodologists and statistical cognition researchers point to evidence from questionnaires and vignette studies to argue that researchers do not, in fact, grasp the core logic of SST.
-In one highly influential study of research psychologists, Oakes [-@Oakes:1986] presented six statements about a hypothetical significance test result to be categorized as true or false (e.g., "[The $p$ value provides] the probability of the null hypothesis being true").
+In one highly influential study of research psychologists, Oakes [@Oakes:1986] presented six statements about a hypothetical significance test result to be categorized as true or false (e.g., "[The $p$ value provides] the probability of the null hypothesis being true").
 Despite all of these statements being false, 97% of the research psychologists categorized at least one as true.
@@
 82).
+
+\EndFirstPage
 
@@
 The $x$-axis monotonically (but nonlinearly) related to the strength of the statistical evidence ($z$ statistic) favoring one group; the $y$-axis is monotonically (but nonlinearly) related to the sample size.
-Underlying numerical values of the statistical evidence and sample sizes were unknown to the participant."}
+Underlying numerical values of the statistical evidence and sample sizes were unknown to the participant.", out.width=".9\\columnwidth"}
 fn = system.file("pkg_html/img/clean_interface.png", package = "MoreyHoekstra2019")
@@
 ```
-
 Participants were also offered the opportunity to sample "random shuffle reports" that were described as the results of experiments with random assignment of elves to groups: that is, the result of experiments in which the null hypothesis was true.
@@
 
-```{r error_rates, fig.cap="Decision error rates as a function of true effect size.
+```{r error-rates, fig.cap="Decision error rates as a function of true effect size.
 The red solid line represents the rate at which participants {\\em correctly} determined the direction of the effect, except for at an effect size of 0, where it represents an incorrect claim of an effect.
 The shaded region around the solid line is one standard error wide.
-The difference between the black and the gray dashed lines shows the proportion of participants who decided that the two groups were the same (as opposed to failing to detect a difference)."}
+The difference between the black and the gray dashed lines shows the proportion of participants who decided that the two groups were the same (as opposed to failing to detect a difference).", fig.showtext=FALSE, out.width=".9\\columnwidth"}
 MoreyHoekstra2019::fig_error_rates()
@@
 
-Decision error rates as a function of true effect size are shown in Figure \ref{fig:error_rates}.
+Decision error rates as a function of true effect size are shown in Figure \ref{fig:error-rates}.
 
@@
 
-```{r response_ev, fig.cap="Statistical evidence underlying participants' decisions.
+```{r response-ev, fig.cap="Statistical evidence underlying participants' decisions.
 The Wilcoxon rank-sum test $p$ value ($x$ axis) between the participants' null samples and experimental samples is used as a rough index of evidential strength.
@@
 Filled circles show correct decisions; hollow circles, incorrect decisions.
-The two hollow squares show sign errors."}
+The two hollow squares show sign errors.", out.width=".9\\columnwidth"}
 MoreyHoekstra2019::fig_response_evidence()
@@
 
-Figure \ref{fig:response_ev} shows the distribution of Wilcoxon $p$ values (arranged by the direction of the decision).
+Figure \ref{fig:response-ev} shows the distribution of Wilcoxon $p$ values (arranged by the direction of the decision).
 Kernel density estimates show the distributions of $p$ values when participants indicated that Sparklies were faster, no detection/same, or that Jinglies were faster.
@@
 
-```{r strategy_table}
+```{r strategy-table}
 MoreyHoekstra2019::christmas_stats_participants %>% 
@@
 
-```{r response_ev2, fig.height=4, fig.cap="The effect of the evidence transformation manipulation on responding.
+```{r response-ev2, fig.height=4, fig.cap="The effect of the evidence transformation manipulation on responding.
 Points on top (narrow scale; $q=7$) and bottom (wide scale; $q=3$) represent participants' decisions as a function of the most extreme experiment sampled.
@@
 Curves show predicted probability by a logistic regression fit with standard errors.", fig.show = "hold", fig.subcap = c("Predicted response probabilities relative to visual extremity.
-Vertical lines show the critical 0.05 for the corresponding null sampling distribution.","Predicted response probabilities relative to the null sampling distributions (implicit $p$ values)."), out.width = "0.90\\textwidth", fig.ncol = 1}
+Vertical lines show the critical 0.05 for the corresponding null sampling distribution.","Predicted response probabilities relative to the null sampling distributions (implicit $p$ values)."), out.width=".9\\columnwidth", fig.ncol = 1}
 glm_obj_x <- MoreyHoekstra2019::fig_evidence_scale_logistic_x()
@@
 
-Figure \ref{fig:response_ev2} (top) shows responses (no detect/same or Jinglies/Sparklies) as a function of the most extreme experiment sampled ($x$ axis) and the transformation.
+Figure \ref{fig:response-ev2} (top) shows responses (no detect/same or Jinglies/Sparklies) as a function of the most extreme experiment sampled ($x$ axis) and the transformation.
 There was a strong effect of the transformation consistent with use of the null sampling distribution; participants randomly assigned to the "narrow" evidence transformation responded "Jinglies/Sparklies" for much less visually extreme evidence (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_x[3:4,3]),3)`, p<.001$).
@@
 
-Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response_ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).
+Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response-ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).
 It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with a small number of participants incorrectly using the nondiagnostic visual extremeness to perform the task.
@@
 
-# References
-
-<div id="refs"></div>
-
-# End notes
-
-* **Data and materials availability**: All materials, code, and data are freely available at https://github.com/richarddmorey/Morey_Hoekstra_StatCognition.
-
-* **Author Contributions**: RDM conceptualized and designed the study in consultation with RH. RDM analysed the data and created the materials and figures. The manuscript was written by RDM and RH.
-
-* The authors declare no competing interests.
-
-* Correspondence and requests for materials should be addressed to Richard D. Morey (richarddmorey@gmail.com).
-
 * Compiled `r Sys.time()` (`r Sys.timezone()`) under `r version$version.string`.

### 2020-05-21 inst/manuscript/manuscript.Rmd
fixed issue with pandoc compilation of vignettes

@@
 
-bibliography: "`r system.file('bib/bibfile.bib', package = 'MoreyHoekstra2019')`"
+bibliography: "``r system.file('bib/bibfile.bib', package = 'MoreyHoekstra2019')``{=latex}"
 
@@
 As Table \ref{tab:strategies} shows that a large majority of participants (`r for_table[1,1]`, `r round( 100*as.numeric(for_table[1,1]) / as.numeric(for_table[1,6]), 2)`%) indicated using strong significance testing strategies.
-We should be cautious, however, because participants were told in the instructions that the shuffle reports could be used for assessing sampling variability.
+We should be cautious in directly interpreting this high number alone, however, because participants were told in the instructions that the shuffle reports could be used for assessing sampling variability.
 We did this to make clear what the shuffle reports were, but without explaining how to use them.
@@
 
-The instructions did not tell the participants *how* they should use the shuffles reports, yet many participants gave detailed accounts.
-Moreover, as we show next, participants' decisions were sensitive to a key manipulation of the shuffles reports in the manner that one would expect if they were using SST logic.
+Secondly---and most importantly---the instructions did not tell the participants *how* they should use the shuffles reports, yet many participants gave detailed accounts.
+As we show next, participants' decisions were sensitive to a key manipulation of the shuffles reports in the manner that one would expect if they were using SST logic.
 

### 2020-05-21 inst/manuscript/manuscript.Rmd
manuscript edits for clarity

@@
 subject:
-  - "subject 1"
-  - "subject 2"
-  - "subject 3"
+  - "Cognition and decision making"
+  - "Psychology and cognitive neuroscience"
+  - "Statistics"
 
@@
 
-Despite statistical significance testing (SST) use in a majority of research projects across fields, there is debate over whether scientists understand SST and can use it competently.
+For most of the past century, the dominant method of statistical inference has been statistical significance testing (SST).
+In a significance test, the statistical evidence in the form of a test statistic is compared to what would be expected under a particular hypothesis (often called the "null" hypothesis).
+If it would be rare to observe evidence as strong as what was observed under this hypothesis, the strength of the evidence is deemed surprising enough to call the assumed hypothesis into doubt.
+The rarity of evidence as strong as what was observed under the assumed hypothesis---the so-called $p$ value---is the typical way that results of significance tests are reported.
+The key feature of SST for our purposes is the assessment of evidence by means of comparing a result to a "null" distribution.
+
+\EndFirstPage
+
+Despite the use of SST in a majority of research projects across fields, there is debate over whether scientists understand SST and can use it competently.
 Methodologists and statistical cognition researchers point to evidence from questionnaires and vignette studies to argue that researchers do not, in fact, grasp the core logic of SST.
@@
 82).
-
-\EndFirstPage
 
@@
 
-```{r error-rates, fig.cap="Decision error rates as a function of true effect size.
+```{r error-rates, fig.cap="Correct decision and error rates as a function of true effect size.
 The red solid line represents the rate at which participants {\\em correctly} determined the direction of the effect, except for at an effect size of 0, where it represents an incorrect claim of an effect.
@@
 
-```{r response-ev2, fig.height=4, fig.cap="The effect of the evidence transformation manipulation on responding.
+```{r response-pvals, fig.cap="Selected two-tailed $p$ values from the null sampling distributions of the narrow (top) and wide (bottom) conditions, projected onto the experimental interface.
+Participants were not shown the $p$ values; they had to intuit them by using samples from the null distribution (``random shuffle reports'').", fig.show = "hold", fig.subcap = c("Implied $p$ values for the ``narrow'' condition.","Implied $p$ values for the ``wide'' condition."), out.width=".9\\columnwidth", fig.ncol = 1}
+fn = MoreyHoekstra2019::fig_evidence_p_vals(q = 7, letter = "")
+knitr::include_graphics(fn)
+fn = MoreyHoekstra2019::fig_evidence_p_vals(q = 3, letter = "")
+knitr::include_graphics(fn)
+```
+
+```{r response-ev2, fig.cap="The effect of the evidence transformation manipulation on responding.
 Points on top (narrow scale; $q=7$) and bottom (wide scale; $q=3$) represent participants' decisions as a function of the most extreme experiment sampled.
+See the methods details for the interpretation of $q$.
 ``N'' indicates a ``no detect'' or ``same'' response; ``J/S'' indicates a response in favor of a difference between the groups.
@@
 In one condition ("wide"), the null sampling distribution was visually about twice as wide as in the other ("narrow").
-If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is visually wider.
+If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider.
+
+The effects of the manipulation are shown in figure \ref{fig:response-pvals}.
 

### 2020-05-21 R/fig_evidence_scale.R

@@ -139,7 +139,7 @@ fig_evidence_scale_logistic_x <- function(pval = .05,
              xlab = "Most extreme evidence / x location",
              ylab = "Prob. Sparklies/Jinglies responses", xaxs = 'i',
              yaxs = 'i', axes=FALSE)
-        legend("left",legend = c("Wide (3)","Narrow (7)"), col = colors_ggplot,
+        legend("left",legend = c("Wide","Narrow"), col = colors_ggplot,
                lwd = 2, bty = 'n', lty = 1:2)
         mtext("Most extreme evidence / x location", 3, par()$mgp[1])
 
@@ -258,7 +258,7 @@ fig_evidence_scale_logistic_p <- function(pval = .05,
              xlab = "Most extreme p value",
              ylab = "Prob. Sparklies/Jinglies response", xaxs = 'i',
              yaxs = 'i', axes=FALSE)
-        legend("left",legend = c("Wide (3)","Narrow (7)"), col = colors_ggplot,
+        legend("left",legend = c("Wide","Narrow"), col = colors_ggplot,
                lwd = 2, bty = 'n', lty = 1:2)
         mtext("Most extreme p value", 3, par()$mgp[1])
 

### 2020-05-21 inst/manuscript/manuscript.Rmd
minor changes to manuscript
small changes to manuscript

@@
 We adopt a similar strategy to test statistical reasoning: we eliminate numerical information from statistical results to test scientists' ability to interpret results with reference to a null sampling distribution (i.e., SST logic).
-Without numerical information, many other strategies and heuristics, such as confidence intervals or Bayesian inference, are difficult or impossible to apply.
+Without numerical information, many other strategies and heuristics, such as confidence intervals or Bayesian inference, are difficult or impossible to apply.[^1]
+
+[^1]: A formal statistical explanation showing that the task is difficult or impossible to perform using non-SST logic is given in Section 3 of Supplement A.
 
@@
 ```
+
 Participants were also offered the opportunity to sample "random shuffle reports" that were described as the results of experiments with random assignment of elves to groups: that is, the result of experiments in which the null hypothesis was true.
@@
 
-All materials, code, and data are freely available at https://github.com/richarddmorey/Morey_Hoekstra_StatCognition.
-
 ## Evidence distributions

### 2020-05-26 inst/manuscript/manuscript.Rmd
small changes to manuscript

@@
 
-[^1]: A formal statistical explanation showing that the task is difficult or impossible to perform using non-SST logic is given in Section 3 of Supplement A.
+[^1]: A formal statistical explanation showing that the task is difficult or impossible to perform using non-SST logic is given in Section 3 of [Supplement A](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/man_supp.html).
 
@@
 Participants were not told how to use these samples, only that they might use them.
+
+Our experiment was contructed such that the only way to assess the evidence in the data was by comparison to a null sampling distribution: either the one provided by the random shuffle reports, or a simpler null that assumes that the evidence will favor one team or the other with 50% probability.
+Thus, the information afforded only the information in a $p$ value, but it was not described as such, and participants had to *discover* for themselves how to use the informaiton.
 
@@
 In one condition ("wide"), the null sampling distribution was visually about twice as wide as in the other ("narrow").
+
+The effects of the manipulation are shown in Figure \ref{fig:response-pvals}.
+The two experimental conditions used different arbitrary monotone mappings from the underlying $Z$-statistic to the visual space.
+Intuitively, this would be like deciding to use $Z^3$ instead of $Z$ in all $Z$ tests; one would need to adjust the significance criteria to account for the cubing (e.g., use $|1.96^3|=7.53$ instead of $|1.96|$ for a $\alpha=0.05$ level test), but the underlying test remains the same.
+The manipulation changes only the visual impression of the sampling distributions, allowing us to see how sensitive their responses are to the null sampling distribution as represented by the random shuffle reports.
+
 If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider.
-
-The effects of the manipulation are shown in figure \ref{fig:response-pvals}.
+When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant, given the $p$ value.
 
@@
 
+There are some major differences between our methods and the methods of previous demonstrations that scientists fail when reasoning about SST.
+We have already mentioned the fact that previous demonstrations typically use survey vignettes, as opposed to our more engaged experimental method.
+A second difference is our sample which is a substantially larger and more diverse convenience sample than previous demonstrations.
+It is possible that the recent so-called "replication crisis" [@Baker:2016] has raised awareness among scientists of deficits of statistical reasoning.
+If this increase in awareness accounts for our more optimistic findings, this would be good news for statistical educators.
+
 Our findings echo other work showing that human reasoning can, under some conditions, be better than previously understood.
@@
 In spite of scientists' real-life statistical behaviour often resembling a "ritual" [@Gigerenzer:etal:2004], when we eliminate the ritual --- no $p$ value, or any other familiar number, was offered --- they think statistically, very often arriving at the correct conclusion about the sign of the effect.
-Effective methodological reform in the sciences may be easier than abandoning significance testing, if we can harness scientists' already existing statistical competence.
+Effective methodological reform in the sciences may be easier than abandoning significance testing, if we can harness scientists' already-existing statistical competence.
 
@@
 Data inclusion criteria included sampling at least one shuffle report and experimental result, working in a scientific field, having at least some University education in science, and that it was their first time participating.
-Details are given in Supplement B.
+Details are given in [Supplement B](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/man_supp.html).
 
@@
 The probability of assignment to either evidence power was 50%.
-Details are given in Table 1.1 in Supplement A.
+Details are given in Table 1.1 in [Supplement A](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/man_supp.html).
 
@@
 After a brief recap of the instructions, participants performed the main task --- sampling either random shuffles or experiments --- until they made a decision about which, if either, elf group was faster.
-They were then asked several open-ended questions about their strategy, some informational questions (results in Supplement B) and debriefed.
+They were then asked several open-ended questions about their strategy, some informational questions (results in [Supplement B](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/man_supp.html)) and debriefed.
 
@@
 This arbitrary transformation was done to ensure that the test statistic's distribution was unfamiliar to the participant.
-See Supplement A for more details, including graphical depictions of the evidence distributions.
+See [Supplement A](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/man_supp.html) for more details, including graphical depictions of the evidence distributions.
 
@@
 The remainder of the disagreements were re-coded separately, and a final round of discussion resolved the remaining disagreements.
-Participants' codings described in detail in Supplement B.
+The coding of participants' responses is described in detail in [Supplement B](https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/man_supp.html).
 

### 2020-05-31 inst/manuscript/supplementA.Rmd
compiled after upgrade to R 4.0

@@
 (Figure B above) shows the distribution of the number of experiments sampled as a function of the true effect size.
-Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`).
+Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print.htest(kt_expt)$statistic`).
 When the effect size is small and difficult to detect, participants do more experimentation before deciding.

### 2020-06-01 R/fig_evidence_p_vals.R
fixed some problems installing on Windows

@@ -22,7 +22,7 @@ fig_evidence_p_vals <- function(q,
                                 ps = c(.005, .025, .05, .125, .25, .5, .75, .875, .95, .975, .995),
                                 letter = ""){
 
-  tf = tempfile(fileext = ".png")
+  tf = gsub( "\\\\", "/", tempfile(fileext = ".png") )
 
   # ima is a global variable
   img_dim = c(dim(ima)[2], dim(ima)[1])

### 2020-06-01 R/fig_recreate_display.R

@@ -23,7 +23,8 @@ fig_recreate_display <- function(id,
                                  ps =
                                    c(.005, .025, .05, .125, .25, .5, .75, .875, .95, .975, .995),
                                  letter = "") {
-  tf = tempfile(fileext = ".png")
+  tf = gsub( "\\\\", "/", tempfile(fileext = ".png") )
+  
 
   dat <- MoreyHoekstra2019::christmas_stats_participants
   samples <- MoreyHoekstra2019::christmas_stats_samples

### 2020-06-01 README.Rmd
small changes ahead of resubmission
recompiled site; added info about potential install failures to readme

@@
 ``` r
-devtools::install_github("richarddmorey/Morey_Hoekstra_StatCognition", build_vignettes = TRUE, depedencies = TRUE)
+devtools::install_github("richarddmorey/Morey_Hoekstra_StatCognition", build_vignettes = TRUE, dependencies = TRUE)
 ```
+
+### Problems installing
+
+If you have difficulties installing the package due to failure to compile the vignettes &mdash; particularly on Windows &mdash; you can try setting `build_vignettes` to `FALSE` in the command above.
+You can also move your R installation and temp folder to a location without spaces.
+This is a good idea even if turning off compilation of the vignettes solves the immediate problem.
+See [here](https://lcolladotor.github.io/2019/09/18/windows-user-space-issues-with-installing-r-packages/) for details.
 

### 2020-06-01 inst/build_site.R

@@ -1,10 +1,11 @@
 library(here)
 
-pkgdown::build_site( here::here() )
-
 rmarkdown::render( input = here("README.Rmd"),
                    output_file = here("README.md"),
                    output_format = "md_document")
+
+pkgdown::build_site( here::here() )
+
 
 which_files = c("MoreyHoekstra2019/inst/doc/manuscript.pdf",
                 "MoreyHoekstra2019/inst/doc/supplementA.html",

### 2020-06-02 inst/manuscript/manuscript.Rmd
minor manuscript changes; ready for submission

@@
 In a significance test, the statistical evidence in the form of a test statistic is compared to what would be expected under a particular hypothesis (often called the "null" hypothesis).
-If it would be rare to observe evidence as strong as what was observed under this hypothesis, the strength of the evidence is deemed surprising enough to call the assumed hypothesis into doubt.
+If it would be surprising to observe evidence as strong as what was observed under this hypothesis, the evidence is strong deemed enough to call the assumed hypothesis into doubt, at least tentatively [see also @Dempster:1964;@Greenland:2019].
 The rarity of evidence as strong as what was observed under the assumed hypothesis---the so-called $p$ value---is the typical way that results of significance tests are reported.
@@
 
-In tests of perception --- e.g., of colorblindness [@Ishihara:1972] --- it is common to eliminate one cue (e.g., brightness) in order assess the ability to use another (e.g., color).
+In tests of perception---e.g., of colorblindness [@Ishihara:1972]---it is common to eliminate one cue (e.g., brightness) in order assess the ability to use another (e.g., color).
 If color is the only useful cue for reading a number, deficits in color vision make the number difficult to read.
@@
 
-Our experiment was contructed such that the only way to assess the evidence in the data was by comparison to a null sampling distribution: either the one provided by the random shuffle reports, or a simpler null that assumes that the evidence will favor one team or the other with 50% probability.
-Thus, the information afforded only the information in a $p$ value, but it was not described as such, and participants had to *discover* for themselves how to use the informaiton.
+Our experiment was constructed such that the only way to assess the evidence in the data was by comparison to a null sampling distribution: either the one provided by the random shuffle reports, or a simpler null that assumes that the evidence will favor one team or the other with 50% probability.
+Thus, the information afforded only the information in a $p$ value, but it was not described as such, and participants had to *discover* for themselves how to use the information.
 
@@
 Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response-ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).
-It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with a small number of participants incorrectly using the nondiagnostic visual extremeness to perform the task.
+It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with a small number of participants incorrectly using the non-diagnostic visual extremeness to perform the task.
 If more people had been fooled by the irrelevant visual manipulation of the sampling distribution, we would expect this effect to be substantially larger.

### 2020-06-04 inst/manuscript/manuscript.Rmd
minor additions for RSOS

@@
 data_accessibility: |
-  All materials, code, and data are freely available at https://github.com/richarddmorey/Morey_Hoekstra_StatCognition.
+  Data and relevant code for this research work are stored in GitHub: https://github.com/richarddmorey/Morey_Hoekstra_StatCognition and have been archived within the Zenodo repository: https://doi.org/10.5281/zenodo.3877106
 
@@
 
+* **Funding**: This research was not supported by external funding.
+
 * Compiled `r Sys.time()` (`r Sys.timezone()`) under `r version$version.string`.

### 2020-08-15 inst/manuscript/manuscript.Rmd
fixed issues with compilations in updated dependencies

@@
 glm_anova_x = broom::tidy(anova(glm_obj_x, test = "Chisq"))
-glm_apa_x = papaja::apa_print.glm(glm_obj_x)
+glm_apa_x = papaja::apa_print(glm_obj_x)
 
 pred_p = outer(c(3,7), c(3,7), function(true, z_calc){
-  plogis(predict.glm(glm_obj_x, 
+  plogis(predict(glm_obj_x, 
               newdata = list(ev = MoreyHoekstra2019:::evz(qnorm(1 - pval / 2), z_calc), 
@@
 glm_anova_p = broom::tidy(anova(glm_obj_p, test = "Chisq"))
-glm_apa_p = papaja::apa_print.glm(glm_obj_p)
+glm_apa_p = papaja::apa_print(glm_obj_p)
 

### 2020-08-15 inst/manuscript/supplementA.Rmd
fixed issues with compilations in updated dependencies

--- 2020-02-07 inst/manuscript/supplementA.Rmd
@@
 (Figure B above) shows the distribution of the number of experiments sampled as a function of the true effect size.
-Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`).
+Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja:::apa_print.htest(kt_expt)$statistic`).
 When the effect size is small and difficult to detect, participants do more experimentation before deciding.

### 2020-09-02 inst/manuscript/manuscript.Rmd
first edits for revision

@@
 In a significance test, the statistical evidence in the form of a test statistic is compared to what would be expected under a particular hypothesis (often called the "null" hypothesis).
-If it would be surprising to observe evidence as strong as what was observed under this hypothesis, the evidence is strong deemed enough to call the assumed hypothesis into doubt, at least tentatively [see also @Dempster:1964;@Greenland:2019].
+If it would be surprising to observe evidence as strong as what was observed under this hypothesis, the evidence is deemed strong enough to call the assumed hypothesis into doubt, at least tentatively [see also @Dempster:1964;@Greenland:2019].
 The rarity of evidence as strong as what was observed under the assumed hypothesis---the so-called $p$ value---is the typical way that results of significance tests are reported.
@@
 However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication).
-It is unclear how well vignette studies (including ones by the present authors [@Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results.
+It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results.
 Conceptual understanding and fluency with common representations are both important, but are distinct.
 
-A second major piece of evidence for misunderstandings of SST logic is reasoning errors in published papers.
-[@Gelman:Stern:2006;@Hoekstra:etal:2006;@Nieuwenhuis:etal:2011;@Weisburd:etal:2003].
+A second major piece of evidence for misunderstandings of SST logic is reasoning errors in published papers [@Gelman:Stern:2006;@Hoekstra:etal:2006;@Nieuwenhuis:etal:2011;@Weisburd:etal:2003].
 Like evidence from vignette studies, however, these errors are difficult to interpret as misunderstandings of SST logic *per se*.
-To avoid conflating problems with statistical reasoning with motivated scientific reasoning, one must test scientists' statistical reasoning where these incentives are not in play.
+These examples show that whatever process lead to the statistical conclusion was flawed in some way, but many processes contribute to such conclusions.
+Cognitive (cites), technological (cites), and social processes (cites) have all been assigned some blame for statistical reasoning failures.
 
-Using a simulated experimental task, we tested researchers' and trainees' understanding and application of SST.
-Our key innovation allowing us to focus on SST was to prevent---or make difficult---the use of alternative strategies.
-If participants have poor understanding of SST, one would predict they would perform poorly and would not be sensitive to SST-related information.
-In fact, scientists performed well, explicitly reported using SST strategies, and were sensitive to SST-related information.
-Our results strongly suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.
+In deciding how to improve statistical reasoning, it is crucial to know where the problems are.
+The *fact* of reasoning problems tells us little about their *source*.
+In assessing potential interventions, however, the source is crucial.
+Some interventions might focus on the social aspects (e.g., decreasing the need for "significant" results for prestige), some on technological aspects (e.g., presenting statistical results in ways that were previously impossible), and some on cognitive aspects (e.g., adopting Bayesian procedures because these are claimed to be better understood).
+
+Instead of the usual human-factors approach, which adopts familiar statistical language and tests participants' fluency with existing procedures, we adopt a more cognitive perspective.
+We aim to test working scientists' understanding of the basic conceptual framework underlying SST using a simulated experimental task.
+To avoid conflating basic reasoning processes and fluency with common statistical terminology, we avoid using common statistics---or, indeed, any numbers---at all.
+
+The key innovation allowing us to focus on SST reasoning was to design an experiment that prevents the use of alternative strategies.
+A key feature of SST is that the use of a null distribution destroys information about effect sizes and sample size.
+In fact, this aspect of SST reasoning is often criticized.
+Alternative methodologies often focus on effect sizes (point estimates, confidence intervals, equivalence, likelihood, Bayesian priors/posteriors).
+We offered our participants only the information in a $p$ value, and participants had to understand or discover how to obtain that information.
+Their task was to use this information to come to a decision about the true sign of an effect through repeated experimentation.
+
+If participants have poor understanding of SST, they would 1) often come to the wrong conclusion, in spite of ample information; 2) show error rates that are only weakly associated with true effect size; 3) be unable to articulate strategies for performing our task; 4) be sensitive to misleading, task-irrelevant information; 5) be insensitive to SST-relevant information.
+The scientists in our sample showed often came to the right conclusion, with their performance showing that they were sensitive to the information they were given.
+They explicitly reported using SST strategies.
+Our results suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.
 
@@
 We adopt a similar strategy to test statistical reasoning: we eliminate numerical information from statistical results to test scientists' ability to interpret results with reference to a null sampling distribution (i.e., SST logic).
-Without numerical information, many other strategies and heuristics, such as confidence intervals or Bayesian inference, are difficult or impossible to apply.[^1]
+Without numerical information, many other strategies and heuristics, such as confidence intervals, or Bayesian inference, are difficult or impossible to apply.[^1]
 
@@
 Participants were scientists or trainees recruited via social media.
-Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy, where the true effect size was a randomly-assigned group difference between 0 and 1 standard deviations ($\delta$).
+Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy[^demo].
 Because the study was run around the Christmas holiday season, it was hoped that the theme would make the task more enjoyable.
@@
 
-```{r interface, fig.cap="The interface on which samples were reported.
+[^demo]: A demonstration version of the task that does not store data can be found at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html.
+
+```{r interface, fig.cap="Examples of the experimental interface with several participants' samples.
 The $x$-axis monotonically (but nonlinearly) related to the strength of the statistical evidence ($z$ statistic) favoring one group; the $y$-axis is monotonically (but nonlinearly) related to the sample size.
-Underlying numerical values of the statistical evidence and sample sizes were unknown to the participant.", out.width=".9\\columnwidth"}
-fn = system.file("pkg_html/img/clean_interface.png", package = "MoreyHoekstra2019")
-knitr::include_graphics(fn)
+Underlying numerical values of the statistical evidence and sample sizes were unknown to the participant.
+Corresponding $p$ values and vertical lines are given for reference; they were not shown to the participants.",fig.subcap=c("Random shuffle reports by one participant in the \"wide\" condition.","Random shuffle reports by one participant in the \"narrow\" condition.","Experimental samples by the same participant as shown in (B).
+This participant responded that the two groups were the same; the true effect size was -0.1, so their response was a false negative."), fig.height=10,fig.ncol = 1, out.width=".8\\columnwidth"}
+
+id = "R_11XYfV1Qseg1Sj0"
+fn1 = MoreyHoekstra2019::fig_recreate_display(id, "null", letter = "")
+
+id = "R_1Cm6UPEmvOHihpr"
+fn2 = MoreyHoekstra2019::fig_recreate_display(id, "null", letter = "")
+fn3 = MoreyHoekstra2019::fig_recreate_display(id, "expt", letter = "")
+
+knitr::include_graphics(c(fn1, fn2, fn3))
 
@@
 
-Participants were also offered the opportunity to sample "random shuffle reports" that were described as the results of experiments with random assignment of elves to groups: that is, the result of experiments in which the null hypothesis was true.
+Participants were randomly assigned to one of 15 effect size conditions: either no difference ($\delta=0$), or $\delta=\pm0.1$, $\pm0.185$, $\pm0.296$, $\pm0.433$, $\pm0.596$,$\pm0.785$, or $\pm1$ standard deviation units.
+Each participant had a 25% probability of being assigned $\delta=0$, with the other 75% being randomly and uniformly distributed across the remaining 14 effect size conditions.
+This true effect size was not revealed to them.
+Their goal was to determine the sign of the effect (ie, which of the two teams is truly faster).
+
+Consistent with the fictional two-sample design, statistical evidence for each "experiment" was sampled from a normal distribution with mean that depended on the (chosen, but unknown) sample size and their randomly assigned effect size:
+\[
+Z \sim \mbox{Normal}(\delta\sqrt{n/2}, 1)
+\]
+This was mapped $Z$ test statistic was mapped into a horizontal location on the interface through an arbitrary function unknown to the participant.
+Participants were randomly assigned to one of two mapping functions: a "wide" function, and a narrow function (see Supplement`~A`{=latex}, section`~1.2`{=latex} for full mathematical details).
+Figures`~`{=latex}\ref{fig:interface}A and B, show the visual effect of this manipulation.
+Statistically, these two conditions were identical; visually, they were not.
+
+This visual manipulation was crucial to study, because it allows assessment of participants' use of the null sampling distributions.
+In addition to being able to sample fictional "experiments", participants could sample "random shuffle reports" that were described as the results of experiments with random assignment of elves to groups: that is, the result of experiments in which the null hypothesis was true.
 These results took no time to return.
@@
 
-Our experiment was constructed such that the only way to assess the evidence in the data was by comparison to a null sampling distribution: either the one provided by the random shuffle reports, or a simpler null that assumes that the evidence will favor one team or the other with 50% probability.
-Thus, the information afforded only the information in a $p$ value, but it was not described as such, and participants had to *discover* for themselves how to use the information.
+Our experiment was constructed such that the only way to assess the evidence in the data was by comparison of the fictional experimental results to a null sampling distribution: either the one provided by the random shuffle reports, or a simpler null that assumes that the evidence will favor one team or the other with 50% probability.
+Thus, the information afforded only the information in a $p$ value, but it was not described as such; participants had to discover for themselves how to use the information.
 
@@
 
+# Participant sampling behavior
+
+```{r sampling, fig.cap="Sampling behavior by effect size.
+Each point represents a participant.
+A: Number of samples from the null distribution as a function of true effect size.
+B: Number of samples of fictional 'experiments' as a function of true effect size.
+Note that the $y$ axis is logarithmically scaled.
+Lines are robust regression fits.
+Positive and negative effect sizes hav been collapsed.", fig.showtext=FALSE}
+MoreyHoekstra2019::fig_sampling_behavior()
+```
+
+```{r}
+kt_null <- kruskal.test(n_null ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
+kt_expt <- kruskal.test(n_expt ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
+```
+
+Participants sought out information that would be necessary for significance tests.
+They made heavy use of shuffle reports (Figure~{=latex}\@ref(fig:sampling)A).
+Across all true effect sizes, participants sampled a median of 152 shuffle reports (range: 1-2034; in both panels A and B, lines show robust regression fits [@Venables:Ripley:2002]).
+
+Participants also made use of "replications" of the fictional experiments.
+Figure\@ref(fig:sampling)B shows the distribution of the number of experiments sampled as a function of the true effect size.
+Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`).
+When the effect size is small and difficult to detect, participants experimented more before deciding.
+
 # Success rates identifying effect sign
@@
 ```{r error-rates, fig.cap="Correct decision and error rates as a function of true effect size.
+Positive and negative effect sizes hav been collapsed.
 The red solid line represents the rate at which participants {\\em correctly} determined the direction of the effect, except for at an effect size of 0, where it represents an incorrect claim of an effect.
@@
 Of the `r null_n` participants for whom the null hypothesis was true (i.e.
-$\delta=0$), `r t1_err_count` participants (`r round(t1_err_count/null_n * 100, 0)`%) incorrectly indicated an effect.
+$\delta=0$), `r t1_err_count` participants (`r round(t1_err_count/null_n * 100, 1)`%) incorrectly indicated an effect.
 This is larger than the typically-accepted 5% false positive rate in many sciences; however, participants were performing a novel task with no recourse to numbers or statistical software.
-Those who did not indicate an effect when $\delta=0$ tended to indicate that they *did not detect* an effect (`r null_nodet_count`; `r round(null_nodet_count/null_n * 100,0)`%), which is the correct conclusion from the SST perspective.
-The other `r null_same_count` (`r round(null_same_count/null_n * 100,0)`%) indicated that the groups were the same, which under SST is a fallacy.
+Those who did not indicate an effect when $\delta=0$ tended to indicate that they *did not detect* an effect (`r null_nodet_count`; `r round(null_nodet_count/null_n * 100,1)`%), which is the correct conclusion from the SST perspective.
+The other `r null_same_count` (`r round(null_same_count/null_n * 100,1)`%) indicated that the groups were the same, which under SST is a fallacy.
 
@@
 
+Signal detection theory gives us another perspective on the error rates shown in Figure`~`{=latex}\ref{fig:error-rates}, allowing us to correct for the baseline of errors that occur in the null condition [@Macmillan:Creelman:2005].
+We combine the "false alarm" rate when $\delta=0$ (`r round(t1_err_count/null_n * 100, 1)`%) with the "hit rates" for all other conditions using a simple signal detection model; see Supplement A, section 8 for model details.
+The fitted model yields $d'$ parameters that range from 1.41 when $\delta=0.1$ to 3.24 when $\delta=1$.
+
+# Use of information in the display
+
 Another way one can evaluate the participants' responses is whether they reflect the information in the display at the time the decision is made, taking into account all points.
-To roughly quantify the evidence for a difference for each participant, we computed $p$ values from the Wilcoxon rank-sum test between the shuffle reports and the fictitious experimental results as they stood when the participant made their decision.
+To roughly quantify the evidence for a difference for each participant, we computed two $p$ values from Wilcoxon tests using the fictitious experimental results as they stood when the participant made their decision: a signed-rank test on the experimental samples alone, and a rank-sum test between the shuffle reports and the experimental samples.
+These two $p$ values indicate the information available to participants using sign-like significance tests and those using the null samples, respectively.
+The rank-sum $p$ value was typically lower, but to fairly account for the information available to the participant, we took the lower of the two.
 Small Wilcoxon $p$ values suggest a difference between the shuffle reports and the experiments.
@@
 ```{r response-ev, fig.cap="Statistical evidence underlying participants' decisions.
-The Wilcoxon rank-sum test $p$ value ($x$ axis) between the participants' null samples and experimental samples is used as a rough index of evidential strength.
+The Wilcoxon $p$ value ($x$ axis) used as a rough index of evidential strength in the display.
 Kernel density estimates for the evidence are shown for three relevant conclusions.
@@
 Filled circles show correct decisions; hollow circles, incorrect decisions.
-The two hollow squares show sign errors.", out.width=".9\\columnwidth"}
+The two asterisks show sign errors.", out.width=".9\\columnwidth", fig.height=4.8}
 MoreyHoekstra2019::fig_response_evidence()
@@
 Kernel density estimates show the distributions of $p$ values when participants indicated that Sparklies were faster, no detection/same, or that Jinglies were faster.
-With few exceptions, participants' conclusions appear reasonable given the information in the display, though a few participants appear to ignore clear evidence of an effect.
+With a few notable exceptions, participants' conclusions appear reasonable given the information in the display, though a few participants appear to ignore clear evidence of an effect.
+
+# Sensitivity to SST-Relevant Information
+
+```{r response-ev2, fig.cap="The effect of the evidence transformation manipulation on responding.
+Points on top (narrow scale; $q=7$) and bottom (wide scale; $q=3$) represent participants' decisions as a function of the most extreme experiment sampled.
+See the methods details for the interpretation of $q$.
+``N'' indicates a ``no detect'' or ``same'' response; ``J/S'' indicates a response in favor of a difference between the groups.
+Curves show predicted probability by a logistic regression fit with standard errors.", fig.show = "hold", fig.subcap = c("Predicted response probabilities relative to visual extremity.
+Vertical lines show the critical 0.05 for the corresponding null sampling distribution.","Predicted response probabilities relative to the null sampling distributions (implicit $p$ values)."), out.width=".9\\columnwidth", fig.ncol = 1}
+glm_obj_x <- MoreyHoekstra2019::fig_evidence_scale_logistic_x()
+
+pval = .05
+
+glm_pars_x = broom::tidy(glm_obj_x)
+glm_anova_x = broom::tidy(anova(glm_obj_x, test = "Chisq"))
+glm_apa_x = papaja::apa_print(glm_obj_x)
+
+pred_p = outer(c(3,7), c(3,7), function(true, z_calc){
+  plogis(predict(glm_obj_x, 
+              newdata = list(ev = MoreyHoekstra2019:::evz(qnorm(1 - pval / 2), z_calc), 
+                 evidence_power = true)
+              )
+  )
+})
+dimnames(pred_p) <- list("true"=c("Wide (3)","Narrow (7)"),
+                         "z_calc"=c("Wide (3)","Narrow (7)"))
+
+glm_obj_p <- MoreyHoekstra2019::fig_evidence_scale_logistic_p()
+
+glm_pars_p = broom::tidy(glm_obj_p)
+glm_anova_p = broom::tidy(anova(glm_obj_p, test = "Chisq"))
+glm_apa_p = papaja::apa_print(glm_obj_p)
+
+```
+
+In addition to a random effect size, participants were randomly assigned to one of two transformations of the location/color test statistic from an underlying $z$ statistic.
+Of particular interest was how the transformation affected responding for the same visual deviation from the center.
+
+The visual effects of the manipulation are shown in Figure \ref{fig:interface}, panels A and B.
+The two experimental conditions used different arbitrary monotone mappings from the underlying $Z$-statistic to the visual space.
+Intuitively, this would be like deciding to use $Z^3$ instead of $Z$ in all $Z$ tests; one would need to adjust the significance criteria to account for the cubing (e.g., use $|1.96^3|=7.53$ instead of $|1.96|$ for a $\alpha=0.05$ level test), but the underlying test remains the same.
+The manipulation changes only the visual impression of the sampling distributions, allowing us to see how sensitive their responses are to the null sampling distribution as represented by the random shuffle reports.
+
+If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider.
+When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant, given the $p$ value.
+
+Figure \ref{fig:response-ev2} (top) shows responses (no detect/same or Jinglies/Sparklies) as a function of the most extreme experiment sampled ($x$ axis) and the transformation.
+There was a strong effect of the transformation consistent with use of the null sampling distribution; participants randomly assigned to the "narrow" evidence transformation responded "Jinglies/Sparklies" for much less visually extreme evidence (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_x[3:4,3]),3)`, p<.001$).
+
+A logistic regression relating responses to the visual extremeness of the evidence and the transformation provides predicted probabilities of responding "Jinglies/Sparklies" when the visual evidence corresponded to $p=0.05$ for the null sampling distribution.
+In both the wide and the narrow conditions, the predicted probability of a "Jinglies/Sparklies" response at the critical value was about 22%, despite that in the wide transformation condition this point was about twice as visually extreme.
+
+Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response-ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).
+It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with some participants incorrectly using the non-diagnostic visual extremeness to perform the task.
+If more people had been fooled by the irrelevant visual manipulation of the sampling distribution, we would expect this effect to be substantially larger.
 
@@
 Secondly---and most importantly---the instructions did not tell the participants *how* they should use the shuffles reports, yet many participants gave detailed accounts.
-As we show next, participants' decisions were sensitive to a key manipulation of the shuffles reports in the manner that one would expect if they were using SST logic.
-
-# Sensitivity to SST-Relevant Information
-
-```{r response-pvals, fig.cap="Selected two-tailed $p$ values from the null sampling distributions of the narrow (top) and wide (bottom) conditions, projected onto the experimental interface.
-Participants were not shown the $p$ values; they had to intuit them by using samples from the null distribution (``random shuffle reports'').", fig.show = "hold", fig.subcap = c("Implied $p$ values for the ``narrow'' condition.","Implied $p$ values for the ``wide'' condition."), out.width=".9\\columnwidth", fig.ncol = 1}
-fn = MoreyHoekstra2019::fig_evidence_p_vals(q = 7, letter = "")
-knitr::include_graphics(fn)
-fn = MoreyHoekstra2019::fig_evidence_p_vals(q = 3, letter = "")
-knitr::include_graphics(fn)
-```
-
-```{r response-ev2, fig.cap="The effect of the evidence transformation manipulation on responding.
-Points on top (narrow scale; $q=7$) and bottom (wide scale; $q=3$) represent participants' decisions as a function of the most extreme experiment sampled.
-See the methods details for the interpretation of $q$.
-``N'' indicates a ``no detect'' or ``same'' response; ``J/S'' indicates a response in favor of a difference between the groups.
-Curves show predicted probability by a logistic regression fit with standard errors.", fig.show = "hold", fig.subcap = c("Predicted response probabilities relative to visual extremity.
-Vertical lines show the critical 0.05 for the corresponding null sampling distribution.","Predicted response probabilities relative to the null sampling distributions (implicit $p$ values)."), out.width=".9\\columnwidth", fig.ncol = 1}
-glm_obj_x <- MoreyHoekstra2019::fig_evidence_scale_logistic_x()
-
-pval = .05
-
-glm_pars_x = broom::tidy(glm_obj_x)
-glm_anova_x = broom::tidy(anova(glm_obj_x, test = "Chisq"))
-glm_apa_x = papaja::apa_print(glm_obj_x)
-
-pred_p = outer(c(3,7), c(3,7), function(true, z_calc){
-  plogis(predict(glm_obj_x, 
-              newdata = list(ev = MoreyHoekstra2019:::evz(qnorm(1 - pval / 2), z_calc), 
-                 evidence_power = true)
-              )
-  )
-})
-dimnames(pred_p) <- list("true"=c("Wide (3)","Narrow (7)"),
-                         "z_calc"=c("Wide (3)","Narrow (7)"))
-
-glm_obj_p <- MoreyHoekstra2019::fig_evidence_scale_logistic_p()
-
-glm_pars_p = broom::tidy(glm_obj_p)
-glm_anova_p = broom::tidy(anova(glm_obj_p, test = "Chisq"))
-glm_apa_p = papaja::apa_print(glm_obj_p)
-
-```
-
-In addition to a random effect size, participants were randomly assigned to one of two transformations of the location/color test statistic from an underlying $z$ statistic.
-Of particular interest was how the transformation affected responding for the same visual deviation from the center.
-In one condition ("wide"), the null sampling distribution was visually about twice as wide as in the other ("narrow").
-
-The effects of the manipulation are shown in Figure \ref{fig:response-pvals}.
-The two experimental conditions used different arbitrary monotone mappings from the underlying $Z$-statistic to the visual space.
-Intuitively, this would be like deciding to use $Z^3$ instead of $Z$ in all $Z$ tests; one would need to adjust the significance criteria to account for the cubing (e.g., use $|1.96^3|=7.53$ instead of $|1.96|$ for a $\alpha=0.05$ level test), but the underlying test remains the same.
-The manipulation changes only the visual impression of the sampling distributions, allowing us to see how sensitive their responses are to the null sampling distribution as represented by the random shuffle reports.
-
-If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider.
-When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant, given the $p$ value.
-
-Figure \ref{fig:response-ev2} (top) shows responses (no detect/same or Jinglies/Sparklies) as a function of the most extreme experiment sampled ($x$ axis) and the transformation.
-There was a strong effect of the transformation consistent with use of the null sampling distribution; participants randomly assigned to the "narrow" evidence transformation responded "Jinglies/Sparklies" for much less visually extreme evidence (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_x[3:4,3]),3)`, p<.001$).
-
-A logistic regression relating responses to the visual extremeness of the evidence and the transformation provides predicted probabilities of responding "Jinglies/Sparklies" when the visual evidence corresponded to $p=0.05$ for the null sampling distribution.
-In both the wide and the narrow conditions, the predicted probability of a "Jinglies/Sparklies" response at the critical value was about 22%, despite that in the wide transformation condition this point was about twice as visually extreme.
-
-Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response-ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).
-It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with a small number of participants incorrectly using the non-diagnostic visual extremeness to perform the task.
-If more people had been fooled by the irrelevant visual manipulation of the sampling distribution, we would expect this effect to be substantially larger.
+Combined with the other reported results, this strongly suggests that our participants---with some exceptions---do understand the basic SST logic and can deploy it to correctly solve novel problems.
 

### 2020-09-02 inst/manuscript/supplementA.Rmd
first edits for revision

--- 2020-02-07 inst/manuscript/supplementA.Rmd
@@
 bibliography: "`r system.file('bib/bibfile.bib', package = 'MoreyHoekstra2019')`"
-nocite: | 
-  @Venables:Ripley:2002
 ---
@@
 ```
-
-# Participant sampling behavior
-
-```{r sampling, fig.cap="Sampling behavior by effect size.
-Each point represents a participant.
-A: Number of samples from the null distribution as a function of true effect size.
-B: Number of samples of fictional 'experiments' as a function of true effect size.
-Note that the $y$ axis is logarithmically scaled.
-Lines are robust regression fits (Venables \\& Ripley, 2002).", fig.showtext=FALSE}
-MoreyHoekstra2019::fig_sampling_behavior()
-```
-
-```{r}
-kt_null <- kruskal.test(n_null ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
-kt_expt <- kruskal.test(n_expt ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
-```
-
-Participants sought out information that would be necessary for significance tests.
-They made heavy use of shuffle reports (Figure A above).
-Across all true effect sizes, participants sampled a median of 152 shuffle reports (range: 1-2034).
-
-Participants also made use of "replications" of the fictional experiments.
-(Figure B above) shows the distribution of the number of experiments sampled as a function of the true effect size.
-Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`).
-When the effect size is small and difficult to detect, participants do more experimentation before deciding.
 
@@
 
+# Signal detection analysis
+
+We can treat the participants as "detectors" of the true effect, and ask how well the scientists were able to distinguish signal from noise, in the aggregate.
+To do this, we fit a signal detection model that allowed $d'$ to vary as a function of the true effect size.
+
+Let $d'_i$ ($i=-7,\ldots,7$) be the means for the 15 effect size conditions from $\delta=-1$ to $\delta=1$.
+We constrain $d'_0=0$ , $d'_i\leq d'_j$ when $i<j$ (monotonicity), and $d_{i}=-d_{-i}$ (symmetry).
+The probabilities of responding "Jinglies" or "Sparklies" in the $i$th effect size condition are
+\[
+\begin{eqnarray}
+Pr_i(\mbox{Sparklies}) &=& \Phi\left(c_1 - d'_i\right)\\
+Pr_i(\mbox{Jinglies}) &=& 1 - \Phi\left(c_2 - d'_i\right)
+\end{eqnarray}
+\]
+where $\Phi$ is the cumulative distribution function of the standard normal distribution and $c_1,c_2$ are criteria (not necessarily symmetric).
+The probability of responding "same" or "cannot detect" is the remaining probability.
+
+The model has 9 parameters (two criteria and seven means, though the monotonicity assumption adds more constraint) for 30 non-redundant data points.
+The model was fit using maximum likelihood (R code is available in the source of this document).
+
+```{r sigdet1, fig.cap="Aggregate signal detection analysis. Fitted d' for each effect size. See text for a description of the fitted model.",fig.showtext=FALSE}
+MoreyHoekstra2019::christmas_stats_participants %>%
+  mutate(
+    response = as.character(response),
+    response = case_when(
+      response %in% c("same","no_detect") ~ "no detect/same",
+      TRUE ~ response
+    )
+  ) %>%
+  group_by(effect_size, response) %>%
+  summarise(n = n()) -> freqs0
+
+freqs0 %>%
+  pivot_wider(names_from = "response", values_from = "n") %>%
+  replace_na(list("no detect/same"=0, "sparklies"=0, "jinglies"=0)) -> freqs
+
+loglike = function(pars, d, return_probs = FALSE){
+
+  mu0 = cumsum(exp(pars[1:7]))
+  mu     = c(-rev(mu0), 0, mu0)
+  crit1  = pars[8]
+  crit_d = exp(pars[9])
+  crits  = c(-Inf, crit1, crit1 + crit_d, Inf)
+
+  Y = as.matrix(freqs[,-1])[,c(2,1,3)]
+
+  pr = t(apply(outer(mu, crits, function(m, c) pnorm(c,m)),
+        1, diff))
+  if(return_probs) return(pr)
+
+  -sum(Y * log(pr))
+}
+
+pars_init = c(rep(-.5,7),-1,1)
+
+g = optim(pars_init, loglike, d=freqs)
+mu0 = cumsum(exp(g$par[1:7]))
+dp     = c(-rev(mu0), 0, mu0)
+
+tibble(dprime = dp, effect_size = freqs$effect_size) %>%
+  ggplot(aes(x = effect_size, y = dp)) +
+  annotate("rect",xmin=-1,xmax=0,ymin=-Inf,ymax=Inf,
+           fill="blue", alpha = .2, linetype=0) +
+  annotate("rect",xmin=1,xmax=0,ymin=-Inf,ymax=Inf,
+           fill="red", alpha = .2, linetype=0) +
+  annotate("text", x = -1, y = 0, angle = 90, color = "blue", label="Sparklies", vjust=1.1, size = 5,
+           family = MoreyHoekstra2019::pkg_options("ggplot_family")
+           ) +
+  annotate("text", x = 1, y = 0, angle = -90, color = "red", label="Jinglies", vjust=1.1, size = 5,
+           family = MoreyHoekstra2019::pkg_options("ggplot_family")
+           ) +
+  geom_line() +
+  geom_point(size = 2) +
+  scale_y_continuous(name = "Estimated d'", expand = c(0,0)) +
+  scale_x_continuous(name = "Effect size", expand = c(0,0)) +
+  coord_cartesian(clip="off") +
+  geom_point(x=0,y=0,size=2, shape = 21, color = "black", fill = "white") +
+  theme_minimal() +
+  theme(
+      panel.grid.minor.y = element_blank(),
+      panel.grid.major.x = element_blank(),
+      panel.grid.minor.x = element_blank(),
+      panel.background = element_blank(),
+      text = element_text(family = MoreyHoekstra2019::pkg_options("ggplot_family"))
+    )
+
+```
+
+Figure \@ref(fig:sigdet1) shows the estimated $d'$ parameters as a function of effect size. They range from `r round(dp[9],3)` when $\delta=.1$ to `r round(dp[15],3)` when $\delta=1$.
+
+```{r sigdet2, fig.cap = "Observed and fitted probabilities for each effect size and response.
+Lines show predicted probabilities; ribbons show where 68% of the observed probabilities should fall given the predicted probabilities.
+These limits are approximate due to the discreteness of the response.",fig.showtext=FALSE}
+
+loglike(g$par, freqs, return_probs = TRUE) %>%
+  `[`(,c(2,1,3)) %>%
+  data.frame() %>%
+  `colnames<-`(colnames(freqs)[-1]) %>%
+  mutate(effect_size = !!freqs$effect_size) %>%
+  pivot_longer(cols=1:3,names_to = "response", values_to = "predicted") %>%
+  left_join(freqs0, by = c("effect_size", "response")) %>%
+  complete(n, fill = list(n=0)) %>%
+  arrange(effect_size) %>%
+  group_by(effect_size) %>%
+  mutate(n_total = sum(n),
+         prob_obs = n / n_total) %>%
+  mutate(
+    pred_lo = qbinom(pnorm(-1), n_total, predicted)/n_total,
+    pred_up = qbinom(pnorm(1), n_total, predicted)/n_total
+    ) -> error_rates
+
+error_rates %>%
+  mutate(response = tools::toTitleCase(response)) %>%
+  ggplot(aes(x=effect_size,y=predicted,group=response,color=response)) +
+  geom_segment(aes(x=effect_size,xend=effect_size), y=0, yend=1, color = "gray", alpha = .5, size = .2) +
+  geom_line() +
+  geom_ribbon(aes(ymin=pred_lo,ymax=pred_up,fill=response), alpha=.3, linetype=0) +
+  scale_x_continuous(limits=c(-1,1),expand=c(0,0), name = "Effect size", breaks = round(unique(error_rates$effect_size),2)) +
+  scale_y_continuous(limits=c(0,1.25),expand=c(0,0), name = "Probability", breaks = seq(0,1,.2)) +
+  scale_color_manual(name="Response", values = c("#FF0000", "#999999", "#0000FF")) +
+  scale_fill_manual(name="Response",values = c("#FF0000", "#999999", "#0000FF")) +
+  geom_point(aes(y = prob_obs)) +
+  coord_cartesian(clip="off") +
+  annotate(
+    "text",
+    x = freqs$effect_size,
+    y = 1.06,
+    label = paste0("(", rowSums(freqs[,2:4]), ")"),
+    size = 3,
+    angle = 90,
+    family = MoreyHoekstra2019::pkg_options("ggplot_family")
+  ) +
+  annotate(
+    "text",
+    x = 0,
+    y = 1.20,
+    label = "(Number of participants)",
+    size = 3,
+    family = MoreyHoekstra2019::pkg_options("ggplot_family")
+    ) +
+    theme_minimal() +
+    theme(
+      panel.grid.minor.y = element_blank(),
+      panel.grid.major.x = element_blank(),
+      panel.grid.minor.x = element_blank(),
+      panel.background = element_blank(),
+      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+      text = element_text(family =MoreyHoekstra2019::pkg_options("ggplot_family")),
+      legend.position = "top"
+    )
+
+```
+
+ Figure \@ref(fig:sigdet2) shows the predicted and observed probabilities of all responses.
+The good fit of this model licenses the interpretation of the $d'$ parameters and the collapsing across the sign of the effect in Figure 3 of the main document.
+
 -------

### 2020-09-02 R/fig_response_evidence.R

@@ -19,30 +19,67 @@ fig_response_evidence <- function(filter_func = function(data, ...) return(data)
     filter_func()
   samples <- MoreyHoekstra2019::christmas_stats_samples %>%
     filter(id %in% dat$id)
-
-
-  w_stats = samples %>%
+  
+  
+samples %>%
     mutate(ev0 = ev * sgn) %>%
-    split(.$id) %>%
-    map( ~ wilcox.test(ev0 ~ type, alternative = "less", data = .)) %>%
-    map_dbl("p.value")
-
-  w_stats = w_stats[match(dat$id, names(w_stats))]
-
-  j = dat$response == "jinglies"
-  s = dat$response == "sparklies"
-  n = dat$response %in% c("same", "no_detect")
-
-  dns = density(qlogis(w_stats[n]), na.rm = TRUE)
-
+    split(.$id) -> data_split
+  
+  names(data_split) %>%
+    map_df(function(id){
+      decision = dat$response[dat$id==id]
+      alt = switch(as.character(decision),
+                   "jinglies" = "greater",
+                   "sparklies" = "less",
+                   "two.sided"
+      )
+      d = data_split[[id]]
+      # This is annoying seems is necessary due to numerical precision
+      # issues with high p values
+      w2_g = wilcox.test(ev0 ~ type, alternative = "greater", data = d)
+      w2_l = wilcox.test(ev0 ~ type, alternative = "less", data = d)
+      w1_g = wilcox.test(ev0 ~ 1, alternative = "greater", data = d %>% filter(type=="expt"))
+      w1_l = wilcox.test(ev0 ~ 1, alternative = "less", data = d %>% filter(type=="expt"))
+      
+      w1_p = min(w1_l$p.value, w1_g$p.value)
+      w2_p = min(w2_l$p.value, w2_g$p.value)
+      
+      w1_winner = ifelse(w1_l$p.value<w1_g$p.value, "sparklies","jinglies")
+      w2_winner = ifelse(w2_l$p.value<w2_g$p.value, "sparklies","jinglies")
+      
+      bind_cols(id = id, two=w2_p, one=w1_p,
+                two_winner = w2_winner, 
+                one_winner = w1_winner)
+    }) %>%
[…]

### 2020-09-02 inst/shiny/explore/server.R

@@ -28,18 +28,51 @@ repl_text = c(
     text_missing = "Text missing"
 )
 
-
-suppressWarnings({w_stats = samples %>%
+samples %>%
     mutate(ev0 = ev * sgn) %>%
-    split(.$id) %>%
-    map( ~ wilcox.test(ev0 ~ type, alternative = "less", data = .)) %>%
-    map_dbl("p.value")})
+    split(.$id) -> data_split
 
-w_stats <- w_stats[match(dat$id, names(w_stats))]
+suppressWarnings({
+    names(data_split) %>%
+        map_df(function(id){
+            decision = dat$response[dat$id==id]
+            alt = switch(as.character(decision),
+                         "jinglies" = "greater",
+                         "sparklies" = "less",
+                         "two.sided"
+            )
+            d = data_split[[id]]
+            # This is annoying seems is necessary due to numerical precision
+            # issues with high p values
+            w2_g = wilcox.test(ev0 ~ type, alternative = "greater", data = d)
+            w2_l = wilcox.test(ev0 ~ type, alternative = "less", data = d)
+            w1_g = wilcox.test(ev0 ~ 1, alternative = "greater", data = d %>% filter(type=="expt"))
+            w1_l = wilcox.test(ev0 ~ 1, alternative = "less", data = d %>% filter(type=="expt"))
+            
+            w1_p = min(w1_l$p.value, w1_g$p.value)
+            w2_p = min(w2_l$p.value, w2_g$p.value)
+            
+            w1_winner = ifelse(w1_l$p.value<w1_g$p.value, "sparklies","jinglies")
+            w2_winner = ifelse(w2_l$p.value<w2_g$p.value, "sparklies","jinglies")
+            
+            bind_cols(id = id, two=w2_p, one=w1_p,
+                      two_winner = w2_winner, 
+                      one_winner = w1_winner)
+        }) %>%
+        mutate(lowest = pmin(one,two),
+               lowest_winner = case_when(
+                   one<two ~ one_winner,
+                   TRUE ~ two_winner
+               ),
+               qplot = case_when(
+                   lowest_winner == "jinglies" ~ -qlogis(2*lowest),
+                   TRUE ~ qlogis(2*lowest)
+               )
+        ) %>% 
+        left_join(dat, by="id") -> w_stats
+})
 
-dat$wilcoxon = ifelse(w_stats < .5,
-                      w_stats*2,
[…]

### 2020-10-21 inst/manuscript/manuscript.Rmd
first draft of revision

@@
 However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication).
-It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results.
+It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014;@Hoekstra:etal:2012]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results.
 Conceptual understanding and fluency with common representations are both important, but are distinct.
@@
 
-Instead of the usual human-factors approach, which adopts familiar statistical language and tests participants' fluency with existing procedures, we adopt a more cognitive perspective.
-We aim to test working scientists' understanding of the basic conceptual framework underlying SST using a simulated experimental task.
 To avoid conflating basic reasoning processes and fluency with common statistical terminology, we avoid using common statistics---or, indeed, any numbers---at all.
+Instead of focusing on familiar statistical language and tests participants' fluency with existing procedures, we adopt a different approach: we test working scientists' understanding of the basic conceptual framework underlying SST using a simulated experimental task.
 
@@
 Participants were scientists or trainees recruited via social media.
-Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy[^demo].
+Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy.
+A demonstration version of the task that does not store data can be found at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html.
 Because the study was run around the Christmas holiday season, it was hoped that the theme would make the task more enjoyable.
@@
 Importantly, the meaning of the colors and locations was unknown to the participants, aside from the monotone relationship with the sample size and statistical evidence.
-
-[^demo]: A demonstration version of the task that does not store data can be found at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html.
 
@@
 Each participant had a 25% probability of being assigned $\delta=0$, with the other 75% being randomly and uniformly distributed across the remaining 14 effect size conditions.
-This true effect size was not revealed to them.
+These true effect sizes were not revealed to the participants.
 Their goal was to determine the sign of the effect (ie, which of the two teams is truly faster).
@@
 \]
-This was mapped $Z$ test statistic was mapped into a horizontal location on the interface through an arbitrary function unknown to the participant.
+The $Z$ test statistic was mapped into a horizontal location on the interface through an arbitrary function unknown to the participant.
 Participants were randomly assigned to one of two mapping functions: a "wide" function, and a narrow function (see Supplement`~A`{=latex}, section`~1.2`{=latex} for full mathematical details).
-Figures`~`{=latex}\ref{fig:interface}A and B, show the visual effect of this manipulation.
+Figures`~`{=latex}\ref{fig:interface}A and B show the visual effect of this manipulation.
 Statistically, these two conditions were identical; visually, they were not.
@@
 Participants sought out information that would be necessary for significance tests.
-They made heavy use of shuffle reports (Figure~{=latex}\@ref(fig:sampling)A).
+They made heavy use of shuffle reports (Figure`~A`{=latex}\@ref(fig:sampling)).
 Across all true effect sizes, participants sampled a median of 152 shuffle reports (range: 1-2034; in both panels A and B, lines show robust regression fits [@Venables:Ripley:2002]).
@@
 
-```{r error-rates, fig.cap="Correct decision and error rates as a function of true effect size.
-Positive and negative effect sizes hav been collapsed.
-The red solid line represents the rate at which participants {\\em correctly} determined the direction of the effect, except for at an effect size of 0, where it represents an incorrect claim of an effect.
-The shaded region around the solid line is one standard error wide.
-The difference between the black and the gray dashed lines shows the proportion of participants who decided that the two groups were the same (as opposed to failing to detect a difference).", fig.showtext=FALSE, out.width=".9\\columnwidth"}
-MoreyHoekstra2019::fig_error_rates()
+```{r error-rates, fig.cap="Observed and fitted probabilities for each effect size and response.
+For negative and positive effect sizes, the correct response \"Sparklies\" or \"Jinglies\" respectively.
+Fitted probabilities are from the signal detection model outlined in Supplement A.
+Lines show predicted probabilities; ribbons show where 68\\% of the observed probabilities should fall given the predicted probabilities.
+These limits are approximate due to the discreteness of the response.", fig.showtext=FALSE, out.width=".9\\columnwidth"}
+MoreyHoekstra2019::fig_error_rates_dprime()
 ```
 
-Decision error rates as a function of true effect size are shown in Figure \ref{fig:error-rates}.
+Decision rates as a function of true effect size are shown in Figure \ref{fig:error-rates}.
 
@@
 Those who did not indicate an effect when $\delta=0$ tended to indicate that they *did not detect* an effect (`r null_nodet_count`; `r round(null_nodet_count/null_n * 100,1)`%), which is the correct conclusion from the SST perspective.
-The other `r null_same_count` (`r round(null_same_count/null_n * 100,1)`%) indicated that the groups were the same, which under SST is a fallacy.
+The other `r null_same_count` (`r round(null_same_count/null_n * 100,1)`%) indicated that the groups were the same, which under SST is typically considered a fallacy.
 
@@
 These two $p$ values indicate the information available to participants using sign-like significance tests and those using the null samples, respectively.
-The rank-sum $p$ value was typically lower, but to fairly account for the information available to the participant, we took the lower of the two.
-Small Wilcoxon $p$ values suggest a difference between the shuffle reports and the experiments.
+The rank-sum $p$ value was typically lower; although it makes little difference to the qualitative results, to fairly account for the information available to the participant, we used the smaller of the two $p$ values.
+In general, smaller $p$ values suggest a larger observed between the shuffle reports and the experiments, allowing us to compare the stimulus the participants were given to their decisions.
 
@@
 Moreover, they report strategies consistent with SST, and the signature of SST reasoning can be seen in their responses.
-Because we removed numerical effect size and sample size information --- making strategies other than pure significance testing difficult or impossible to apply --- our results are strong evidence that scientists *can* successfully deploy SST logic.
+Because we removed numerical effect size and sample size information --- making strategies other than pure significance testing difficult or impossible to apply --- our results are evidence that scientists *can* successfully deploy SST logic.
 It is still an open question what causes SST to be misunderstood so often, but it seems comprehension of its underlying logic can be ruled out for many scientists.
 
-There are some major differences between our methods and the methods of previous demonstrations that scientists fail when reasoning about SST.
+Our findings echo other work showing that human reasoning can, under some conditions, be better than previously understood.
+[@Cosmides:Tooby:1992;@Gigerenzer:Hoffrage:1995].
+Suggestions that SST be discontinued due to scientists' apparent misunderstandings may have been hasty.
+Of course, there may be other reasons to abandon SST, but our work shows that given the opportunity, scientists successfully deploy basic SST logic.
+In spite of scientists' real-life statistical behaviour often resembling a "ritual" [@Gigerenzer:etal:2004], when we eliminate the ritual --- no $p$ value, or any other familiar number, was offered --- they think statistically, very often arriving at the correct conclusion about the sign of the effect.
+
+We wish to emphasize what we cannot, and do not, argue.
+First, we cannot argue that simply because scientists can successfully use SST logic, that they do in real situations, or that specific instantiations of SST, such as $p$ values, are used well.
+We specifically set out to abstract the logic away from the typical ways and situations in which scientists use the logic.
+This has the benefit of being helping to identify where problems might be, but the downside that generalizing the results will require further work.
+We are also not addressing other potential arguments against SST, such as philosophical ones.
+
+Although our conclusions are different those of other studies, there are some major differences between our methods and the methods of previous demonstrations that scientists fail when reasoning about SST.
 We have already mentioned the fact that previous demonstrations typically use survey vignettes, as opposed to our more engaged experimental method.
@@
 
-Our findings echo other work showing that human reasoning can, under some conditions, be better than previously understood.
-[@Cosmides:Tooby:1992;@Gigerenzer:Hoffrage:1995].
-Suggestions that SST be discontinued due to scientists' apparent misunderstandings may have been hasty.
-Of course, there may be other reasons why some favor abandoning SST, but our work shows that given the opportunity, scientists successfully deploy basic SST logic.
-In spite of scientists' real-life statistical behaviour often resembling a "ritual" [@Gigerenzer:etal:2004], when we eliminate the ritual --- no $p$ value, or any other familiar number, was offered --- they think statistically, very often arriving at the correct conclusion about the sign of the effect.
-Effective methodological reform in the sciences may be easier than abandoning significance testing, if we can harness scientists' already-existing statistical competence.
+Finally, we hope to provide a fresh method and perspective on a long-standing debate in statistical cognition.
+Simulation-based approaches to teaching statistics have long been touted [@Cumming:etal:1995;@Rossman:Chance:2014].
+Simulation-based approaches to *studying* scientists' statistical reasoning may also profitable, particularly in studying reasoning that is difficult for participants to articulate formally.
+If we are to reform statistical education and practice in the sciences, we should base that reform on diverse lines of evidence about scientists' reasoning.
+Harnessing scientists' already-existing statistical competence may make effective methodological reform in the sciences easier than wholesale abandoning of SST.
 

### 2020-10-21 inst/manuscript/supplementA.Rmd
first draft of revision

--- 2020-05-31 inst/manuscript/supplementA.Rmd
@@
 ```
-
-# Participant sampling behavior
-
-```{r sampling, fig.cap="Sampling behavior by effect size.
-Each point represents a participant.
-A: Number of samples from the null distribution as a function of true effect size.
-B: Number of samples of fictional 'experiments' as a function of true effect size.
-Note that the $y$ axis is logarithmically scaled.
-Lines are robust regression fits (Venables \\& Ripley, 2002).", fig.showtext=FALSE}
-MoreyHoekstra2019::fig_sampling_behavior()
-```
-
-```{r}
-kt_null <- kruskal.test(n_null ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
-kt_expt <- kruskal.test(n_expt ~ factor(abs(effect_size)), data = MoreyHoekstra2019::christmas_stats_participants)
-```
-
-Participants sought out information that would be necessary for significance tests.
-They made heavy use of shuffle reports (Figure A above).
-Across all true effect sizes, participants sampled a median of 152 shuffle reports (range: 1-2034).
-
-Participants also made use of "replications" of the fictional experiments.
-(Figure B above) shows the distribution of the number of experiments sampled as a function of the true effect size.
-Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print.htest(kt_expt)$statistic`).
-When the effect size is small and difficult to detect, participants do more experimentation before deciding.
 
@@
 
+# Signal detection analysis
+
+We can treat the participants as "detectors" of the true effect, and ask how well the scientists were able to distinguish signal from noise, in the aggregate.
+To do this, we fit a signal detection model that allowed $d'$ to vary as a function of the true effect size.
+
+Let $d'_i$ ($i=-7,\ldots,7$) be the means for the 15 effect size conditions from $\delta=-1$ to $\delta=1$.
+We constrain $d'_0=0$ , $d'_i\leq d'_j$ when $i<j$ (monotonicity), and $d_{i}=-d_{-i}$ (symmetry).
+The probabilities of responding "Jinglies" or "Sparklies" in the $i$th effect size condition are
+\[
+\begin{eqnarray}
+Pr_i(\mbox{Sparklies}) &=& \Phi\left(c_1 - d'_i\right)\\
+Pr_i(\mbox{Jinglies}) &=& 1 - \Phi\left(c_2 - d'_i\right)
+\end{eqnarray}
+\]
+where $\Phi$ is the cumulative distribution function of the standard normal distribution and $c_1,c_2$ are criteria (not necessarily symmetric).
+The probability of responding "same" or "cannot detect" is the remaining probability.
+
+The model has 9 parameters (two criteria and seven means, though the monotonicity assumption adds more constraint) for 30 non-redundant data points.
+The model was fit using maximum likelihood (R code is available in the source of this document).
+
+```{r sigdet1, fig.cap="Aggregate signal detection analysis. Fitted d' for each effect size. See text for a description of the fitted model.",fig.showtext=FALSE}
+
+MoreyHoekstra2019::christmas_stats_participants %>%
+  mutate(
+    response = as.character(response),
+    response = case_when(
+      response %in% c("same","no_detect") ~ "no detect/same",
+      TRUE ~ response
+    )
+  ) %>%
+  group_by(effect_size, response) %>%
+  summarise(n = n()) %>%
+  pivot_wider(names_from = "response", values_from = "n") %>%
+  replace_na(list("no detect/same"=0, "sparklies"=0, "jinglies"=0)) %>%
+  MoreyHoekstra2019:::sigdet_model_fit() -> sd_results
+
+dp = sd_results$dp
+
+MoreyHoekstra2019::fig_dprime_effect_size()
+
+```
+
+Figure \@ref(fig:sigdet1) shows the estimated $d'$ parameters as a function of effect size. They range from `r round(dp[9],3)` when $\delta=.1$ to `r round(dp[15],3)` when $\delta=1$.
+
+```{r sigdet2, fig.cap = "Observed and fitted probabilities for each effect size and response.
+Lines show predicted probabilities; ribbons show where 68% of the observed probabilities should fall given the predicted probabilities.
+These limits are approximate due to the discreteness of the response.",fig.showtext=FALSE}
+
+MoreyHoekstra2019::fig_error_rates_dprime()
+```
+
+ Figure \@ref(fig:sigdet2) shows the predicted and observed probabilities of all responses.
+The good fit of this model licenses the interpretation of the $d'$ parameters and the collapsing across the sign of the effect in Figure 3 of the main document.
+
 -------

### 2020-10-21 R/fig_error_rates.R

@@ -176,3 +176,236 @@ fig_error_rates <- function(
   invisible(g)
 
 }
+
+#' Error/correct response rates as a function of effect size, with model fits
+#'
+#' @param filter_func function to filter the data before plotting
+#' @param print_plot (logical) set to FALSE to prevent printing the ggplot object
+#' @return Invisibly returns the ggplot object
+#' @import dplyr
+#' @import ggplot2
+#' @importFrom tidyr complete pivot_longer
+#' @export
+#'
+#' @examples
+#' suppressWarnings(
+#'   fig_error_rates_dprime()
+#' )
+fig_error_rates_dprime <- function(
+  filter_func = function(data, ...) return(data),
+  print_plot = TRUE
+) {
+  
+  MoreyHoekstra2019::christmas_stats_participants %>%
+    filter_func() %>%
+    mutate(
+      response = as.character(response),
+      response = case_when(
+        response %in% c("same","no_detect") ~ "no detect/same",
+        TRUE ~ response
+      )
+    ) %>%
+    group_by(effect_size, response) %>%
+    summarise(n = n()) -> freqs0
+  
+  freqs0 %>%
+    pivot_wider(names_from = "response", values_from = "n") %>%
+    replace_na(list("no detect/same"=0, "sparklies"=0, "jinglies"=0)) -> freqs
+  
+  
+  sd_results = sigdet_model_fit(freqs)
+  
+  mu0 = sd_results$mu0
+  dp  = sd_results$dp
+  probs  = sd_results$probs
+  
+  probs %>%
+    `[`(,c(2,1,3)) %>%
+    data.frame() %>%
+    `colnames<-`(colnames(freqs)[-1]) %>%
+    mutate(effect_size = !!freqs$effect_size) %>%
+    pivot_longer(cols=1:3,names_to = "response", values_to = "predicted") %>%
+    left_join(freqs0, by = c("effect_size", "response")) %>%
+    complete(n, fill = list(n=0)) %>%
+    arrange(effect_size) %>%
+    group_by(effect_size) %>%
+    mutate(n_total = sum(n),
+           prob_obs = n / n_total) %>%
+    mutate(
+      pred_lo = qbinom(pnorm(-1), n_total, predicted)/n_total,
+      pred_up = qbinom(pnorm(1), n_total, predicted)/n_total
+    ) -> error_rates
+  
+  error_rates %>%
[…]

### 2020-11-04 inst/manuscript/manuscript.Rmd
Revision 1, RSOS (package)

@@
 However, if scientists' understanding of significance testing is truly as poor as thought, one could argue such drastic action is required.
-We show using a novel experimental method that scientists do, in fact, understand the logic of significance testing and can use it effectively.
-This suggests that scientists may not be as statistically-challenged as often believed, and that reforms should take this into account.
+We present evidence using a novel reasoning task that scientists may understand the logic of significance testing better than previously thought.
+Scientists may not be as statistically-challenged as often believed; reforms should take this into account.
 
@@
 
-The interpretation of studies of researchers' understanding of SST are limited, however, by their methodology.
+The interpretation of studies of researchers' understanding of SST is limited, however, by their methodology.
 A typical study presents a vignette describing research results.
-Statistical results are offered to the participants (e.g., a $t$ statistic and $p$ value), who are then asked to explicitly give or endorse various interpretations that are thought to represent participants' understanding of SST.
+Statistical results are offered to the participants (e.g., a $t$ statistic and $p$ value), who are then asked to explicitly give or endorse various interpretations.
+These responses are taken to represent participants' understanding, or misunderstanding, of SST.
 However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication).
-It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014;@Hoekstra:etal:2012]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results.
+It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014;@Hoekstra:etal:2012]) tap understanding of the core logic of SST rather than, say, familiarity with the technical terminology used to present statistical results.
 Conceptual understanding and fluency with common representations are both important, but are distinct.
@@
 These examples show that whatever process lead to the statistical conclusion was flawed in some way, but many processes contribute to such conclusions.
-Cognitive (cites), technological (cites), and social processes (cites) have all been assigned some blame for statistical reasoning failures.
+Cognitive [e.g., @Pashler:Harris:2012], technological [e.g., @Kennedy-Shaffer:2019], and social processes [e.g., @Lilienfeld:2017] have all been assigned some blame for statistical reasoning failures.
 
-In deciding how to improve statistical reasoning, it is crucial to know where the problems are.
+In deciding how to improve statistical reasoning, it is crucial to know where the problems lie.
 The *fact* of reasoning problems tells us little about their *source*.
@@
 
-To avoid conflating basic reasoning processes and fluency with common statistical terminology, we avoid using common statistics---or, indeed, any numbers---at all.
+To avoid conflating basic reasoning failures and lack of fluency with common statistical terminology, we avoid using common statistics---or, indeed, any numbers---at all.
 Instead of focusing on familiar statistical language and tests participants' fluency with existing procedures, we adopt a different approach: we test working scientists' understanding of the basic conceptual framework underlying SST using a simulated experimental task.
@@
 The key innovation allowing us to focus on SST reasoning was to design an experiment that prevents the use of alternative strategies.
-A key feature of SST is that the use of a null distribution destroys information about effect sizes and sample size.
-In fact, this aspect of SST reasoning is often criticized.
-Alternative methodologies often focus on effect sizes (point estimates, confidence intervals, equivalence, likelihood, Bayesian priors/posteriors).
+A critical feature of SST is that the use of a null distribution destroys information about effect sizes and sample size.
+In fact, this aspect of SST reasoning is often criticized, while alternative methodologies focus on effect sizes (point estimates, confidence intervals, equivalence, likelihood, Bayesian priors/posteriors).
 We offered our participants only the information in a $p$ value, and participants had to understand or discover how to obtain that information.
@@
 If participants have poor understanding of SST, they would 1) often come to the wrong conclusion, in spite of ample information; 2) show error rates that are only weakly associated with true effect size; 3) be unable to articulate strategies for performing our task; 4) be sensitive to misleading, task-irrelevant information; 5) be insensitive to SST-relevant information.
-The scientists in our sample showed often came to the right conclusion, with their performance showing that they were sensitive to the information they were given.
-They explicitly reported using SST strategies.
+The scientists in our sample often came to the right conclusion, and their performance showing sensitivity to the SST-relevant information they were given.
+Moreover, they explicitly reported using SST strategies.
 Our results suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.
@@
 
-In tests of perception---e.g., of colorblindness [@Ishihara:1972]---it is common to eliminate one cue (e.g., brightness) in order assess the ability to use another (e.g., color).
-If color is the only useful cue for reading a number, deficits in color vision make the number difficult to read.
-We adopt a similar strategy to test statistical reasoning: we eliminate numerical information from statistical results to test scientists' ability to interpret results with reference to a null sampling distribution (i.e., SST logic).
+In tests of perception, it is common to eliminate one cue in order assess the ability to use another: e.g., eliminating brightness cues to test colorblindness [@Ishihara:1972].
+If color is the only useful cue for reading a number on a card, deficits in color vision make the number difficult to read.
+We adopt a similar strategy to test statistical reasoning: we eliminate numerical information from statistical results to test scientists' ability to interpret results with reference to a null sampling distribution, a central element of SST logic.
 Without numerical information, many other strategies and heuristics, such as confidence intervals, or Bayesian inference, are difficult or impossible to apply.[^1]
@@
 Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy.
-A demonstration version of the task that does not store data can be found at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html.
-Because the study was run around the Christmas holiday season, it was hoped that the theme would make the task more enjoyable.
+A demonstration version of the task can be found at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html.
+Because the study was run around the Christmas holiday season, we hoped the theme would make the task more engaging.
 The numerical information for an experiment, including sample size and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}).
@@
 These true effect sizes were not revealed to the participants.
-Their goal was to determine the sign of the effect (ie, which of the two teams is truly faster).
+Their goal was to determine the sign of the effect (i.e., which of the two teams is truly faster).
 
@@
 
+```{r}
+MoreyHoekstra2019::christmas_stats_participants %>%
+  group_by(shuffle) %>%
+  summarise(n=n()) %>%
+  mutate(p = n / sum(n),
+         yes = grepl(pattern = "yes", shuffle)) -> shuffle_understanding
+
+shuffle_understanding %>% filter(yes) %>%
+  pull(n) -> shuffle_ns
+
+shuffle_understanding %>% filter(yes) %>%
+  pull(p) -> shuffle_ps
+
+shuffle_pct = round(shuffle_ps * 100,1)
+
+```
+
 This visual manipulation was crucial to study, because it allows assessment of participants' use of the null sampling distributions.
@@
 
-After sampling as many "experiments" and "random shuffle reports" as they liked, participants could report whether they believed Jinglies or Sparklies were better, that they could not detect a difference, that there was no difference, or that they were bored and wanted to stop.
-Following their decision they were asked several open-ended questions about their strategy.
-Our key questions are whether participants can effectively find the "truth", whether they report strategies consistent with SST, and whether their behaviour shows evidence of SST use.
+After sampling as many "experiments" and "random shuffle reports" as they liked, participants could report whether they believed Jinglies or Sparklies were the better team, that they could not detect a difference, that there was no difference, or that they were bored and wanted to stop.
+Following their decision they were asked several open-ended questions about their strategy, along with some opinion and demographic questions.
+Our central questions are whether participants can effectively find the "truth", whether they report strategies consistent with SST, and whether their behaviour shows evidence of strategic SST use.
 
@@
 Here, we report the results of `r total_n` scientists or trainees who completed the statistical reasoning task.
-We provide an interactive app for exploring participants' responses at https://richarddmorey.shinyapps.io/explore/.
 
@@
 Participants also made use of "replications" of the fictional experiments.
-Figure\@ref(fig:sampling)B shows the distribution of the number of experiments sampled as a function of the true effect size.
+Figure`~`{=latex}\@ref(fig:sampling)B shows the distribution of the number of experiments sampled as a function of the true effect size.
 Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`).
@@
 
-When there was a true effect ($|\delta|>0$), correct decisions increased as a function of effect size, plateauing at about 95%.
-Of the `r total_n - null_n` participants for whom $|\delta|>0$, only `r sign_err_count` (`r round(sign_err_count/(total_n - null_n)*100,1)`%) indicated the incorrect team [a sign, or Type S, error; @Gelman:Carlin:2014].
+When there was a true effect ($\delta\neq0$), correct decisions increased as a function of effect size, plateauing at about 95%.
+Of the `r total_n - null_n` participants for whom $\delta\neq0$, only `r sign_err_count` (`r round(sign_err_count/(total_n - null_n)*100,1)`%) indicated the incorrect team [a sign, or Type S, error; @Gelman:Carlin:2014].
 For larger effect sizes, participants never incorrectly indicated that the two groups were the same.
 
-Signal detection theory gives us another perspective on the error rates shown in Figure`~`{=latex}\ref{fig:error-rates}, allowing us to correct for the baseline of errors that occur in the null condition [@Macmillan:Creelman:2005].
+Signal detection theory gives us another perspective on the decision rate shown in Figure`~`{=latex}\ref{fig:error-rates}, allowing us to correct for the baseline of errors that occur in the null condition [@Macmillan:Creelman:2005].
 We combine the "false alarm" rate when $\delta=0$ (`r round(t1_err_count/null_n * 100, 1)`%) with the "hit rates" for all other conditions using a simple signal detection model; see Supplement A, section 8 for model details.
@@
 
-Another way one can evaluate the participants' responses is whether they reflect the information in the display at the time the decision is made, taking into account all points.
+Another way of evaluating participants' responses is whether they reflect the information in the display at the time the decision is made, taking into account all points.
 To roughly quantify the evidence for a difference for each participant, we computed two $p$ values from Wilcoxon tests using the fictitious experimental results as they stood when the participant made their decision: a signed-rank test on the experimental samples alone, and a rank-sum test between the shuffle reports and the experimental samples.
 These two $p$ values indicate the information available to participants using sign-like significance tests and those using the null samples, respectively.
-The rank-sum $p$ value was typically lower; although it makes little difference to the qualitative results, to fairly account for the information available to the participant, we used the smaller of the two $p$ values.
+The rank-sum $p$ value is based on more information and so was typically lower.
+It makes little difference to the qualitative results, but to fairly account for the information available to the participant, we used the smaller of the two $p$ values.
 In general, smaller $p$ values suggest a larger observed between the shuffle reports and the experiments, allowing us to compare the stimulus the participants were given to their decisions.
@@
 With a few notable exceptions, participants' conclusions appear reasonable given the information in the display, though a few participants appear to ignore clear evidence of an effect.
+We provide an interactive app for exploring participants' individual responses at https://richarddmorey.shinyapps.io/explore/.
 
@@
 
-In addition to a random effect size, participants were randomly assigned to one of two transformations of the location/color test statistic from an underlying $z$ statistic.
+In addition to a random effect size, participants were also randomly assigned to one of two transformations of the location/color test statistic from an underlying $z$ statistic.
 Of particular interest was how the transformation affected responding for the same visual deviation from the center.
@@
 If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider.
-When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant, given the $p$ value.
+When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant given the $p$ value.
 
@@
 Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response-ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$).
-It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with some participants incorrectly using the non-diagnostic visual extremeness to perform the task.
-If more people had been fooled by the irrelevant visual manipulation of the sampling distribution, we would expect this effect to be substantially larger.
+It is noteworthy that when the responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with some participants incorrectly using the non-diagnostic visual extremeness to perform the task.
+If more people had been fooled by the irrelevant width of the null sampling distribution, we would expect this effect to be substantially larger.
 
@@
 
-We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategies), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports.
+We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategies), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports (see Supplement`~`{=latex}A, section 7 for coding details).
 
@@
 
-As Table \ref{tab:strategies} shows that a large majority of participants (`r for_table[1,1]`, `r round( 100*as.numeric(for_table[1,1]) / as.numeric(for_table[1,6]), 2)`%) indicated using strong significance testing strategies.
+As Table \ref{tab:strategies} shows, a large majority of participants (`r for_table[1,1]`, `r round( 100*as.numeric(for_table[1,1]) / as.numeric(for_table[1,6]), 2)`%) indicated using strong significance testing strategies.
 We should be cautious in directly interpreting this high number alone, however, because participants were told in the instructions that the shuffle reports could be used for assessing sampling variability.
-We did this to make clear what the shuffle reports were, but without explaining how to use them.
+We did this to make clear what the shuffle reports were, but without explaining *how* to use them.
 To some extent, then, the text responses may reflect the instructions.
@@
 
-Secondly---and most importantly---the instructions did not tell the participants *how* they should use the shuffles reports, yet many participants gave detailed accounts.
+Secondly---and most importantly---the instructions did not tell the participants how they should use the shuffles reports, yet many participants gave detailed accounts.
 Combined with the other reported results, this strongly suggests that our participants---with some exceptions---do understand the basic SST logic and can deploy it to correctly solve novel problems.
@@
 
-Although it has previously been suggested that scientists have dramatic misunderstandings of SST logic, scientists and trainees in our experiment demonstrate both understanding and the ability to use it to come to the correct conclusion in a simulated statistical task.
-Moreover, they report strategies consistent with SST, and the signature of SST reasoning can be seen in their responses.
+Although it has previously been suggested that scientists have dramatic misunderstandings of SST logic, scientists and trainees in our experiment demonstrate both understanding and the ability to use the logic to come to the correct conclusion in a simulated statistical task.
+Moreover, they report strategies consistent with SST, and signatures of SST reasoning can be seen in their responses.
 Because we removed numerical effect size and sample size information --- making strategies other than pure significance testing difficult or impossible to apply --- our results are evidence that scientists *can* successfully deploy SST logic.
-It is still an open question what causes SST to be misunderstood so often, but it seems comprehension of its underlying logic can be ruled out for many scientists.
+It is still an open question what causes typical SST reports to be misunderstood so often, but we have not found evidence that the problem is misapprehension of its underlying logic.
 
-Our findings echo other work showing that human reasoning can, under some conditions, be better than previously understood.
+Our findings echo other demonstrations that human reasoning can, under some conditions, be better than previously understood.
 [@Cosmides:Tooby:1992;@Gigerenzer:Hoffrage:1995].
-Suggestions that SST be discontinued due to scientists' apparent misunderstandings may have been hasty.
+Suggestions that SST be discontinued due to scientists' apparent misunderstandings may be hasty.
 Of course, there may be other reasons to abandon SST, but our work shows that given the opportunity, scientists successfully deploy basic SST logic.
@@
 We wish to emphasize what we cannot, and do not, argue.
-First, we cannot argue that simply because scientists can successfully use SST logic, that they do in real situations, or that specific instantiations of SST, such as $p$ values, are used well.
-We specifically set out to abstract the logic away from the typical ways and situations in which scientists use the logic.
+First, we cannot argue that simply because scientists can successfully use SST logic, that they *do* in real situations, or that specific instantiations of SST, such as $p$ values, are used well.
+We specifically set out to abstract the logic away from the typical situations in which scientists use the logic.
 This has the benefit of being helping to identify where problems might be, but the downside that generalizing the results will require further work.
-We are also not addressing other potential arguments against SST, such as philosophical ones.
-
-Although our conclusions are different those of other studies, there are some major differences between our methods and the methods of previous demonstrations that scientists fail when reasoning about SST.
-We have already mentioned the fact that previous demonstrations typically use survey vignettes, as opposed to our more engaged experimental method.
-A second difference is our sample which is a substantially larger and more diverse convenience sample than previous demonstrations.
-It is possible that the recent so-called "replication crisis" [@Baker:2016] has raised awareness among scientists of deficits of statistical reasoning.
-If this increase in awareness accounts for our more optimistic findings, this would be good news for statistical educators.
+We also cannot address other potential arguments against SST, such as philosophical ones.
 
@@
 If we are to reform statistical education and practice in the sciences, we should base that reform on diverse lines of evidence about scientists' reasoning.
-Harnessing scientists' already-existing statistical competence may make effective methodological reform in the sciences easier than wholesale abandoning of SST.
+Understanding and harnessing scientists' already-existing competence in statistical reasoning is essential to developing effective methodological reforms.
 

### 2022-11-30 inst/manuscript/supplementA.Rmd
fixed vignette issues that kept package from installing

@@
     tally() %>%
+    ungroup() %>%
     complete(effect_size, true_winner, response, `# strong`, fill = list(n = 0)) %>%
@@
     tally( wt = n ) %>%
+    ungroup() %>%
     complete(dec_type, effect_size, `# strong`, fill = list(n = 0)) %>%

### 2025-03-07 inst/manuscript/manuscript.Rmd
small changes to prepare for resubmission

@@
 However, if scientists' understanding of significance testing is truly as poor as thought, one could argue such drastic action is required.
-We present evidence using a novel reasoning task that scientists may understand the logic of significance testing better than previously thought.
-Scientists may not be as statistically-challenged as often believed; reforms should take this into account.
+We present evidence using a novel reasoning task that scientists successfully use the logic of significance testing, which has implications for statistical reform.
 
@@
 However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication).
-It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014;@Hoekstra:etal:2012]) tap understanding of the core logic of SST rather than, say, familiarity with the technical terminology used to present statistical results.
-Conceptual understanding and fluency with common representations are both important, but are distinct.
+It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014;@Hoekstra:etal:2012]) tap understanding of the core logic of SST rather than, say, the human factors of a particular practical implementation of SST (e.g., $p$ values).
+Conceptual understanding and fluency with common implementations are both important, but are distinct.
 
@@
 If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider.
-When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant given the $p$ value.
+Visual extremeness is misleading by itself; extremeness only matters relative to the null sampling distribution.
+When we break down responses by the visual extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  statistical extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant given the $p$ value.
 
@@
 We specifically set out to abstract the logic away from the typical situations in which scientists use the logic.
-This has the benefit of being helping to identify where problems might be, but the downside that generalizing the results will require further work.
+This has the benefit of helping to identify where problems might be, but has the downside that generalizing the results will require further work.
 We also cannot address other potential arguments against SST, such as philosophical ones.
@@
 Simulation-based approaches to teaching statistics have long been touted [@Cumming:etal:1995;@Rossman:Chance:2014].
-Simulation-based approaches to *studying* scientists' statistical reasoning may also profitable, particularly in studying reasoning that is difficult for participants to articulate formally.
+Simulation-based approaches to *studying* scientists' statistical reasoning may also be profitable, particularly in studying reasoning that is difficult for participants to formally articulate.
 If we are to reform statistical education and practice in the sciences, we should base that reform on diverse lines of evidence about scientists' reasoning.

### 2025-03-09 inst/manuscript/manuscript.Rmd
more small changes to prepare for resubmission

@@
 Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`).
-When the effect size is small and difficult to detect, participants experimented more before deciding.
+For the small effect sizes, null samples and experimental samples would be expected to appear quite similar; in this case, participants experimented more before deciding.
 
@@
 Filled circles show correct decisions; hollow circles, incorrect decisions.
-The two asterisks show sign errors.", out.width=".9\\columnwidth", fig.height=4.8}
+The two asterisks show sign errors.
+Some correct decisions with extreme $p$ values are omitted from the left-hand side of the graph for clarity.", out.width=".9\\columnwidth", fig.height=4.8}
 MoreyHoekstra2019::fig_response_evidence()
